### 2018-02-05 Paper-1-Vitis.Rmd
Create Paper-1-Vitis.Rmd

---
title: "Paper 1: Vitis"
output: html_notebook
---

# Intro

The size and shape (physiognomy) of woody, dicotyledonous angiosperm leaves are correlated to climate.  These relationships have been used to develop paleoclimate proxies. Many plant communities, particularly in Northern Hemisphere temperate regions, have an inverse correlation between temperature and the degree of leaf dissection and the presence of teeth, as well as tooth size and abundance. One of the underlying assumptions of leaf physiognomic-paleoclimate proxies is that leaf traits can plastically respond to climate such that that leaves reliably track changes in temperature and precipitation.  Additionally, it is also assumed that leaf traits allometrically scale as a leaf matures through the growing season, such that juvenile and mature leaves have similar leaf traits. However, little work has been done to determine how quickly leaves plastically respond to changes in climate or whether all leaves on a plant respond in the same way.

We used Digital Leaf Physiognomy (DiLP) to measure leaf characters of multiple *Vitis* species from two growing seasons with different temperatures and precipitation. The same vines were sampled each year in order to compare leaf shape on a vine-by-vine and leaf-by-leaf basis.

```{r}
library(tidyverse)
```

### 2018-02-06 Paper-1-Vitis.Rmd
Commit from RStudio

@@
 library(tidyverse)
+library(readxl)
+read_excel("R_DiLP_all_species.xlsx")
+read_excel("R_DiLP_Vitis_all_added.xlsx")
 ```
+
+# t-tests
+First, we tested whether there were significant differences between the two growing seasons for each of the ratios measured, as well as the total number of teeth.
+To do this, we ran a series of t-tests.

### 2018-02-06 Paper-1-Vitis.Rmd
Commit from Rstudio

@@
 read_excel("R_DiLP_all_species.xlsx")
+excel_sheets("R_DiLP_all_species.xlsx")
 read_excel("R_DiLP_Vitis_all_added.xlsx")
+excel_sheets("R_DiLP_Vitis_all_added.xlsx")
 ```
@@
 To do this, we ran a series of t-tests.
+
+We first tested each of the variables for *Vitis acerifolia*.
+
+```{r}
+ac_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "acerifolia.all.90")
+str(ac_all)
+ac_all$year <- as.factor(ac_all$year)
+ac_all$tooth.area <- as.numeric(ac_all$tooth.area)
+ac_all$tooth.area.perimeter <- as.numeric(ac_all$tooth.area.perimeter)
+ac_all$tooth.area.int.perimeter <- as.numeric(ac_all$tooth.area.int.perimeter)
+ac_all$average.tooth.area <- as.numeric(ac_all$average.tooth.area)
+t.test(ac_all$total.teeth~ac_all$year, var.equal = TRUE)
+t.test(ac_all$feret.diameter.ratio~ac_all$year, var.equal = TRUE)
+
+```

### 2018-02-06 Paper-1-Vitis.Rmd
Commit from RStudio

--- 2018-02-06 Paper-1-Vitis.Rmd
@@
 library(tidyverse)
+library(skimr)
 library(readxl)
 read_excel("R_DiLP_all_species.xlsx")
+excel_sheets("R_DiLP_all_species.xlsx")
 read_excel("R_DiLP_Vitis_all_added.xlsx")
+excel_sheets("R_DiLP_Vitis_all_added.xlsx")
 ```
@@
 To do this, we ran a series of t-tests.
+
+We first tested each of the variables for *Vitis acerifolia*.
+
+```{r}
+ac_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "acerifolia.all.90")
+str(ac_all)
+ac_all$year <- as.factor(ac_all$year)
+ac_all$tooth.area <- as.numeric(ac_all$tooth.area)
+ac_all$tooth.area.perimeter <- as.numeric(ac_all$tooth.area.perimeter)
+ac_all$tooth.area.int.perimeter <- as.numeric(ac_all$tooth.area.int.perimeter)
+ac_all$average.tooth.area <- as.numeric(ac_all$average.tooth.area)
+ac_all$tooth.area.blade.area <- as.numeric(ac_all$tooth.area.blade.area)
+t.test(ac_all$total.teeth~ac_all$year, var.equal = TRUE)
+t.test(ac_all$feret.diameter.ratio~ac_all$year, var.equal = TRUE)
+t.test(ac_all$feret.diameter.ratio~ac_all$year, var.equal = TRUE)
+t.test(ac_all$tooth.area.perimeter~ac_all$year, var.equal = TRUE)
+t.test(ac_all$tooth.area.int.perimeter~ac_all$year, var.equal = TRUE)
+t.test(ac_all$average.tooth.area~ac_all$year, var.equal = TRUE)
+t.test(ac_all$tooth.area.blade.area~ac_all$year, var.equal = TRUE)
+t.test(ac_all$teeth.perimeter~ac_all$year, var.equal = TRUE)
+t.test(ac_all$teeth.int.perimeter~ac_all$year, var.equal = TRUE)
+t.test(ac_all$perimeter.area~ac_all$year, var.equal = TRUE)
+t.test(ac_all$perimeter.ratio~ac_all$year, var.equal = TRUE)
+t.test(ac_all$compactness~ac_all$year, var.equal = TRUE)
+t.test(ac_all$shape.factor~ac_all$year, var.equal = TRUE)
+t.test(ac_all$teeth.blade.area~ac_all$year, var.equal = TRUE)
+```
+Only one variable showed a statistically significant different changes between growing seasons: tooth area to blade area (TABA).

### 2018-02-06 Paper-1-Vitis.Rmd
Commit from RStudio

--- 2018-02-06 Paper-1-Vitis.Rmd
@@
 To do this, we ran a series of t-tests.
+
+We first tested each of the variables for *Vitis acerifolia*.
+
+```{r}
+ac_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "acerifolia.all.90")
+str(ac_all)
+ac_all$year <- as.factor(ac_all$year)
+ac_all$tooth.area <- as.numeric(ac_all$tooth.area)
+ac_all$tooth.area.perimeter <- as.numeric(ac_all$tooth.area.perimeter)
+ac_all$tooth.area.int.perimeter <- as.numeric(ac_all$tooth.area.int.perimeter)
+ac_all$average.tooth.area <- as.numeric(ac_all$average.tooth.area)
+ac_all$tooth.area.blade.area <- as.numeric(ac_all$tooth.area.blade.area)
+t.test(ac_all$total.teeth~ac_all$year, var.equal = TRUE)
+t.test(ac_all$feret.diameter.ratio~ac_all$year, var.equal = TRUE)
+t.test(ac_all$tooth.area.perimeter~ac_all$year, var.equal = TRUE)
+t.test(ac_all$tooth.area.int.perimeter~ac_all$year, var.equal = TRUE)
+t.test(ac_all$average.tooth.area~ac_all$year, var.equal = TRUE)
+t.test(ac_all$tooth.area.blade.area~ac_all$year, var.equal = TRUE)
+t.test(ac_all$teeth.perimeter~ac_all$year, var.equal = TRUE)
+t.test(ac_all$teeth.int.perimeter~ac_all$year, var.equal = TRUE)
+t.test(ac_all$perimeter.area~ac_all$year, var.equal = TRUE)
+t.test(ac_all$perimeter.ratio~ac_all$year, var.equal = TRUE)
+t.test(ac_all$compactness~ac_all$year, var.equal = TRUE)
+t.test(ac_all$shape.factor~ac_all$year, var.equal = TRUE)
+t.test(ac_all$teeth.blade.area~ac_all$year, var.equal = TRUE)
+```
+Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).
+
+We next tested each of the variables for *Vitis aestivalis*.
+
+```{r}
+av_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "aestivalis.all.90")
+av_all$year <- as.factor(av_all$year)
+av_all$petiole.width.area <- as.numeric(av_all$petiole.width.area)
+av_all$petiole.area <- as.numeric(av_all$petiole.area)
+av_all$major.feret <- as.numeric(av_all$major.feret)
+av_all$tooth.area <- as.numeric(av_all$tooth.area)
+av_all$feret.diameter.ratio <- as.numeric(av_all$feret.diameter.ratio)
+av_all$tooth.area.perimeter <- as.numeric(av_all$tooth.area.perimeter)
+av_all$tooth.area.int.perimeter <- as.numeric(av_all$tooth.area.int.perimeter)
+av_all$average.tooth.area <- as.numeric(av_all$average.tooth.area)
+av_all$tooth.area.blade.area <- as.numeric(av_all$tooth.area.blade.area)
+str(av_all)
+t.test(av_all$total.teeth~av_all$year, var.equal = TRUE)
+t.test(av_all$feret.diameter.ratio~av_all$year, var.equal = TRUE)
+t.test(av_all$tooth.area.perimeter~av_all$year, var.equal = TRUE)
+t.test(av_all$tooth.area.int.perimeter~av_all$year, var.equal = TRUE)
+t.test(av_all$average.tooth.area~av_all$year, var.equal = TRUE)
+t.test(av_all$tooth.area.blade.area~av_all$year, var.equal = TRUE)
+t.test(av_all$teeth.perimeter~av_all$year, var.equal = TRUE)
+t.test(av_all$teeth.int.perimeter~av_all$year, var.equal = TRUE)
+t.test(av_all$perimeter.area~av_all$year, var.equal = TRUE)
+t.test(av_all$perimeter.ratio~av_all$year, var.equal = TRUE)
+t.test(av_all$compactness~av_all$year, var.equal = TRUE)
+t.test(av_all$shape.factor~av_all$year, var.equal = TRUE)
+t.test(av_all$teeth.blade.area~av_all$year, var.equal = TRUE)
+```
+Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).

### 2018-02-06 Paper-1-Vitis.Rmd
Commit from RStudio

@@
 Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).
+
+We next tested each of the variables for *Vitis amurensis*.
+
+```{r}
+am_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "amurensis.all.90")
+am_all$year <- as.factor(am_all$year)
+am_all$petiole.width <- as.numeric(am_all$petiole.width)
+am_all$petiole.width.area <- as.numeric(am_all$petiole.width.area)
+am_all$petiole.area <- as.numeric(am_all$petiole.area)
+am_all$blade.area <- as.numeric(am_all$blade.area)
+am_all$feret.diameter <- as.numeric(am_all$feret.diameter)
+am_all$major.feret <- as.numeric(am_all$major.feret)
+am_all$leaf.area <- as.numeric(am_all$leaf.area)
+am_all$perimeter <- as.numeric(am_all$perimeter)
+am_all$internal.perimeter <- as.numeric(am_all$internal.perimeter)
+am_all$feret.diameter.ratio <- as.numeric(am_all$feret.diameter.ratio)
+am_all$tooth.area.perimeter <- as.numeric(am_all$tooth.area.perimeter)
+am_all$tooth.area.int.perimeter <- as.numeric(am_all$tooth.area.int.perimeter)
+am_all$tooth.area.blade.area <- as.numeric(am_all$tooth.area.blade.area)
+am_all$teeth.perimeter <- as.numeric(am_all$teeth.perimeter)
+am_all$teeth.int.perimeter <- as.numeric(am_all$teeth.int.perimeter)
+am_all$perimeter.area <- as.numeric(am_all$perimeter.area)
+am_all$perimeter.ratio <- as.numeric(am_all$perimeter.ratio)
+am_all$compactness <- as.numeric(am_all$compactness)
+am_all$shape.factor <- as.numeric(am_all$shape.factor)
+am_all$teeth.blade.area <- as.numeric(am_all$teeth.blade.area)
+str(am_all)
+t.test(am_all$total.teeth~am_all$year, var.equal = TRUE)
+t.test(am_all$feret.diameter.ratio~am_all$year, var.equal = TRUE)
+t.test(am_all$tooth.area.perimeter~am_all$year, var.equal = TRUE)
+t.test(am_all$tooth.area.int.perimeter~am_all$year, var.equal = TRUE)
+t.test(am_all$average.tooth.area~am_all$year, var.equal = TRUE)
+t.test(am_all$tooth.area.blade.area~am_all$year, var.equal = TRUE)
+t.test(am_all$teeth.perimeter~am_all$year, var.equal = TRUE)
+t.test(am_all$teeth.int.perimeter~am_all$year, var.equal = TRUE)
+t.test(am_all$perimeter.area~am_all$year, var.equal = TRUE)
+t.test(am_all$perimeter.ratio~am_all$year, var.equal = TRUE)
+t.test(am_all$compactness~am_all$year, var.equal = TRUE)
+t.test(am_all$shape.factor~am_all$year, var.equal = TRUE)
+t.test(am_all$teeth.blade.area~am_all$year, var.equal = TRUE)
+```

### 2018-02-06 Paper-1-Vitis.Rmd
Commit from RStudio

--- 2018-02-06 Paper-1-Vitis.Rmd
@@
 ac_all$tooth.area.blade.area <- as.numeric(ac_all$tooth.area.blade.area)
+skim(ac_all)
 t.test(ac_all$total.teeth~ac_all$year, var.equal = TRUE)
-t.test(ac_all$feret.diameter.ratio~ac_all$year, var.equal = TRUE)
 t.test(ac_all$feret.diameter.ratio~ac_all$year, var.equal = TRUE)
@@
 ```
-Only one variable showed a statistically significant different changes between growing seasons: tooth area to blade area (TABA).
+Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).
+
+We next tested each of the variables for *Vitis aestivalis*.
+
+```{r}
+av_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "aestivalis.all.90")
+av_all$year <- as.factor(av_all$year)
+av_all$petiole.width.area <- as.numeric(av_all$petiole.width.area)
+av_all$petiole.area <- as.numeric(av_all$petiole.area)
+av_all$major.feret <- as.numeric(av_all$major.feret)
+av_all$tooth.area <- as.numeric(av_all$tooth.area)
+av_all$feret.diameter.ratio <- as.numeric(av_all$feret.diameter.ratio)
+av_all$tooth.area.perimeter <- as.numeric(av_all$tooth.area.perimeter)
+av_all$tooth.area.int.perimeter <- as.numeric(av_all$tooth.area.int.perimeter)
+av_all$average.tooth.area <- as.numeric(av_all$average.tooth.area)
+av_all$tooth.area.blade.area <- as.numeric(av_all$tooth.area.blade.area)
+str(av_all)
+skim(av_all)
+t.test(av_all$total.teeth~av_all$year, var.equal = TRUE)
+t.test(av_all$feret.diameter.ratio~av_all$year, var.equal = TRUE)
+t.test(av_all$tooth.area.perimeter~av_all$year, var.equal = TRUE)
+t.test(av_all$tooth.area.int.perimeter~av_all$year, var.equal = TRUE)
+t.test(av_all$average.tooth.area~av_all$year, var.equal = TRUE)
+t.test(av_all$tooth.area.blade.area~av_all$year, var.equal = TRUE)
+t.test(av_all$teeth.perimeter~av_all$year, var.equal = TRUE)
+t.test(av_all$teeth.int.perimeter~av_all$year, var.equal = TRUE)
+t.test(av_all$perimeter.area~av_all$year, var.equal = TRUE)
+t.test(av_all$perimeter.ratio~av_all$year, var.equal = TRUE)
+t.test(av_all$compactness~av_all$year, var.equal = TRUE)
+t.test(av_all$shape.factor~av_all$year, var.equal = TRUE)
+t.test(av_all$teeth.blade.area~av_all$year, var.equal = TRUE)
+```
+Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).
+
+We next tested each of the variables for *Vitis amurensis*.
+
+```{r}
+am_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "amurensis.all.90")
+am_all$year <- as.factor(am_all$year)
+am_all$petiole.width <- as.numeric(am_all$petiole.width)
+am_all$petiole.width.area <- as.numeric(am_all$petiole.width.area)
+am_all$petiole.area <- as.numeric(am_all$petiole.area)
+am_all$blade.area <- as.numeric(am_all$blade.area)
+am_all$feret.diameter <- as.numeric(am_all$feret.diameter)
+am_all$major.feret <- as.numeric(am_all$major.feret)
+am_all$leaf.area <- as.numeric(am_all$leaf.area)
+am_all$perimeter <- as.numeric(am_all$perimeter)
+am_all$internal.perimeter <- as.numeric(am_all$internal.perimeter)
+am_all$feret.diameter.ratio <- as.numeric(am_all$feret.diameter.ratio)
+am_all$tooth.area.perimeter <- as.numeric(am_all$tooth.area.perimeter)
+am_all$tooth.area.int.perimeter <- as.numeric(am_all$tooth.area.int.perimeter)
+am_all$tooth.area.blade.area <- as.numeric(am_all$tooth.area.blade.area)
+am_all$teeth.perimeter <- as.numeric(am_all$teeth.perimeter)
+am_all$teeth.int.perimeter <- as.numeric(am_all$teeth.int.perimeter)
+am_all$perimeter.area <- as.numeric(am_all$perimeter.area)
+am_all$perimeter.ratio <- as.numeric(am_all$perimeter.ratio)
+am_all$compactness <- as.numeric(am_all$compactness)
+am_all$shape.factor <- as.numeric(am_all$shape.factor)
+am_all$teeth.blade.area <- as.numeric(am_all$teeth.blade.area)
+str(am_all)
+skim(am_all)
+t.test(am_all$total.teeth~am_all$year, var.equal = TRUE)
+t.test(am_all$feret.diameter.ratio~am_all$year, var.equal = TRUE)
+t.test(am_all$tooth.area.perimeter~am_all$year, var.equal = TRUE)
+t.test(am_all$tooth.area.int.perimeter~am_all$year, var.equal = TRUE)
+t.test(am_all$average.tooth.area~am_all$year, var.equal = TRUE)
+t.test(am_all$tooth.area.blade.area~am_all$year, var.equal = TRUE)
+t.test(am_all$teeth.perimeter~am_all$year, var.equal = TRUE)
+t.test(am_all$teeth.int.perimeter~am_all$year, var.equal = TRUE)
+t.test(am_all$perimeter.area~am_all$year, var.equal = TRUE)
+t.test(am_all$perimeter.ratio~am_all$year, var.equal = TRUE)
+t.test(am_all$compactness~am_all$year, var.equal = TRUE)
+t.test(am_all$shape.factor~am_all$year, var.equal = TRUE)
+t.test(am_all$teeth.blade.area~am_all$year, var.equal = TRUE)
+```
+
+Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF).
+
+We finally tested each of the variables for *Vitis riparia*.
+
+```{r}
+r_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "riparia.all.90")
+r_all$year <- as.factor(r_all$year)
+r_all$petiole.width <- as.numeric(r_all$petiole.width)
+r_all$petiole.width.area <- as.numeric(r_all$petiole.width.area)
+r_all$petiole.area <- as.numeric(r_all$petiole.area)
+r_all$internal.perimeter <- as.numeric(r_all$internal.perimeter)
+r_all$tooth.area <- as.numeric(r_all$tooth.area)
+r_all$tooth.area.perimeter <- as.numeric(r_all$tooth.area.perimeter)
+r_all$tooth.area.int.perimeter <- as.numeric(r_all$tooth.area.int.perimeter)
+r_all$average.tooth.area <- as.numeric(r_all$average.tooth.area)
+r_all$tooth.area.blade.area <- as.numeric(r_all$tooth.area.blade.area)
+r_all$teeth.int.perimeter <- as.numeric(r_all$teeth.int.perimeter)
+r_all$perimeter.ratio <- as.numeric(r_all$perimeter.ratio)
+r_all$teeth.blade.area <- as.numeric(r_all$teeth.blade.area)
+str(r_all)
+skim(r_all)
+t.test(r_all$total.teeth~r_all$year, var.equal = TRUE)
+t.test(r_all$feret.diameter.ratio~r_all$year, var.equal = TRUE)
+t.test(r_all$tooth.area.perimeter~r_all$year, var.equal = TRUE)
+t.test(r_all$tooth.area.int.perimeter~r_all$year, var.equal = TRUE)
+t.test(r_all$average.tooth.area~r_all$year, var.equal = TRUE)
+t.test(r_all$tooth.area.blade.area~r_all$year, var.equal = TRUE)
+t.test(r_all$teeth.perimeter~r_all$year, var.equal = TRUE)
+t.test(r_all$teeth.int.perimeter~r_all$year, var.equal = TRUE)
+t.test(r_all$perimeter.area~r_all$year, var.equal = TRUE)
+t.test(r_all$perimeter.ratio~r_all$year, var.equal = TRUE)
+t.test(r_all$compactness~r_all$year, var.equal = TRUE)
+t.test(r_all$shape.factor~r_all$year, var.equal = TRUE)
+t.test(r_all$teeth.blade.area~r_all$year, var.equal = TRUE)
+```
+
+Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).

### 2018-02-08 Paper-1-Vitis.Rmd
Commit from RStudio

@@
 Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).
+
+# Principal Components Analysis (PCA)
+
+```{r}
+library(pcaMethods)
+excel_sheets("R_DiLP_Vitis_all_added.xlsx")
+```
+
+First, we applied a PCA to *Vitis acerifolia*.
+```{r}
+ac_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "acerifolia.all.ln")
+ac_ln_mat <- data.matrix(ac_ln, rownames.force = NA)
+ac_ln_fixed <- ac_ln_mat[,-(1:3)]
+ac_ln_fixed[ac_ln_fixed == 0] <- NA
+pca_ac_all <- pca(ac_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
+summary(pca_ac_all)
+loadings(pca_ac_all)
+scores_pca_ac <- scores(pca_ac_all)
+scores_pca_ac_df <- as.data.frame(scores_pca_ac)
+str(scores_pca_ac_df)
+#ggplot(scores_pca_ac_df, aes(x = PC1, y = PC2, label = rownames(scores_pca_ac_df))) +
+  #geom_col(x = PC1, y = PC2) +
+  #ggtitle("PCA plot of Vitis acerifolia")
+#attach row names
+```

### 2018-02-09 Paper-1-Vitis.Rmd
Commit from RStudio

--- 2018-02-06 Paper-1-Vitis.Rmd
@@
 Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).
+
+# Principal Components Analysis (PCA)
+
+```{r}
+library(pcaMethods)
+excel_sheets("R_DiLP_Vitis_all_added.xlsx")
+```
+
+First, we applied a PCA to *Vitis acerifolia*.
+```{r}
+ac_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "acerifolia.all.ln")
+ac_ln_mat <- data.matrix(ac_ln, rownames.force = NA)
+ac_ln_fixed <- ac_ln_mat[,-(1:3)]
+ac_ln_fixed[ac_ln_fixed == 0] <- NA
+pca_ac_all <- pca(ac_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
+bpca_ac_ll <- pca(ac_ln_fixed, method = "bpca", nPcs = 5)
+summary(bpca_ac_ll)
+loadings(bpca_ac_ll)
+#summary(pca_ac_all)
+#loadings(pca_ac_all)
+#print(pca_ac_all)
+#slplot(pca_ac_all)
+#loadings_pca_ac <- loadings(pca_ac_all)
+#loadings_pca_ac_df <- as.data.frame(loadings_pca_ac)
+#str(loadings_pca_ac_df)
+#ggplot(loadings_pca_ac_df, aes(x = PC1, y = PC2, label = rownames(loadings_pca_ac_df))) +
+  #geom_col(x = PC1, y = PC2) +
+  #ggtitle("PCA plot of Vitis acerifolia")
+#attach row names
+beep(sound = 3, expr = NULL)
+```

### 2018-02-12 Paper-1-Vitis.Rmd
Commit from RStudio

--- 2018-02-06 Paper-1-Vitis.Rmd
@@
 Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).
+
+# Principal Components Analysis (PCA)
+
+```{r}
+library(pcaMethods)
+excel_sheets("R_DiLP_Vitis_all_added.xlsx")
+```
+
+First, we applied a PCA to *Vitis acerifolia*.
+```{r}
+ac_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "acerifolia.all.ln")
+ac_ln_mat <- data.matrix(ac_ln, rownames.force = NA)
+ac_ln_fixed <- ac_ln_mat[,-(1:3)]
+ac_ln_fixed[ac_ln_fixed == 0] <- NA
+#pca_ac_all <- pca(ac_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
+bpca_ac_ll <- pca(ac_ln_fixed, method = "bpca", nPcs = 5)
+summary(bpca_ac_ll)
+loadings(bpca_ac_ll)
+#summary(pca_ac_all)
+#loadings(pca_ac_all)
+print(bpca_ac_ll)
+#print(pca_ac_all)
+#slplot(pca_ac_all)
+#loadings_pca_ac <- loadings(pca_ac_all)
+#loadings_pca_ac_df <- as.data.frame(loadings_pca_ac)
+#str(loadings_pca_ac_df)
+#ggplot(loadings_pca_ac_df, aes(x = PC1, y = PC2, label = rownames(loadings_pca_ac_df))) +
+  #geom_col(x = PC1, y = PC2) +
+  #ggtitle("PCA plot of Vitis acerifolia")
+#attach row names
+beep(sound = 3, expr = NULL)
+```

### 2018-02-12 Paper-1-Vitis.Rmd
Commit from RStudio

--- 2018-02-08 Paper-1-Vitis.Rmd
@@
 library(tidyverse)
+library(tidyr)
 library(skimr)
@@
 ac_ln_fixed[ac_ln_fixed == 0] <- NA
-pca_ac_all <- pca(ac_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
-summary(pca_ac_all)
-loadings(pca_ac_all)
-scores_pca_ac <- scores(pca_ac_all)
-scores_pca_ac_df <- as.data.frame(scores_pca_ac)
-str(scores_pca_ac_df)
-#ggplot(scores_pca_ac_df, aes(x = PC1, y = PC2, label = rownames(scores_pca_ac_df))) +
+#pca_ac_all <- pca(ac_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
+bpca_ac_ll <- pca(ac_ln_fixed, method = "bpca", nPcs = 5)
+summary(bpca_ac_ll)
+loadings(bpca_ac_ll)
+#summary(pca_ac_all)
+#loadings(pca_ac_all)
+print(bpca_ac_ll)
+#print(pca_ac_all)
+#slplot(pca_ac_all)
+#loadings_pca_ac <- loadings(pca_ac_all)
+#loadings_pca_ac_df <- as.data.frame(loadings_pca_ac)
+#str(loadings_pca_ac_df)
+#ggplot(loadings_pca_ac_df, aes(x = PC1, y = PC2, label = rownames(loadings_pca_ac_df))) +
   #geom_col(x = PC1, y = PC2) +
@@
 ```
+
+We then applied a PCA to *Vitis aestivalis*.
+
+```{r}
+av_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "aestivalis.all.ln")
+av_ln_mat <- data.matrix(av_ln, rownames.force = NA)
+av_ln_fixed <- av_ln_mat[,-(1:3)]
+av_ln_fixed[av_ln_fixed == 0] <- NA
+#pca_av_all <- pca(av_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
+bpca_av_ll <- pca(av_ln_fixed, method = "bpca", nPcs = 5)
+summary(bpca_av_ll)
+loadings(bpca_av_ll)
+#summary(pca_av_all)
+#loadings(pca_av_all)
+print(bpca_av_ll)
+#print(pca_av_all)
+#slplot(pca_av_all)
+```
+
+We then applied a PCA to *Vitis amurensis*.
+
+```{r}
+am_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "amurensis.all.ln")
+am_ln_mat <- data.matrix(am_ln, rownames.force = NA)
+am_ln_fixed <- am_ln_mat[,-(1:3)]
+am_ln_fixed[am_ln_fixed == 0] <- NA
+#pca_am_all <- pca(am_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
+bpca_am_ll <- pca(am_ln_fixed, method = "bpca", nPcs = 5)
+summary(bpca_am_ll)
+loadings(bpca_am_ll)
+#summary(pca_am_all)
+#loadings(pca_am_all)
+print(bpca_am_ll)
+#print(pca_am_all)
+#slplot(pca_am_all)
+```
+
+Finally, we applied a PCA to *Vitis riparia*.
+
+```{r}
+r_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "riparia.all.ln")
+r_ln_mat <- data.matrix(r_ln, rownames.force = NA)
+r_ln_fixed <- r_ln_mat[,-(1:3)]
+r_ln_fixed[r_ln_fixed == 0] <- NA
+#pca_r_all <- pca(r_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
+bpca_r_ll <- pca(r_ln_fixed, method = "bpca", nPcs = 5)
+summary(bpca_r_ll)
+loadings(bpca_r_ll)
+#summary(pca_r_all)
+#loadings(pca_r_all)
+print(bpca_r_ll)
+#print(pca_r_all)
+#slplot(pca_r_all)
+beep(sound = 3, expr = NULL)
+```
+
+# Generalized Linear Modeling
+
+First, we tested *Vitis acerifolia*.
+
+```{r}
+ac_glm <- glm(leaf~total.teeth, data = ac_all, family = gaussian())
+summary(ac_glm)
+plot(ac_glm)
+```

### 2018-02-14 Paper-1-Vitis.Rmd
Commit from RStudio

@@
 #slplot(pca_r_all)
-beep(sound = 3, expr = NULL)
 ```
@@
 
+We ran a GLM for leaf number and total teeth.
+
 ```{r}
-ac_glm <- glm(leaf~total.teeth, data = ac_all, family = gaussian())
-summary(ac_glm)
-plot(ac_glm)
+skim(ac_all)
+ac_tt_glm <- glm(leaf~total.teeth, data = ac_all, family = gaussian())
+summary(ac_tt_glm)
+plot(ac_tt_glm)
 ```
+
+We ran a GLM for leaf number and feret diameter ratio.
+
+```{r}
+ac_fdr_glm <- glm(leaf~feret.diameter.ratio, data = ac_all, family = gaussian())
+summary(ac_fdr_glm)
+plot(ac_fdr_glm)
+```
+
+We ran a GLM for leaf number and tooth area:perimeter.
+
+```{r}
+ac_tap_glm <- glm(leaf~tooth.area.perimeter, data = ac_all, family = gaussian())
+summary(ac_tap_glm)
+plot(ac_tap_glm)
+```
+
+We ran a GLM for leaf number and tooth area:internal perimeter.
+
+```{r}
+ac_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = ac_all, family = gaussian())
+summary(ac_taip_glm)
+plot(ac_taip_glm)
+```
+
+We ran a GLM for leaf number and average tooth area.
+
+```{r}
+ac_ata_glm <- glm(leaf~average.tooth.area, data = ac_all, family = gaussian())
+summary(ac_ata_glm)
+plot(ac_ata_glm)
+```
+
+We ran a GLM for leaf number and tooth area:blade area..
+
+```{r}
+ac_taba_glm <- glm(leaf~tooth.area.blade.area, data = ac_all, family = gaussian())
+summary(ac_taba_glm)
+plot(ac_taba_glm)
+```
+
+We ran a GLM for leaf number and teeth:perimeter.
+
+```{r}
+ac_tp_glm <- glm(leaf~teeth.perimeter, data = ac_all, family = gaussian())
+summary(ac_tp_glm)
+plot(ac_tp_glm)
+```
+
+We ran a GLM for leaf number and teeth:internal perimeter.
+
+```{r}
+ac_tip_glm <- glm(leaf~teeth.int.perimeter, data = ac_all, family = gaussian())
+summary(ac_tip_glm)
+plot(ac_tip_glm)
+```
+
+We ran a GLM for leaf number and perimeter:area.
+
+```{r}
+ac_pa_glm <- glm(leaf~perimeter.area, data = ac_all, family = gaussian())
+summary(ac_pa_glm)
+plot(ac_pa_glm)
+```
+
+We ran a GLM for leaf number and perimeter ratio.
+
+```{r}
+ac_pr_glm <- glm(leaf~perimeter.ratio, data = ac_all, family = gaussian())
+summary(ac_pr_glm)
+plot(ac_pr_glm)
+```
+
+We ran a GLM for leaf number and compactness.
+
+```{r}
+ac_cp_glm <- glm(leaf~compactness, data = ac_all, family = gaussian())
+summary(ac_cp_glm)
+plot(ac_cp_glm)
+```
+
+We ran a GLM for leaf number and shape factor.
+
+```{r}
+ac_sf_glm <- glm(leaf~shape.factor, data = ac_all, family = gaussian())
+summary(ac_sf_glm)
+plot(ac_sf_glm)
+```
+
+We ran a GLM for leaf number and teeth:blade area.
+
+```{r}
+ac_tba_glm <- glm(leaf~teeth.blade.area, data = ac_all, family = gaussian())
+summary(ac_tba_glm)
+plot(ac_tba_glm)
+beep(sound = 3, expr = NULL)
+```
+
+We ran a GLM for leaf number and multiple variables. 
+```{r}
+ac_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = ac_all, family = gaussian())
+summary(ac_clim_glm)
+plot(ac_clim_glm)
+```

### 2018-02-21 Paper-1-Vitis.Rmd
Commit from RStudio

@@
 plot(ac_tba_glm)
-beep(sound = 3, expr = NULL)
 ```
@@
 plot(ac_clim_glm)
+beep(sound = 3, expr = NULL)
 ```

### 2018-02-21 Paper-1-Vitis.Rmd
Commit from RStudio

--- 2018-02-14 Paper-1-Vitis.Rmd
@@
 ```
+
+We ran a GLM for leaf number and all variables. 
+```{r}
+ac_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = ac_all, family = gaussian())
+summary(ac_all_glm)
+plot(ac_all_glm)
+```
+
+We tested *Vitis aestivalis*.
+
+We ran a GLM for leaf number and total teeth.
+
+```{r}
+skim(av_all)
+av_tt_glm <- glm(leaf~total.teeth, data = av_all, family = gaussian())
+summary(av_tt_glm)
+plot(av_tt_glm)
+```
+
+We ran a GLM for leaf number and feret diameter ratio.
+
+```{r}
+av_fdr_glm <- glm(leaf~feret.diameter.ratio, data = av_all, family = gaussian())
+summary(av_fdr_glm)
+plot(av_fdr_glm)
+```
+
+We ran a GLM for leaf number and tooth area:perimeter.
+
+```{r}
+av_tap_glm <- glm(leaf~tooth.area.perimeter, data = av_all, family = gaussian())
+summary(av_tap_glm)
+plot(av_tap_glm)
+```
+
+We ran a GLM for leaf number and tooth area:internal perimeter.
+
+```{r}
+av_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = av_all, family = gaussian())
+summary(av_taip_glm)
+plot(av_taip_glm)
+```
+
+We ran a GLM for leaf number and average tooth area.
+
+```{r}
+av_ata_glm <- glm(leaf~average.tooth.area, data = av_all, family = gaussian())
+summary(av_ata_glm)
+plot(av_ata_glm)
+```
+
+We ran a GLM for leaf number and tooth area:blade area..
+
+```{r}
+av_taba_glm <- glm(leaf~tooth.area.blade.area, data = av_all, family = gaussian())
+summary(av_taba_glm)
+plot(av_taba_glm)
+```
+
+We ran a GLM for leaf number and teeth:perimeter.
+
+```{r}
+av_tp_glm <- glm(leaf~teeth.perimeter, data = av_all, family = gaussian())
+summary(av_tp_glm)
+plot(av_tp_glm)
+```
+
+We ran a GLM for leaf number and teeth:internal perimeter.
+
+```{r}
+av_tip_glm <- glm(leaf~teeth.int.perimeter, data = av_all, family = gaussian())
+summary(av_tip_glm)
+plot(av_tip_glm)
+```
+
+We ran a GLM for leaf number and perimeter:area.
+
+```{r}
+av_pa_glm <- glm(leaf~perimeter.area, data = av_all, family = gaussian())
+summary(av_pa_glm)
+plot(av_pa_glm)
+```
+
+We ran a GLM for leaf number and perimeter ratio.
+
+```{r}
+av_pr_glm <- glm(leaf~perimeter.ratio, data = av_all, family = gaussian())
+summary(av_pr_glm)
+plot(av_pr_glm)
+```
+
+We ran a GLM for leaf number and compactness.
+
+```{r}
+av_cp_glm <- glm(leaf~compactness, data = av_all, family = gaussian())
+summary(av_cp_glm)
+plot(av_cp_glm)
+```
+
+We ran a GLM for leaf number and shape factor.
+
+```{r}
+av_sf_glm <- glm(leaf~shape.factor, data = av_all, family = gaussian())
+summary(av_sf_glm)
+plot(av_sf_glm)
+```
+
+We ran a GLM for leaf number and teeth:blade area.
+
+```{r}
+av_tba_glm <- glm(leaf~teeth.blade.area, data = av_all, family = gaussian())
+summary(av_tba_glm)
+plot(av_tba_glm)
+```
+
+We ran a GLM for leaf number and multiple variables. 
+```{r}
+av_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = av_all, family = gaussian())
+summary(av_clim_glm)
+plot(av_clim_glm)
+```
+
+We ran a GLM for leaf number and all variables. 
+```{r}
+av_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = av_all, family = gaussian())
+summary(av_all_glm)
+plot(av_all_glm)
+```
+
+we tested *Vitis amurensis*.
+
+We ran a GLM for leaf number and total teeth.
+
+```{r}
+skim(am_all)
+am_tt_glm <- glm(leaf~total.teeth, data = am_all, family = gaussian())
+summary(am_tt_glm)
+plot(am_tt_glm)
+```
+
+We ran a GLM for leaf number and feret diameter ratio.
+
+```{r}
+am_fdr_glm <- glm(leaf~feret.diameter.ratio, data = am_all, family = gaussian())
+summary(am_fdr_glm)
+plot(am_fdr_glm)
+```
+
+We ran a GLM for leaf number and tooth area:perimeter.
+
+```{r}
+am_tap_glm <- glm(leaf~tooth.area.perimeter, data = am_all, family = gaussian())
+summary(am_tap_glm)
+plot(am_tap_glm)
+```
+
+We ran a GLM for leaf number and tooth area:internal perimeter.
+
+```{r}
+am_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = am_all, family = gaussian())
+summary(am_taip_glm)
+plot(am_taip_glm)
+```
+
+We ran a GLM for leaf number and average tooth area.
+
+```{r}
+am_ata_glm <- glm(leaf~average.tooth.area, data = am_all, family = gaussian())
+summary(am_ata_glm)
+plot(am_ata_glm)
+```
+
+We ran a GLM for leaf number and tooth area:blade area..
+
+```{r}
+am_taba_glm <- glm(leaf~tooth.area.blade.area, data = am_all, family = gaussian())
+summary(am_taba_glm)
+plot(am_taba_glm)
+```
+
+We ran a GLM for leaf number and teeth:perimeter.
+
+```{r}
+am_tp_glm <- glm(leaf~teeth.perimeter, data = am_all, family = gaussian())
+summary(am_tp_glm)
+plot(am_tp_glm)
+```
+
+We ran a GLM for leaf number and teeth:internal perimeter.
+
+```{r}
+am_tip_glm <- glm(leaf~teeth.int.perimeter, data = am_all, family = gaussian())
+summary(am_tip_glm)
+plot(am_tip_glm)
+```
+
+We ran a GLM for leaf number and perimeter:area.
+
+```{r}
+am_pa_glm <- glm(leaf~perimeter.area, data = am_all, family = gaussian())
+summary(am_pa_glm)
+plot(am_pa_glm)
+```
+
+We ran a GLM for leaf number and perimeter ratio.
+
+```{r}
+am_pr_glm <- glm(leaf~perimeter.ratio, data = am_all, family = gaussian())
+summary(am_pr_glm)
+plot(am_pr_glm)
+```
+
+We ran a GLM for leaf number and compactness.
+
+```{r}
+am_cp_glm <- glm(leaf~compactness, data = am_all, family = gaussian())
+summary(am_cp_glm)
+plot(am_cp_glm)
+```
+
+We ran a GLM for leaf number and shape factor.
+
+```{r}
+am_sf_glm <- glm(leaf~shape.factor, data = am_all, family = gaussian())
+summary(am_sf_glm)
+plot(am_sf_glm)
+```
+
+We ran a GLM for leaf number and teeth:blade area.
+
+```{r}
+am_tba_glm <- glm(leaf~teeth.blade.area, data = am_all, family = gaussian())
+summary(am_tba_glm)
+plot(am_tba_glm)
+```
+
+We ran a GLM for leaf number and multiple variables. 
+```{r}
+am_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = am_all, family = gaussian())
+summary(am_clim_glm)
+plot(am_clim_glm)
+```
+
+We ran a GLM for leaf number and all variables. 
+```{r}
+am_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = am_all, family = gaussian())
+summary(am_all_glm)
+plot(am_all_glm)
+```
+
+We tested *Vitis riparia*.
+
+We ran a GLM for leaf number and total teeth.
+
+```{r}
+skim(r_all)
+r_tt_glm <- glm(leaf~total.teeth, data = r_all, family = gaussian())
+summary(r_tt_glm)
+plot(r_tt_glm)
+```
+
+We ran a GLM for leaf number and feret diameter ratio.
+
+```{r}
+r_fdr_glm <- glm(leaf~feret.diameter.ratio, data = r_all, family = gaussian())
+summary(r_fdr_glm)
+plot(r_fdr_glm)
+```
+
+We ran a GLM for leaf number and tooth area:perimeter.
+
+```{r}
+r_tap_glm <- glm(leaf~tooth.area.perimeter, data = r_all, family = gaussian())
+summary(r_tap_glm)
+plot(r_tap_glm)
+```
+
+We ran a GLM for leaf number and tooth area:internal perimeter.
+
+```{r}
+r_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = r_all, family = gaussian())
+summary(r_taip_glm)
+plot(r_taip_glm)
+```
+
+We ran a GLM for leaf number and average tooth area.
+
+```{r}
+r_ata_glm <- glm(leaf~average.tooth.area, data = r_all, family = gaussian())
+summary(r_ata_glm)
+plot(r_ata_glm)
+```
+
+We ran a GLM for leaf number and tooth area:blade area.
+
+```{r}
+r_taba_glm <- glm(leaf~tooth.area.blade.area, data = r_all, family = gaussian())
+summary(r_taba_glm)
+plot(r_taba_glm)
+```
+
+We ran a GLM for leaf number and teeth:perimeter.
+
+```{r}
+r_tp_glm <- glm(leaf~teeth.perimeter, data = r_all, family = gaussian())
+summary(r_tp_glm)
+plot(r_tp_glm)
+```
+
+We ran a GLM for leaf number and teeth:internal perimeter.
+
+```{r}
+r_tip_glm <- glm(leaf~teeth.int.perimeter, data = r_all, family = gaussian())
+summary(r_tip_glm)
+plot(r_tip_glm)
+```
+
+We ran a GLM for leaf number and perimeter:area.
+
+```{r}
+r_pa_glm <- glm(leaf~perimeter.area, data = r_all, family = gaussian())
+summary(r_pa_glm)
+plot(r_pa_glm)
+```
+
+We ran a GLM for leaf number and perimeter ratio.
+
+```{r}
+r_pr_glm <- glm(leaf~perimeter.ratio, data = r_all, family = gaussian())
+summary(r_pr_glm)
+plot(r_pr_glm)
+```
+
+We ran a GLM for leaf number and compactness.
+
+```{r}
+r_cp_glm <- glm(leaf~compactness, data = r_all, family = gaussian())
+summary(r_cp_glm)
+plot(r_cp_glm)
+```
+
+We ran a GLM for leaf number and shape factor.
+
+```{r}
+r_sf_glm <- glm(leaf~shape.factor, data = r_all, family = gaussian())
+summary(r_sf_glm)
+plot(r_sf_glm)
+```
+
+We ran a GLM for leaf number and teeth:blade area.
+
+```{r}
+r_tba_glm <- glm(leaf~teeth.blade.area, data = r_all, family = gaussian())
+summary(r_tba_glm)
+plot(r_tba_glm)
+```
+
+We ran a GLM for leaf number and multiple variables. 
+```{r}
+r_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = r_all, family = gaussian())
+summary(r_clim_glm)
+plot(r_clim_glm)
+```
+
+We ran a GLM for leaf number and all variables. 
+```{r}
+r_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = r_all, family = gaussian())
+summary(r_all_glm)
+plot(r_all_glm)
+beep(sound = 3, expr = NULL)
+```

### 2018-10-10 Vitis_GLM.Rmd
lm attempt 1

--- 2018-02-21 Paper-1-Vitis.Rmd
@@
 ---
-title: "Paper 1: Vitis"
+title: "Paper 1 GLM"
 output: html_notebook
@@
 
-```{r}
-library(tidyverse)
-library(tidyr)
+```{r setup}
+library(readxl)
+library(dplyr)
 library(skimr)
-library(readxl)
-read_excel("R_DiLP_all_species.xlsx")
-excel_sheets("R_DiLP_all_species.xlsx")
-read_excel("R_DiLP_Vitis_all_added.xlsx")
-excel_sheets("R_DiLP_Vitis_all_added.xlsx")
+library(ggplot2)
+library(segmented)
 ```
 
-# t-tests
-First, we tested whether there were significant differences between the two growing seasons for each of the ratios measured, as well as the total number of teeth.
-To do this, we ran a series of t-tests.
-
-We first tested each of the variables for *Vitis acerifolia*.
-
-```{r}
-ac_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "acerifolia.all.90")
-str(ac_all)
-ac_all$year <- as.factor(ac_all$year)
-ac_all$tooth.area <- as.numeric(ac_all$tooth.area)
-ac_all$tooth.area.perimeter <- as.numeric(ac_all$tooth.area.perimeter)
-ac_all$tooth.area.int.perimeter <- as.numeric(ac_all$tooth.area.int.perimeter)
-ac_all$average.tooth.area <- as.numeric(ac_all$average.tooth.area)
-ac_all$tooth.area.blade.area <- as.numeric(ac_all$tooth.area.blade.area)
-skim(ac_all)
-t.test(ac_all$total.teeth~ac_all$year, var.equal = TRUE)
-t.test(ac_all$feret.diameter.ratio~ac_all$year, var.equal = TRUE)
-t.test(ac_all$tooth.area.perimeter~ac_all$year, var.equal = TRUE)
-t.test(ac_all$tooth.area.int.perimeter~ac_all$year, var.equal = TRUE)
-t.test(ac_all$average.tooth.area~ac_all$year, var.equal = TRUE)
-t.test(ac_all$tooth.area.blade.area~ac_all$year, var.equal = TRUE)
-t.test(ac_all$teeth.perimeter~ac_all$year, var.equal = TRUE)
-t.test(ac_all$teeth.int.perimeter~ac_all$year, var.equal = TRUE)
-t.test(ac_all$perimeter.area~ac_all$year, var.equal = TRUE)
-t.test(ac_all$perimeter.ratio~ac_all$year, var.equal = TRUE)
-t.test(ac_all$compactness~ac_all$year, var.equal = TRUE)
-t.test(ac_all$shape.factor~ac_all$year, var.equal = TRUE)
-t.test(ac_all$teeth.blade.area~ac_all$year, var.equal = TRUE)
-```
-Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).
-
-We next tested each of the variables for *Vitis aestivalis*.
-
-```{r}
-av_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "aestivalis.all.90")
-av_all$year <- as.factor(av_all$year)
-av_all$petiole.width.area <- as.numeric(av_all$petiole.width.area)
-av_all$petiole.area <- as.numeric(av_all$petiole.area)
-av_all$major.feret <- as.numeric(av_all$major.feret)
-av_all$tooth.area <- as.numeric(av_all$tooth.area)
-av_all$feret.diameter.ratio <- as.numeric(av_all$feret.diameter.ratio)
-av_all$tooth.area.perimeter <- as.numeric(av_all$tooth.area.perimeter)
-av_all$tooth.area.int.perimeter <- as.numeric(av_all$tooth.area.int.perimeter)
-av_all$average.tooth.area <- as.numeric(av_all$average.tooth.area)
-av_all$tooth.area.blade.area <- as.numeric(av_all$tooth.area.blade.area)
-str(av_all)
-skim(av_all)
-t.test(av_all$total.teeth~av_all$year, var.equal = TRUE)
-t.test(av_all$feret.diameter.ratio~av_all$year, var.equal = TRUE)
-t.test(av_all$tooth.area.perimeter~av_all$year, var.equal = TRUE)
-t.test(av_all$tooth.area.int.perimeter~av_all$year, var.equal = TRUE)
-t.test(av_all$average.tooth.area~av_all$year, var.equal = TRUE)
-t.test(av_all$tooth.area.blade.area~av_all$year, var.equal = TRUE)
-t.test(av_all$teeth.perimeter~av_all$year, var.equal = TRUE)
-t.test(av_all$teeth.int.perimeter~av_all$year, var.equal = TRUE)
-t.test(av_all$perimeter.area~av_all$year, var.equal = TRUE)
-t.test(av_all$perimeter.ratio~av_all$year, var.equal = TRUE)
-t.test(av_all$compactness~av_all$year, var.equal = TRUE)
-t.test(av_all$shape.factor~av_all$year, var.equal = TRUE)
-t.test(av_all$teeth.blade.area~av_all$year, var.equal = TRUE)
-```
-Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).
-
-We next tested each of the variables for *Vitis amurensis*.
-
-```{r}
-am_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "amurensis.all.90")
-am_all$year <- as.factor(am_all$year)
-am_all$petiole.width <- as.numeric(am_all$petiole.width)
-am_all$petiole.width.area <- as.numeric(am_all$petiole.width.area)
-am_all$petiole.area <- as.numeric(am_all$petiole.area)
-am_all$blade.area <- as.numeric(am_all$blade.area)
-am_all$feret.diameter <- as.numeric(am_all$feret.diameter)
-am_all$major.feret <- as.numeric(am_all$major.feret)
-am_all$leaf.area <- as.numeric(am_all$leaf.area)
-am_all$perimeter <- as.numeric(am_all$perimeter)
-am_all$internal.perimeter <- as.numeric(am_all$internal.perimeter)
-am_all$feret.diameter.ratio <- as.numeric(am_all$feret.diameter.ratio)
-am_all$tooth.area.perimeter <- as.numeric(am_all$tooth.area.perimeter)
-am_all$tooth.area.int.perimeter <- as.numeric(am_all$tooth.area.int.perimeter)
-am_all$tooth.area.blade.area <- as.numeric(am_all$tooth.area.blade.area)
-am_all$teeth.perimeter <- as.numeric(am_all$teeth.perimeter)
-am_all$teeth.int.perimeter <- as.numeric(am_all$teeth.int.perimeter)
-am_all$perimeter.area <- as.numeric(am_all$perimeter.area)
-am_all$perimeter.ratio <- as.numeric(am_all$perimeter.ratio)
-am_all$compactness <- as.numeric(am_all$compactness)
-am_all$shape.factor <- as.numeric(am_all$shape.factor)
-am_all$teeth.blade.area <- as.numeric(am_all$teeth.blade.area)
-str(am_all)
-skim(am_all)
-t.test(am_all$total.teeth~am_all$year, var.equal = TRUE)
-t.test(am_all$feret.diameter.ratio~am_all$year, var.equal = TRUE)
-t.test(am_all$tooth.area.perimeter~am_all$year, var.equal = TRUE)
-t.test(am_all$tooth.area.int.perimeter~am_all$year, var.equal = TRUE)
-t.test(am_all$average.tooth.area~am_all$year, var.equal = TRUE)
-t.test(am_all$tooth.area.blade.area~am_all$year, var.equal = TRUE)
-t.test(am_all$teeth.perimeter~am_all$year, var.equal = TRUE)
-t.test(am_all$teeth.int.perimeter~am_all$year, var.equal = TRUE)
-t.test(am_all$perimeter.area~am_all$year, var.equal = TRUE)
-t.test(am_all$perimeter.ratio~am_all$year, var.equal = TRUE)
-t.test(am_all$compactness~am_all$year, var.equal = TRUE)
-t.test(am_all$shape.factor~am_all$year, var.equal = TRUE)
-t.test(am_all$teeth.blade.area~am_all$year, var.equal = TRUE)
-```
-
-Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF).
-
-We finally tested each of the variables for *Vitis riparia*.
-
-```{r}
-r_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "riparia.all.90")
-r_all$year <- as.factor(r_all$year)
-r_all$petiole.width <- as.numeric(r_all$petiole.width)
-r_all$petiole.width.area <- as.numeric(r_all$petiole.width.area)
-r_all$petiole.area <- as.numeric(r_all$petiole.area)
-r_all$internal.perimeter <- as.numeric(r_all$internal.perimeter)
-r_all$tooth.area <- as.numeric(r_all$tooth.area)
-r_all$tooth.area.perimeter <- as.numeric(r_all$tooth.area.perimeter)
-r_all$tooth.area.int.perimeter <- as.numeric(r_all$tooth.area.int.perimeter)
-r_all$average.tooth.area <- as.numeric(r_all$average.tooth.area)
-r_all$tooth.area.blade.area <- as.numeric(r_all$tooth.area.blade.area)
-r_all$teeth.int.perimeter <- as.numeric(r_all$teeth.int.perimeter)
-r_all$perimeter.ratio <- as.numeric(r_all$perimeter.ratio)
-r_all$teeth.blade.area <- as.numeric(r_all$teeth.blade.area)
-str(r_all)
-skim(r_all)
-t.test(r_all$total.teeth~r_all$year, var.equal = TRUE)
-t.test(r_all$feret.diameter.ratio~r_all$year, var.equal = TRUE)
-t.test(r_all$tooth.area.perimeter~r_all$year, var.equal = TRUE)
-t.test(r_all$tooth.area.int.perimeter~r_all$year, var.equal = TRUE)
-t.test(r_all$average.tooth.area~r_all$year, var.equal = TRUE)
-t.test(r_all$tooth.area.blade.area~r_all$year, var.equal = TRUE)
-t.test(r_all$teeth.perimeter~r_all$year, var.equal = TRUE)
-t.test(r_all$teeth.int.perimeter~r_all$year, var.equal = TRUE)
-t.test(r_all$perimeter.area~r_all$year, var.equal = TRUE)
-t.test(r_all$perimeter.ratio~r_all$year, var.equal = TRUE)
-t.test(r_all$compactness~r_all$year, var.equal = TRUE)
-t.test(r_all$shape.factor~r_all$year, var.equal = TRUE)
-t.test(r_all$teeth.blade.area~r_all$year, var.equal = TRUE)
-```
-
-Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).
-
-# Principal Components Analysis (PCA)
-
-```{r}
-library(pcaMethods)
-excel_sheets("R_DiLP_Vitis_all_added.xlsx")
-```
-
-First, we applied a PCA to *Vitis acerifolia*.
-```{r}
-ac_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "acerifolia.all.ln")
-ac_ln_mat <- data.matrix(ac_ln, rownames.force = NA)
-ac_ln_fixed <- ac_ln_mat[,-(1:3)]
-ac_ln_fixed[ac_ln_fixed == 0] <- NA
-#pca_ac_all <- pca(ac_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
-bpca_ac_ll <- pca(ac_ln_fixed, method = "bpca", nPcs = 5)
-summary(bpca_ac_ll)
-loadings(bpca_ac_ll)
-#summary(pca_ac_all)
-#loadings(pca_ac_all)
-print(bpca_ac_ll)
-#print(pca_ac_all)
-#slplot(pca_ac_all)
-#loadings_pca_ac <- loadings(pca_ac_all)
-#loadings_pca_ac_df <- as.data.frame(loadings_pca_ac)
-#str(loadings_pca_ac_df)
-#ggplot(loadings_pca_ac_df, aes(x = PC1, y = PC2, label = rownames(loadings_pca_ac_df))) +
-  #geom_col(x = PC1, y = PC2) +
-  #ggtitle("PCA plot of Vitis acerifolia")
-#attach row names
-```
-
-We then applied a PCA to *Vitis aestivalis*.
-
-```{r}
-av_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "aestivalis.all.ln")
-av_ln_mat <- data.matrix(av_ln, rownames.force = NA)
-av_ln_fixed <- av_ln_mat[,-(1:3)]
-av_ln_fixed[av_ln_fixed == 0] <- NA
-#pca_av_all <- pca(av_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
-bpca_av_ll <- pca(av_ln_fixed, method = "bpca", nPcs = 5)
-summary(bpca_av_ll)
-loadings(bpca_av_ll)
-#summary(pca_av_all)
-#loadings(pca_av_all)
-print(bpca_av_ll)
-#print(pca_av_all)
-#slplot(pca_av_all)
-```
-
-We then applied a PCA to *Vitis amurensis*.
-
-```{r}
-am_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "amurensis.all.ln")
-am_ln_mat <- data.matrix(am_ln, rownames.force = NA)
-am_ln_fixed <- am_ln_mat[,-(1:3)]
-am_ln_fixed[am_ln_fixed == 0] <- NA
-#pca_am_all <- pca(am_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
-bpca_am_ll <- pca(am_ln_fixed, method = "bpca", nPcs = 5)
-summary(bpca_am_ll)
-loadings(bpca_am_ll)
-#summary(pca_am_all)
-#loadings(pca_am_all)
-print(bpca_am_ll)
-#print(pca_am_all)
-#slplot(pca_am_all)
-```
-
-Finally, we applied a PCA to *Vitis riparia*.
-
-```{r}
-r_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "riparia.all.ln")
-r_ln_mat <- data.matrix(r_ln, rownames.force = NA)
-r_ln_fixed <- r_ln_mat[,-(1:3)]
-r_ln_fixed[r_ln_fixed == 0] <- NA
-#pca_r_all <- pca(r_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
-bpca_r_ll <- pca(r_ln_fixed, method = "bpca", nPcs = 5)
-summary(bpca_r_ll)
-loadings(bpca_r_ll)
-#summary(pca_r_all)
-#loadings(pca_r_all)
-print(bpca_r_ll)
-#print(pca_r_all)
-#slplot(pca_r_all)
-```
-
-# Generalized Linear Modeling
+# Linear Modeling
 
@@
 
-We ran a GLM for leaf number and total teeth.
+We ran a LM for leaf number and total teeth.
+```{r}
+acerifolia <-filter(R_DiLP_all_species, species == "acerifolia")
+skim(acerifolia, total.teeth)
+ac_tt_plot <- ggplot(acerifolia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic()
+ac_tt_plot <- ac_tt_plot +labs(x = "Leaf number",
+                               y = "Total teeth")
+ac_tt_plot
+
+```
 
 ```{r}
-skim(ac_all)
-ac_tt_glm <- glm(leaf~total.teeth, data = ac_all, family = gaussian())
-summary(ac_tt_glm)
-plot(ac_tt_glm)
+ac_tt_lm <- lm(total.teeth ~ leaf, data = acerifolia)
+summary(ac_tt_lm)
 ```
 
-We ran a GLM for leaf number and feret diameter ratio.
-
 ```{r}
-ac_fdr_glm <- glm(leaf~feret.diameter.ratio, data = ac_all, family = gaussian())
-summary(ac_fdr_glm)
-plot(ac_fdr_glm)
+ac_tt_seg <- segmented(ac_tt_lm,
+                       seg.Z = ~ leaf,
+                       psi = NA)
+summary(ac_tt_seg)
 ```
 
-We ran a GLM for leaf number and tooth area:perimeter.
+First, we tested *Vitis aestivalis*.
 
+We ran a GLM for leaf number and total teeth.
 ```{r}
-ac_tap_glm <- glm(leaf~tooth.area.perimeter, data = ac_all, family = gaussian())
-summary(ac_tap_glm)
-plot(ac_tap_glm)
+aestivalis <-filter(R_DiLP_all_species, species == "aestivalis")
 ```
 
-We ran a GLM for leaf number and tooth area:internal perimeter.
+First, we tested *Vitis amurensis*.
 
+We ran a GLM for leaf number and total teeth.
 ```{r}
-ac_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = ac_all, family = gaussian())
-summary(ac_taip_glm)
-plot(ac_taip_glm)
+amurensis <-filter(R_DiLP_all_species, species == "amurensis")
 ```
 
-We ran a GLM for leaf number and average tooth area.
+First, we tested *Vitis riparia*.
 
+We ran a GLM for leaf number and total teeth.
 ```{r}
-ac_ata_glm <- glm(leaf~average.tooth.area, data = ac_all, family = gaussian())
-summary(ac_ata_glm)
-plot(ac_ata_glm)
+riparia <-filter(R_DiLP_all_species, species == "riparia")
 ```
-
-We ran a GLM for leaf number and tooth area:blade area..
-
-```{r}
-ac_taba_glm <- glm(leaf~tooth.area.blade.area, data = ac_all, family = gaussian())
-summary(ac_taba_glm)
-plot(ac_taba_glm)
-```
-
-We ran a GLM for leaf number and teeth:perimeter.
-
-```{r}
-ac_tp_glm <- glm(leaf~teeth.perimeter, data = ac_all, family = gaussian())
-summary(ac_tp_glm)
-plot(ac_tp_glm)
-```
-
-We ran a GLM for leaf number and teeth:internal perimeter.
-
-```{r}
-ac_tip_glm <- glm(leaf~teeth.int.perimeter, data = ac_all, family = gaussian())
-summary(ac_tip_glm)
-plot(ac_tip_glm)
-```
-
-We ran a GLM for leaf number and perimeter:area.
-
-```{r}
-ac_pa_glm <- glm(leaf~perimeter.area, data = ac_all, family = gaussian())
-summary(ac_pa_glm)
-plot(ac_pa_glm)
-```
-
-We ran a GLM for leaf number and perimeter ratio.
-
-```{r}
-ac_pr_glm <- glm(leaf~perimeter.ratio, data = ac_all, family = gaussian())
-summary(ac_pr_glm)
-plot(ac_pr_glm)
-```
-
-We ran a GLM for leaf number and compactness.
-
-```{r}
-ac_cp_glm <- glm(leaf~compactness, data = ac_all, family = gaussian())
-summary(ac_cp_glm)
-plot(ac_cp_glm)
-```
-
-We ran a GLM for leaf number and shape factor.
-
-```{r}
-ac_sf_glm <- glm(leaf~shape.factor, data = ac_all, family = gaussian())
-summary(ac_sf_glm)
-plot(ac_sf_glm)
-```
-
-We ran a GLM for leaf number and teeth:blade area.
-
-```{r}
-ac_tba_glm <- glm(leaf~teeth.blade.area, data = ac_all, family = gaussian())
-summary(ac_tba_glm)
-plot(ac_tba_glm)
-```
-
-We ran a GLM for leaf number and multiple variables. 
-```{r}
-ac_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = ac_all, family = gaussian())
-summary(ac_clim_glm)
-plot(ac_clim_glm)
-beep(sound = 3, expr = NULL)
-```

### 2018-10-10 Vitis_GLM.Rmd
acerifolia tt breakpoint SUCCESS

@@
 ```{r}
-acerifolia <-filter(R_DiLP_all_species, species == "acerifolia")
+acerifolia <-filter(Vitis_90_ratios, species == "acerifolia")
 skim(acerifolia, total.teeth)
-ac_tt_plot <- ggplot(acerifolia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic()
+ac_tt_plot <- ggplot(acerifolia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
 ac_tt_plot <- ac_tt_plot +labs(x = "Leaf number",
@@
 ```
+We need the means for each leaf character for both years combined.
+```{r}
+ac_l_means <-  acerifolia %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tt = mean(total.teeth),
+            avg.fdr = mean(feret.diameter.ratio),
+            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
+            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
+            avg.ta = mean(average.tooth.area, na.rm = TRUE),
+            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
+            avg.tp = mean(teeth.perimeter),
+            avg.tip = mean(teeth.int.perimeter),
+            avg.pa = mean(perimeter.area),
+            avg.pr = mean(perimeter.ratio),
+            avg.cp = mean(compactness),
+            avg.sf = mean(shape.factor),
+            avg.tba = mean(teeth.blade.area))
+```
 
 ```{r}
-ac_tt_lm <- lm(total.teeth ~ leaf, data = acerifolia)
+ac_tt_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
+ac_tt_mean_plot <- ac_tt_mean_plot +labs(x = "Leaf number",
+                               y = "Total teeth")
+ac_tt_mean_plot
+```
+
+```{r}
+ac_tt_lm <- lm(avg.tt ~ leaf, data = ac_l_means)
 summary(ac_tt_lm)
@@
                        seg.Z = ~ leaf,
-                       psi = NA)
+                       psi = list(leaf = c(2, 7)))
 summary(ac_tt_seg)
+```
+
+```{r}
+ac_tt_seg$psi
+slope(ac_tt_seg)
+ac_tt_fitted <- fitted(ac_tt_seg)
+ac_tt_model <- data.frame(Leaf = ac_l_means$leaf, Teeth = ac_tt_fitted)
+ggplot(ac_tt_model, aes(x = Leaf, y = Teeth)) + geom_line()
+ac_tt_lines <- ac_tt_seg$psi[, 2]
+ac_tt_mean_break <- ac_tt_plot + geom_vline(xintercept = ac_tt_lines, linetype = "dashed")
+ac_tt_mean_break
 ```

### 2018-10-19 Vitis_GLM.Rmd
VAC PA

@@
 
+# t-tests
+First, we tested whether there were significant differences between the two growing seasons for each of the ratios measured, as well as the total number of teeth.
+To do this, we ran a series of t-tests.
+
+We first tested each of the variables for *Vitis acerifolia*. Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).
+
+We next tested each of the variables for *Vitis aestivalis*.
+Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).
+
+We next tested each of the variables for *Vitis amurensis*.
+Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF).
+
+We finally tested each of the variables for *Vitis riparia*.
+Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).
+
 # Linear Modeling
@@
 
-We ran a LM for leaf number and total teeth.
-```{r}
-acerifolia <-filter(Vitis_90_ratios, species == "acerifolia")
-skim(acerifolia, total.teeth)
-ac_tt_plot <- ggplot(acerifolia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tt_plot <- ac_tt_plot +labs(x = "Leaf number",
-                               y = "Total teeth")
-ac_tt_plot
-
-```
 We need the means for each leaf character for both years combined.
@@
 
+We wanted to test leaf number and total teeth.
+```{r}
+acerifolia <-filter(Vitis_90_ratios, species == "acerifolia")
+skim(acerifolia, total.teeth)
+ac_tt_plot <- ggplot(acerifolia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_tt_plot <- ac_tt_plot +labs(x = "Leaf number",
+                               y = "Total teeth")
+ac_tt_plot
+
+```
+
 ```{r}
@@
 
+We ran a LM for leaf number and total teeth.
 ```{r}
@@
                        seg.Z = ~ leaf,
-                       psi = list(leaf = c(2, 7)))
+                       psi = list(leaf = c(3, 7, 10)))
 summary(ac_tt_seg)
@@
 
-First, we tested *Vitis aestivalis*.
+We wanted to test leaf number and feret diameter ratio.
+```{r}
+skim(acerifolia, total.teeth)
+ac_fdr_plot <- ggplot(acerifolia, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_fdr_plot <- ac_fdr_plot +labs(x = "Leaf number",
+                               y = "Feret Diameter Ratio")
+ac_fdr_plot
+
+```
+
+```{r}
+ac_fdr_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
+ac_fdr_mean_plot <- ac_fdr_mean_plot +labs(x = "Leaf number",
+                               y = "Feret Diameter Ratio")
+ac_fdr_mean_plot
+```
+
+We ran a LM for leaf number and feret diamter ratio.
+```{r}
+ac_fdr_lm <- lm(avg.fdr ~ leaf, data = ac_l_means)
+summary(ac_fdr_lm)
+```
+
+```{r}
+ac_fdr_seg <- segmented(ac_fdr_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_fdr_seg)
+```
+
+```{r}
+ac_fdr_seg$psi
+slope(ac_fdr_seg)
+ac_fdr_fitted <- fitted(ac_fdr_seg)
+ac_fdr_model <- data.frame(Leaf = ac_l_means$leaf, FDR = ac_fdr_fitted)
+ggplot(ac_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
+ac_fdr_lines <- ac_fdr_seg$psi[, 2]
+ac_fdr_mean_break <- ac_fdr_plot + geom_vline(xintercept = ac_fdr_lines, linetype = "dashed")
+ac_fdr_mean_break
+```
+We tested leaf number and tooth area:perimeter.
+```{r}
+skim(acerifolia, tooth.area.perimeter)
+ac_tap_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_tap_plot <- ac_tap_plot +labs(x = "Leaf number",
+                               y = "Tooth area: perimeter")
+ac_tap_plot
+
+```
+
+```{r}
+ac_tap_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
+ac_tap_mean_plot <- ac_tap_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth Area: Perimeter")
+ac_tap_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:perimeter.
+```{r}
+ac_tap_lm <- lm(avg.tap ~ leaf, data = ac_l_means)
+summary(ac_tap_lm)
+```
+
+```{r}
+ac_tap_seg <- segmented(ac_tap_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_tap_seg)
+```
+
+```{r}
+ac_tap_seg$psi
+slope(ac_tap_seg)
+ac_tap_fitted <- fitted(ac_tap_seg)
+ac_tap_model <- data.frame(Leaf = ac_l_means$leaf, TAP = ac_tap_fitted)
+ggplot(ac_tap_model, aes(x = Leaf, y = TAP)) + geom_line()
+ac_tap_lines <- ac_tap_seg$psi[, 2]
+ac_tap_mean_break <- ac_tap_plot + geom_vline(xintercept = ac_tap_lines, linetype = "dashed")
+ac_tap_mean_break
+```
+
+We tested leaf number and tooth area:internal perimeter.
+```{r}
+skim(acerifolia, tooth.area.int.perimeter)
+ac_taip_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_taip_plot <- ac_taip_plot +labs(x = "Leaf number",
+                               y = "Tooth area: internal perimeter")
+ac_taip_plot
+
+```
+
+```{r}
+ac_taip_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
+ac_taip_mean_plot <- ac_taip_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth Area: Internal Perimeter")
+ac_taip_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:internal perimeter.
+```{r}
+ac_taip_lm <- lm(avg.taip ~ leaf, data = ac_l_means)
+summary(ac_taip_lm)
+```
+
+```{r}
+ac_taip_seg <- segmented(ac_taip_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_taip_seg)
+```
+
+```{r}
+ac_taip_seg$psi
+slope(ac_taip_seg)
+ac_taip_fitted <- fitted(ac_taip_seg)
+ac_taip_model <- data.frame(Leaf = ac_l_means$leaf, TAIP = ac_taip_fitted)
+ggplot(ac_taip_model, aes(x = Leaf, y = TAIP)) + geom_line()
+ac_taip_lines <- ac_taip_seg$psi[, 2]
+ac_taip_mean_break <- ac_taip_plot + geom_vline(xintercept = ac_taip_lines, linetype = "dashed")
+ac_taip_mean_break
+```
+
+We tested leaf number and average tooth area.
+```{r}
+skim(acerifolia, average.tooth.area)
+ac_ta_plot <- ggplot(acerifolia, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_ta_plot <- ac_ta_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+ac_ta_plot
+
+```
+
+```{r}
+ac_ta_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.ta)) + geom_line() + theme_classic()
+ac_ta_mean_plot <- ac_ta_mean_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+ac_ta_mean_plot
+```
+
+We ran a LM for leaf number and average tooth area.
+```{r}
+ac_ta_lm <- lm(avg.ta ~ leaf, data = ac_l_means)
+summary(ac_ta_lm)
+```
+
+```{r}
+ac_ta_seg <- segmented(ac_ta_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_ta_seg)
+```
+
+```{r}
+ac_ta_seg$psi
+slope(ac_ta_seg)
+ac_ta_fitted <- fitted(ac_ta_seg)
+ac_ta_model <- data.frame(Leaf = ac_l_means$leaf, TA = ac_ta_fitted)
+ggplot(ac_ta_model, aes(x = Leaf, y = TA)) + geom_line()
+ac_ta_lines <- ac_ta_seg$psi[, 2]
+ac_ta_mean_break <- ac_ta_plot + geom_vline(xintercept = ac_ta_lines, linetype = "dashed")
+ac_ta_mean_break
+```
+
+We tested leaf number and teeth:perimeter.
+```{r}
+skim(acerifolia, teeth.perimeter)
+ac_tp_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_tp_plot <- ac_tp_plot +labs(x = "Leaf number",
+                               y = "Teeth:perimeter")
+ac_tp_plot
+
+```
+
+```{r}
+ac_tp_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
+ac_tp_mean_plot <- ac_tp_mean_plot +labs(x = "Leaf number",
+                               y = "Teeth:perimeter")
+ac_tp_mean_plot
+```
+
+We ran a LM for leaf number and teeth:perimeter.
+```{r}
+ac_tp_lm <- lm(avg.tp ~ leaf, data = ac_l_means)
+summary(ac_tp_lm)
+```
+
+```{r}
+ac_tp_seg <- segmented(ac_tp_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_tp_seg)
+```
+
+```{r}
+ac_tp_seg$psi
+slope(ac_tp_seg)
+ac_tp_fitted <- fitted(ac_tp_seg)
+ac_tp_model <- data.frame(Leaf = ac_l_means$leaf, TP = ac_tp_fitted)
+ggplot(ac_tp_model, aes(x = Leaf, y = TP)) + geom_line()
+ac_tp_lines <- ac_tp_seg$psi[, 2]
+ac_tp_mean_break <- ac_tp_plot + geom_vline(xintercept = ac_tp_lines, linetype = "dashed")
+ac_tp_mean_break
+```
+
+We tested leaf number and teeth:internal perimeter.
+```{r}
+skim(acerifolia, teeth.int.perimeter)
+ac_tip_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_tip_plot <- ac_tip_plot +labs(x = "Leaf number",
+                               y = "Teeth:internal perimeter")
+ac_tip_plot
+
+```
+
+```{r}
+ac_tip_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
+ac_tip_mean_plot <- ac_tip_mean_plot +labs(x = "Leaf number",
+                               y = "Teeth:internal perimeter")
+ac_tip_mean_plot
+```
+
+We ran a LM for leaf number and teeth:internal perimeter.
+```{r}
+ac_tip_lm <- lm(avg.tip ~ leaf, data = ac_l_means)
+summary(ac_tip_lm)
+```
+
+```{r}
+ac_tip_seg <- segmented(ac_tip_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_tip_seg)
+```
+
+```{r}
+ac_tip_seg$psi
+slope(ac_tip_seg)
+ac_tip_fitted <- fitted(ac_tip_seg)
+ac_tip_model <- data.frame(Leaf = ac_l_means$leaf, TIP = ac_tip_fitted)
+ggplot(ac_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
+ac_tip_lines <- ac_tip_seg$psi[, 2]
+ac_tip_mean_break <- ac_tip_plot + geom_vline(xintercept = ac_tip_lines, linetype = "dashed")
+ac_tip_mean_break
+```
+
+We tested leaf number and perimeter:area.
+```{r}
+skim(acerifolia, perimeter.area)
+ac_pa_plot <- ggplot(acerifolia, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_pa_plot <- ac_pa_plot +labs(x = "Leaf number",
+                               y = "Perimeter:Area")
+ac_pa_plot
+
+```
+
+```{r}
+ac_pa_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
+ac_pa_mean_plot <- ac_pa_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter:Area")
+ac_pa_mean_plot
+```
+
+We ran a LM for leaf number and perimeter:area.
+```{r}
+ac_pa_lm <- lm(avg.pa ~ leaf, data = ac_l_means)
+summary(ac_pa_lm)
+```
+
+```{r}
+ac_pa_seg <- segmented(ac_pa_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_pa_seg)
+```
+
+```{r}
+ac_pa_seg$psi
+slope(ac_pa_seg)
+ac_pa_fitted <- fitted(ac_pa_seg)
+ac_pa_model <- data.frame(Leaf = ac_l_means$leaf, PA = ac_pa_fitted)
+ggplot(ac_pa_model, aes(x = Leaf, y = PA)) + geom_line()
+ac_pa_lines <- ac_pa_seg$psi[, 2]
+ac_pa_mean_break <- ac_pa_plot + geom_vline(xintercept = ac_pa_lines, linetype = "dashed")
+ac_pa_mean_break
+```
+
+We then tested *Vitis aestivalis*.
 
@@
 ```{r}
-aestivalis <-filter(R_DiLP_all_species, species == "aestivalis")
+aestivalis <-filter(Vitis_90_ratios, species == "aestivalis")
+```
+
+We need the means for each leaf character for both years combined.
+```{r}
+av_l_means <-  aestivalis %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tt = mean(total.teeth),
+            avg.fdr = mean(feret.diameter.ratio, na.rm = TRUE),
+            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
+            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
+            avg.ta = mean(average.tooth.area, na.rm = TRUE),
+            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
+            avg.tp = mean(teeth.perimeter),
+            avg.tip = mean(teeth.int.perimeter),
+            avg.pa = mean(perimeter.area),
+            avg.pr = mean(perimeter.ratio),
+            avg.cp = mean(compactness),
+            avg.sf = mean(shape.factor),
+            avg.tba = mean(teeth.blade.area))
 ```

### 2018-10-19 Vitis_GLM.Rmd
Vitis acerifolia except TABA

@@
 
+We tested leaf number and perimeter ratio.
+```{r}
+skim(acerifolia, perimeter.ratio)
+ac_pr_plot <- ggplot(acerifolia, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_pr_plot <- ac_pr_plot +labs(x = "Leaf number",
+                               y = "Perimeter Ratio")
+ac_pr_plot
+
+```
+
+```{r}
+ac_pr_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
+ac_pr_mean_plot <- ac_pr_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter Ratio")
+ac_pr_mean_plot
+```
+
+We ran a LM for leaf number and perimeter ratio.
+```{r}
+ac_pr_lm <- lm(avg.pr ~ leaf, data = ac_l_means)
+summary(ac_pr_lm)
+```
+
+```{r}
+ac_pr_seg <- segmented(ac_pr_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_pr_seg)
+```
+
+```{r}
+ac_pr_seg$psi
+slope(ac_pr_seg)
+ac_pr_fitted <- fitted(ac_pr_seg)
+ac_pr_model <- data.frame(Leaf = ac_l_means$leaf, PR = ac_pr_fitted)
+ggplot(ac_pr_model, aes(x = Leaf, y = PR)) + geom_line()
+ac_pr_lines <- ac_pr_seg$psi[, 2]
+ac_pr_mean_break <- ac_pr_plot + geom_vline(xintercept = ac_pr_lines, linetype = "dashed")
+ac_pr_mean_break
+```
+
+We tested leaf number and compactness.
+```{r}
+skim(acerifolia, compactness)
+ac_cp_plot <- ggplot(acerifolia, aes(x = leaf, y = compactness)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_cp_plot <- ac_cp_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+ac_cp_plot
+
+```
+
+```{r}
+ac_cp_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.cp)) + geom_line() + theme_classic()
+ac_cp_mean_plot <- ac_cp_mean_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+ac_cp_mean_plot
+```
+
+We ran a LM for leaf number and compactness.
+```{r}
+ac_cp_lm <- lm(avg.cp ~ leaf, data = ac_l_means)
+summary(ac_cp_lm)
+```
+
+```{r}
+ac_cp_seg <- segmented(ac_cp_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_cp_seg)
+```
+
+```{r}
+ac_cp_seg$psi
+slope(ac_cp_seg)
+ac_cp_fitted <- fitted(ac_cp_seg)
+ac_cp_model <- data.frame(Leaf = ac_l_means$leaf, CP = ac_cp_fitted)
+ggplot(ac_cp_model, aes(x = Leaf, y = CP)) + geom_line()
+ac_cp_lines <- ac_cp_seg$psi[, 2]
+ac_cp_mean_break <- ac_cp_plot + geom_vline(xintercept = ac_cp_lines, linetype = "dashed")
+ac_cp_mean_break
+```
+
+We tested leaf number and shape factor.
+```{r}
+skim(acerifolia, shape.factor)
+ac_sf_plot <- ggplot(acerifolia, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_sf_plot <- ac_sf_plot +labs(x = "Leaf number",
+                               y = "Shape Factor")
+ac_sf_plot
+
+```
+
+```{r}
+ac_sf_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
+ac_sf_mean_plot <- ac_sf_mean_plot +labs(x = "Leaf number",
+                               y = "Shape Factor")
+ac_sf_mean_plot
+```
+
+We ran a LM for leaf number and shape factor.
+```{r}
+ac_sf_lm <- lm(avg.sf ~ leaf, data = ac_l_means)
+summary(ac_sf_lm)
+```
+
+```{r}
+ac_sf_seg <- segmented(ac_sf_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_sf_seg)
+```
+
+```{r}
+ac_sf_seg$psi
+slope(ac_sf_seg)
+ac_sf_fitted <- fitted(ac_sf_seg)
+ac_sf_model <- data.frame(Leaf = ac_l_means$leaf, SF = ac_sf_fitted)
+ggplot(ac_sf_model, aes(x = Leaf, y = SF)) + geom_line()
+ac_sf_lines <- ac_sf_seg$psi[, 2]
+ac_sf_mean_break <- ac_sf_plot + geom_vline(xintercept = ac_sf_lines, linetype = "dashed")
+ac_sf_mean_break
+```
+
+We tested leaf number and teeth:blade area.
+```{r}
+skim(acerifolia, teeth.blade.area)
+ac_tba_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_tba_plot <- ac_tba_plot +labs(x = "Leaf number",
+                               y = "Teeth:Blade Area")
+ac_tba_plot
+
+```
+
+```{r}
+ac_tba_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
+ac_tba_mean_plot <- ac_tba_mean_plot +labs(x = "Leaf number",
+                               y = "Teeth:Blade Area")
+ac_tba_mean_plot
+```
+
+We ran a LM for leaf number and teeth:blade area.
+```{r}
+ac_tba_lm <- lm(avg.tba ~ leaf, data = ac_l_means)
+summary(ac_tba_lm)
+```
+
+```{r}
+ac_tba_seg <- segmented(ac_tba_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_tba_seg)
+```
+
+```{r}
+ac_tba_seg$psi
+slope(ac_tba_seg)
+ac_tba_fitted <- fitted(ac_tba_seg)
+ac_tba_model <- data.frame(Leaf = ac_l_means$leaf, TBA = ac_tba_fitted)
+ggplot(ac_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
+ac_tba_lines <- ac_tba_seg$psi[, 2]
+ac_tba_mean_break <- ac_tba_plot + geom_vline(xintercept = ac_tba_lines, linetype = "dashed")
+ac_tba_mean_break
+```
+
 We then tested *Vitis aestivalis*.

### 2018-10-23 Vitis_GLM.Rmd
V. acerifolia leaf lm

@@
 library(ggplot2)
+library(gganimate)
 library(segmented)
+Vitis_90_ratios <- read_excel("R_DiLP_all_species.xlsx", 
+     sheet = "all.90", col_types = c("text", 
+         "numeric", "numeric", "numeric", 
+         "skip", "skip", "skip", "skip", 
+         "skip", "skip", "skip", "skip", 
+         "skip", "skip", "skip", "numeric", 
+         "skip", "numeric", "numeric", "numeric", 
+         "numeric", "numeric", "numeric", 
+         "numeric", "numeric", "numeric", 
+         "numeric", "numeric", "numeric"))
 ```
@@
 ```{r}
+acerifolia <-filter(Vitis_90_ratios, species == "acerifolia")
 ac_l_means <-  acerifolia %>% 
@@
 
-We wanted to test leaf number and total teeth.
+However we need to split tooth area:blade area for each year.
 ```{r}
-acerifolia <-filter(Vitis_90_ratios, species == "acerifolia")
+ac_13 <- filter(acerifolia, year == 2013)
+ac_15 <- filter(acerifolia, year == 2015)
+```
+
+```{r}
+ac_taba13_mean <-  acerifolia %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
+
+ac_taba15_mean <- acerifolia %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
+```
+
+We tested leaf number and total teeth.
+```{r}
 skim(acerifolia, total.teeth)
@@
 ac_tt_mean_break
+ggsave("ac_tt_break.png")
 ```
@@
 ac_fdr_mean_break
+ggsave("ac_fdr_break.png")
 ```
@@
 ac_tap_mean_break
+ggsave("ac_tap_break.png")
 ```
@@
 ac_taip_mean_break
+ggsave("ac_taip_break.png")
 ```
@@
 ac_ta_mean_break
+ggsave("ac_ta_break.png")
 ```
@@
 ac_tp_mean_break
+ggsave("ac_tp_break.png")
 ```
@@
 ac_tip_mean_break
+ggsave("ac_tip_break.png")
 ```
@@
 ac_pa_mean_break
+ggsave("ac_pa_break.png")
 ```
@@
 ac_pr_mean_break
+ggsave("ac_pr_break.png")
 ```
@@
 ac_cp_mean_break
+ggsave("ac_cp_break.png")
 ```
@@
 ac_sf_mean_break
+ggsave("ac_sf_break.png")
 ```
@@
 ac_tba_mean_break
+ggsave("ac_tba_break.png")
+```
+
+We tested leaf number and tooth area:blade area for 2013.
+```{r}
+ac_taba13_plot <- ggplot(ac_13, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_taba13_plot <- ac_taba13_plot +labs(x = "Leaf number",
+                               y = "Tooth Area:Blade Area")
+ac_taba13_plot
+
+```
+
+```{r}
+ac_taba13_mean_plot <- ggplot(ac_taba13_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
+ac_taba13_mean_plot <- ac_taba13_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth Area:Blade Area")
+ac_taba13_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:blade area in 2013.
+```{r}
+ac_taba13_lm <- lm(avg.taba ~ leaf, data = ac_taba13_mean)
+summary(ac_taba13_lm)
+```
+
+```{r}
+ac_taba13_seg <- segmented(ac_taba13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_taba13_seg)
+```
+
+```{r}
+ac_taba13_seg$psi
+slope(ac_taba13_seg)
+ac_taba13_fitted <- fitted(ac_taba13_seg)
+ac_taba13_model <- data.frame(Leaf = ac_taba13_mean$leaf, TABA = ac_taba13_fitted)
+ggplot(ac_taba13_model, aes(x = Leaf, y = TABA)) + geom_line()
+ac_taba13_lines <- ac_taba13_seg$psi[, 2]
+ac_taba13_mean_break <- ac_taba13_plot + geom_vline(xintercept = ac_taba13_lines, linetype = "dashed")
+ac_taba13_mean_break
+ggsave("ac_taba13_break.png")
+```
+
+We tested leaf number and tooth area:blade area for 2015.
+```{r}
+ac_taba15_plot <- ggplot(ac_15, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_taba15_plot <- ac_taba15_plot +labs(x = "Leaf number",
+                               y = "Tooth Area:Blade Area")
+ac_taba15_plot
+
+```
+
+```{r}
+ac_taba15_mean_plot <- ggplot(ac_taba15_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
+ac_taba15_mean_plot <- ac_taba15_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth Area:Blade Area")
+ac_taba15_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:blade area in 2015.
+```{r}
+ac_taba15_lm <- lm(avg.taba ~ leaf, data = ac_taba15_mean)
+summary(ac_taba15_lm)
+```
+
+```{r}
+ac_taba15_seg <- segmented(ac_taba15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(ac_taba15_seg)
+```
+
+```{r}
+ac_taba15_seg$psi
+slope(ac_taba15_seg)
+ac_taba15_fitted <- fitted(ac_taba15_seg)
+ac_taba15_model <- data.frame(Leaf = ac_taba15_mean$leaf, TABA = ac_taba15_fitted)
+ggplot(ac_taba15_model, aes(x = Leaf, y = TABA)) + geom_line()
+ac_taba15_lines <- ac_taba15_seg$psi[, 2]
+ac_taba15_mean_break <- ac_taba15_plot + geom_vline(xintercept = ac_taba15_lines, linetype = "dashed")
+ac_taba15_mean_break
+ggsave("ac_taba15_break.png")
 ```

### 2018-10-23 Vitis_GLM.Rmd
V. aestivalis all but sig diff

@@
 ```
+
 We tested leaf number and tooth area:perimeter.
@@
 
-We ran a GLM for leaf number and total teeth.
 ```{r}
@@
             avg.tba = mean(teeth.blade.area))
+skim(av_l_means)
+```
+
+Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF), so we need to split these variables for each year.
+```{r}
+av_13 <- filter(aestivalis, year == 2013)
+av_15 <- filter(aestivalis, year == 2015)
+```
+
+Total teeth
+```{r}
+av_tt13_mean <-  aestivalis %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
+
+ac_tt15_mean <- aestivalis %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
+```
+
+Tooth area: blade area
+```{r}
+av_taba13_mean <-  aestivalis %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
+
+ac_taba15_mean <- aestivalis %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
+```
+
+Compactness
+```{r}
+av_cp13_mean <-  aestivalis %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.cp = mean(compactness, na.rm = TRUE))
+
+ac_cp15_mean <- aestivalis %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.cp = mean(compactness, na.rm = TRUE))
+```
+
+Shape factor
+```{r}
+av_sf13_mean <-  aestivalis %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
+
+ac_sf15_mean <- aestivalis %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
+```
+
+We wanted to test leaf number and feret diameter ratio.
+```{r}
+av_fdr_plot <- ggplot(aestivalis, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_fdr_plot <- av_fdr_plot +labs(x = "Leaf number",
+                               y = "Feret Diameter Ratio")
+av_fdr_plot
+```
+
+```{r}
+av_fdr_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
+av_fdr_mean_plot <- av_fdr_mean_plot +labs(x = "Leaf number",
+                               y = "Feret Diameter Ratio")
+av_fdr_mean_plot
+```
+
+We ran a LM for leaf number and feret diamter ratio.
+```{r}
+av_fdr_lm <- lm(avg.fdr ~ leaf, data = av_l_means)
+summary(av_fdr_lm)
+```
+
+```{r}
+av_fdr_seg <- segmented(av_fdr_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_fdr_seg)
+```
+
+```{r}
+av_fdr_seg$psi
+slope(av_fdr_seg)
+av_fdr_fitted <- fitted(av_fdr_seg)
+av_fdr_model <- data.frame(Leaf = av_l_means$leaf, FDR = av_fdr_fitted)
+ggplot(av_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
+av_fdr_lines <- av_fdr_seg$psi[, 2]
+av_fdr_mean_break <- av_fdr_plot + geom_vline(xintercept = av_fdr_lines, linetype = "dashed")
+av_fdr_mean_break
+ggsave("av_fdr_break.png")
+```
+
+We tested leaf number and tooth area:perimeter.
+```{r}
+av_tap_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_tap_plot <- av_tap_plot +labs(x = "Leaf number",
+                               y = "Tooth area: perimeter")
+av_tap_plot
+```
+
+```{r}
+av_tap_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
+av_tap_mean_plot <- av_tap_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth Area: Perimeter")
+av_tap_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:perimeter.
+```{r}
+av_tap_lm <- lm(avg.tap ~ leaf, data = av_l_means)
+summary(av_tap_lm)
+```
+
+```{r}
+av_tap_seg <- segmented(av_tap_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(4, 7, 10)))
+summary(av_tap_seg)
+```
+
+```{r}
+av_tap_seg$psi
+slope(av_tap_seg)
+av_tap_fitted <- fitted(av_tap_seg)
+av_tap_model <- data.frame(Leaf = av_l_means$leaf, TAP = av_tap_fitted)
+ggplot(av_tap_model, aes(x = Leaf, y = TAP)) + geom_line()
+av_tap_lines <- av_tap_seg$psi[, 2]
+av_tap_mean_break <- av_tap_plot + geom_vline(xintercept = av_tap_lines, linetype = "dashed")
+av_tap_mean_break
+ggsave("av_tap_break.png")
+```
+
+We tested leaf number and tooth area:internal perimeter.
+```{r}
+av_taip_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_taip_plot <- av_taip_plot +labs(x = "Leaf number",
+                               y = "Tooth area: internal perimeter")
+av_taip_plot
+```
+
+```{r}
+av_taip_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
+av_taip_mean_plot <- av_taip_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth Area: Internal Perimeter")
+av_taip_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:internal perimeter.
+```{r}
+av_taip_lm <- lm(avg.taip ~ leaf, data = av_l_means)
+summary(av_taip_lm)
+```
+
+```{r}
+av_taip_seg <- segmented(av_taip_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_taip_seg)
+```
+
+```{r}
+av_taip_seg$psi
+slope(av_taip_seg)
+av_taip_fitted <- fitted(av_taip_seg)
+av_taip_model <- data.frame(Leaf = av_l_means$leaf, TAIP = av_taip_fitted)
+ggplot(av_taip_model, aes(x = Leaf, y = TAIP)) + geom_line()
+av_taip_lines <- av_taip_seg$psi[, 2]
+av_taip_mean_break <- av_taip_plot + geom_vline(xintercept = av_taip_lines, linetype = "dashed")
+av_taip_mean_break
+ggsave("av_taip_break.png")
+```
+
+We tested leaf number and average tooth area.
+```{r}
+av_ta_plot <- ggplot(aestivalis, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_ta_plot <- av_ta_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+av_ta_plot
+
+```
+
+```{r}
+av_ta_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.ta)) + geom_line() + theme_classic()
+av_ta_mean_plot <- av_ta_mean_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+av_ta_mean_plot
+```
+
+We ran a LM for leaf number and average tooth area.
+```{r}
+av_ta_lm <- lm(avg.ta ~ leaf, data = av_l_means)
+summary(av_ta_lm)
+```
+
+```{r}
+av_ta_seg <- segmented(av_ta_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_ta_seg)
+```
+
+```{r}
+av_ta_seg$psi
+slope(av_ta_seg)
+av_ta_fitted <- fitted(av_ta_seg)
+av_ta_model <- data.frame(Leaf = av_l_means$leaf, TA = av_ta_fitted)
+ggplot(av_ta_model, aes(x = Leaf, y = TA)) + geom_line()
+av_ta_lines <- av_ta_seg$psi[, 2]
+av_ta_mean_break <- av_ta_plot + geom_vline(xintercept = av_ta_lines, linetype = "dashed")
+av_ta_mean_break
+ggsave("av_ta_break.png")
+```
+
+We tested leaf number and teeth:perimeter.
+```{r}
+av_tp_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_tp_plot <- av_tp_plot +labs(x = "Leaf number",
+                               y = "Teeth:perimeter")
+av_tp_plot
+
+```
+
+```{r}
+av_tp_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
+av_tp_mean_plot <- av_tp_mean_plot +labs(x = "Leaf number",
+                               y = "Teeth:perimeter")
+av_tp_mean_plot
+```
+
+We ran a LM for leaf number and teeth:perimeter.
+```{r}
+av_tp_lm <- lm(avg.tp ~ leaf, data = av_l_means)
+summary(av_tp_lm)
+```
+
+```{r}
+av_tp_seg <- segmented(av_tp_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_tp_seg)
+```
+
+```{r}
+av_tp_seg$psi
+slope(av_tp_seg)
+av_tp_fitted <- fitted(av_tp_seg)
+av_tp_model <- data.frame(Leaf = av_l_means$leaf, TP = av_tp_fitted)
+ggplot(av_tp_model, aes(x = Leaf, y = TP)) + geom_line()
+av_tp_lines <- av_tp_seg$psi[, 2]
+av_tp_mean_break <- av_tp_plot + geom_vline(xintercept = av_tp_lines, linetype = "dashed")
+av_tp_mean_break
+ggsave("av_tp_break.png")
+```
+
+We tested leaf number and teeth:internal perimeter.
+```{r}
+av_tip_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_tip_plot <- av_tip_plot +labs(x = "Leaf number",
+                               y = "Teeth:internal perimeter")
+av_tip_plot
+
+```
+
+```{r}
+av_tip_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
+av_tip_mean_plot <- av_tip_mean_plot +labs(x = "Leaf number",
+                               y = "Teeth:internal perimeter")
+av_tip_mean_plot
+```
+
+We ran a LM for leaf number and teeth:internal perimeter.
+```{r}
+av_tip_lm <- lm(avg.tip ~ leaf, data = av_l_means)
+summary(av_tip_lm)
+```
+
+```{r}
+av_tip_seg <- segmented(av_tip_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_tip_seg)
+```
+
+```{r}
+av_tip_seg$psi
+slope(av_tip_seg)
+av_tip_fitted <- fitted(av_tip_seg)
+av_tip_model <- data.frame(Leaf = av_l_means$leaf, TIP = av_tip_fitted)
+ggplot(av_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
+av_tip_lines <- av_tip_seg$psi[, 2]
+av_tip_mean_break <- av_tip_plot + geom_vline(xintercept = av_tip_lines, linetype = "dashed")
+av_tip_mean_break
+ggsave("av_tip_break.png")
+```
+
+We tested leaf number and perimeter:area.
+```{r}
+av_pa_plot <- ggplot(aestivalis, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_pa_plot <- av_pa_plot +labs(x = "Leaf number",
+                               y = "Perimeter:Area")
+av_pa_plot
+
+```
+
+```{r}
+av_pa_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
+av_pa_mean_plot <- av_pa_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter:Area")
+av_pa_mean_plot
+```
+
+We ran a LM for leaf number and perimeter:area.
+```{r}
+av_pa_lm <- lm(avg.pa ~ leaf, data = av_l_means)
+summary(av_pa_lm)
+```
+
+```{r}
+av_pa_seg <- segmented(av_pa_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(2, 7, 10)))
+summary(av_pa_seg)
+```
+
+```{r}
+ac_pa_seg$psi
+slope(ac_pa_seg)
+ac_pa_fitted <- fitted(ac_pa_seg)
+ac_pa_model <- data.frame(Leaf = ac_l_means$leaf, PA = ac_pa_fitted)
+ggplot(ac_pa_model, aes(x = Leaf, y = PA)) + geom_line()
+ac_pa_lines <- ac_pa_seg$psi[, 2]
+ac_pa_mean_break <- ac_pa_plot + geom_vline(xintercept = ac_pa_lines, linetype = "dashed")
+ac_pa_mean_break
+ggsave("ac_pa_break.png")
+```
+
+We tested leaf number and perimeter ratio.
+```{r}
+av_pr_plot <- ggplot(aestivalis, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_pr_plot <- av_pr_plot +labs(x = "Leaf number",
+                               y = "Perimeter Ratio")
+av_pr_plot
+```
+
+```{r}
+av_pr_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
+av_pr_mean_plot <- av_pr_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter Ratio")
+av_pr_mean_plot
+```
+
+We ran a LM for leaf number and perimeter ratio.
+```{r}
+av_pr_lm <- lm(avg.pr ~ leaf, data = av_l_means)
+summary(av_pr_lm)
+```
+
+```{r}
+av_pr_seg <- segmented(av_pr_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_pr_seg)
+```
+
+```{r}
+av_pr_seg$psi
+slope(av_pr_seg)
+av_pr_fitted <- fitted(av_pr_seg)
+av_pr_model <- data.frame(Leaf = av_l_means$leaf, PR = av_pr_fitted)
+ggplot(av_pr_model, aes(x = Leaf, y = PR)) + geom_line()
+av_pr_lines <- av_pr_seg$psi[, 2]
+av_pr_mean_break <- av_pr_plot + geom_vline(xintercept = av_pr_lines, linetype = "dashed")
+av_pr_mean_break
+ggsave("av_pr_break.png")
+```
+
+We tested leaf number and teeth:blade area.
+```{r}
+av_tba_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_tba_plot <- av_tba_plot +labs(x = "Leaf number",
+                               y = "Teeth:Blade Area")
+av_tba_plot
+
+```
+
+```{r}
+av_tba_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
+av_tba_mean_plot <- av_tba_mean_plot +labs(x = "Leaf number",
+                               y = "Teeth:Blade Area")
+av_tba_mean_plot
+```
+
+We ran a LM for leaf number and teeth:blade area.
+```{r}
+av_tba_lm <- lm(avg.tba ~ leaf, data = av_l_means)
+summary(av_tba_lm)
+```
+
+```{r}
+av_tba_seg <- segmented(av_tba_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(2, 4, 6)))
+summary(av_tba_seg)
+```
+
+```{r}
+av_tba_seg$psi
+slope(av_tba_seg)
+av_tba_fitted <- fitted(av_tba_seg)
+av_tba_model <- data.frame(Leaf = av_l_means$leaf, TBA = av_tba_fitted)
+ggplot(av_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
+av_tba_lines <- av_tba_seg$psi[, 2]
+av_tba_mean_break <- av_tba_plot + geom_vline(xintercept = av_tba_lines, linetype = "dashed")
+av_tba_mean_break
+ggsave("av_tba_break.png")
 ```

### 2018-10-24 Vitis_GLM.Rmd
V. aestivalis lm

@@
 
-ac_tt15_mean <- aestivalis %>% 
+av_tt15_mean <- aestivalis %>% 
   filter(year == 2015) %>% 
@@
 
-ac_taba15_mean <- aestivalis %>% 
+av_taba15_mean <- aestivalis %>% 
   filter(year == 2015) %>% 
@@
 
-ac_cp15_mean <- aestivalis %>% 
+av_cp15_mean <- aestivalis %>% 
   filter(year == 2015) %>% 
@@
 
-ac_sf15_mean <- aestivalis %>% 
+av_sf15_mean <- aestivalis %>% 
   filter(year == 2015) %>% 
@@
 
-First, we tested *Vitis amurensis*.
+We tested leaf number and tooth area:blade area for 2013.
+```{r}
+av_taba13_plot <- ggplot(av_13, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_taba13_plot <- av_taba13_plot +labs(x = "Leaf number",
+                               y = "Tooth Area:Blade Area")
+av_taba13_plot
+
+```
+
+```{r}
+av_taba13_mean_plot <- ggplot(av_taba13_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
+av_taba13_mean_plot <- av_taba13_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth Area:Blade Area")
+av_taba13_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:blade area in 2013.
+```{r}
+av_taba13_lm <- lm(avg.taba ~ leaf, data = av_taba13_mean)
+summary(av_taba13_lm)
+```
+
+```{r}
+av_taba13_seg <- segmented(av_taba13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_taba13_seg)
+```
+
+```{r}
+av_taba13_seg$psi
+slope(av_taba13_seg)
+av_taba13_fitted <- fitted(av_taba13_seg)
+av_taba13_model <- data.frame(Leaf = av_taba13_mean$leaf, TABA = av_taba13_fitted)
+ggplot(av_taba13_model, aes(x = Leaf, y = TABA)) + geom_line()
+av_taba13_lines <- av_taba13_seg$psi[, 2]
+av_taba13_mean_break <- av_taba13_plot + geom_vline(xintercept = av_taba13_lines, linetype = "dashed")
+av_taba13_mean_break
+ggsave("av_taba13_break.png")
+```
+
+We tested leaf number and tooth area:blade area for 2015.
+```{r}
+av_taba15_plot <- ggplot(av_15, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_taba15_plot <- av_taba15_plot +labs(x = "Leaf number",
+                               y = "Tooth Area:Blade Area")
+av_taba15_plot
+
+```
+
+```{r}
+av_taba15_mean_plot <- ggplot(av_taba15_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
+av_taba15_mean_plot <- av_taba15_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth Area:Blade Area")
+av_taba15_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:blade area in 2015.
+```{r}
+av_taba15_lm <- lm(avg.taba ~ leaf, data = av_taba15_mean)
+summary(av_taba15_lm)
+```
+
+```{r}
+av_taba15_seg <- segmented(av_taba15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(2, 7, 10)))
+summary(av_taba15_seg)
+```
+
+```{r}
+av_taba15_seg$psi
+slope(av_taba15_seg)
+av_taba15_fitted <- fitted(av_taba15_seg)
+av_taba15_model <- data.frame(Leaf = av_taba15_mean$leaf, TABA = av_taba15_fitted)
+ggplot(av_taba15_model, aes(x = Leaf, y = TABA)) + geom_line()
+av_taba15_lines <- av_taba15_seg$psi[, 2]
+av_taba15_mean_break <- av_taba15_plot + geom_vline(xintercept = av_taba15_lines, linetype = "dashed")
+av_taba15_mean_break
+ggsave("av_taba15_break.png")
+```
+
+We tested leaf number and tooth teeth for 2013.
+```{r}
+av_tt13_plot <- ggplot(av_13, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_tt13_plot <- av_tt13_plot +labs(x = "Leaf number",
+                               y = "Total Teeth")
+av_tt13_plot
+
+```
+
+```{r}
+av_tt13_mean_plot <- ggplot(av_tt13_mean, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
+av_tt13_mean_plot <- av_tt13_mean_plot +labs(x = "Leaf number",
+                               y = "Total Teeth")
+av_tt13_mean_plot
+```
+
+We ran a LM for leaf number and total teeth in 2013.
+```{r}
+av_tt13_lm <- lm(avg.tt ~ leaf, data = av_tt13_mean)
+summary(av_tt13_lm)
+```
+
+```{r}
+av_tt13_seg <- segmented(av_tt13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_tt13_seg)
+```
+
+```{r}
+av_tt13_seg$psi
+slope(av_tt13_seg)
+av_tt13_fitted <- fitted(av_tt13_seg)
+av_tt13_model <- data.frame(Leaf = av_tt13_mean$leaf, TT = av_tt13_fitted)
+ggplot(av_tt13_model, aes(x = Leaf, y = TT)) + geom_line()
+av_tt13_lines <- av_tt13_seg$psi[, 2]
+av_tt13_mean_break <- av_tt13_plot + geom_vline(xintercept = av_tt13_lines, linetype = "dashed")
+av_tt13_mean_break
+ggsave("av_tt13_break.png")
+```
+
+We tested leaf number and total teeth for 2015.
+```{r}
+av_tt15_plot <- ggplot(av_15, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_tt15_plot <- av_tt15_plot +labs(x = "Leaf number",
+                               y = "Total Teeth")
+av_tt15_plot
+
+```
+
+```{r}
+av_tt15_mean_plot <- ggplot(av_tt15_mean, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
+av_tt15_mean_plot <- av_tt15_mean_plot +labs(x = "Leaf number",
+                               y = "Total teeth")
+av_tt15_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:blade area in 2015.
+```{r}
+av_tt15_lm <- lm(avg.tt ~ leaf, data = av_tt15_mean)
+summary(av_tt15_lm)
+```
+
+```{r}
+av_tt15_seg <- segmented(av_tt15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_tt15_seg)
+```
+
+```{r}
+av_tt15_seg$psi
+slope(av_tt15_seg)
+av_tt15_fitted <- fitted(av_tt15_seg)
+av_tt15_model <- data.frame(Leaf = av_tt15_mean$leaf, TT = av_tt15_fitted)
+ggplot(av_tt15_model, aes(x = Leaf, y = TT)) + geom_line()
+av_tt15_lines <- av_tt15_seg$psi[, 2]
+av_tt15_mean_break <- av_tt15_plot + geom_vline(xintercept = av_tt15_lines, linetype = "dashed")
+av_tt15_mean_break
+ggsave("av_tt15_break.png")
+```
+
+We tested leaf number and compactness for 2013.
+```{r}
+av_cp13_plot <- ggplot(av_13, aes(x = leaf, y = compactness)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_cp13_plot <- av_cp13_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+av_cp13_plot
+
+```
+
+```{r}
+av_cp13_mean_plot <- ggplot(av_cp13_mean, aes(x = leaf, y = avg.cp)) + geom_line() + theme_classic()
+av_cp13_mean_plot <- av_cp13_mean_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+av_cp13_mean_plot
+```
+
+We ran a LM for leaf number and compactness in 2013.
+```{r}
+av_cp13_lm <- lm(avg.cp ~ leaf, data = av_cp13_mean)
+summary(av_cp13_lm)
+```
+
+```{r}
+av_cp13_seg <- segmented(av_cp13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_cp13_seg)
+```
+
+```{r}
+av_cp13_seg$psi
+slope(av_cp13_seg)
+av_cp13_fitted <- fitted(av_cp13_seg)
+av_cp13_model <- data.frame(Leaf = av_cp13_mean$leaf, CP = av_cp13_fitted)
+ggplot(av_cp13_model, aes(x = Leaf, y = CP)) + geom_line()
+av_cp13_lines <- av_cp13_seg$psi[, 2]
+av_cp13_mean_break <- av_cp13_plot + geom_vline(xintercept = av_cp13_lines, linetype = "dashed")
+av_cp13_mean_break
+ggsave("av_cp13_break.png")
+```
+
+We tested leaf number and compactness for 2015.
+```{r}
+av_cp15_plot <- ggplot(av_15, aes(x = leaf, y = compactness)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_cp15_plot <- av_cp15_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+av_cp15_plot
+```
+
+```{r}
+av_cp15_mean_plot <- ggplot(av_cp15_mean, aes(x = leaf, y = avg.cp)) + geom_line() + theme_classic()
+av_cp15_mean_plot <- av_cp15_mean_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+av_cp15_mean_plot
+```
+
+We ran a LM for leaf number and compactness in 2015.
+```{r}
+av_cp15_lm <- lm(avg.cp ~ leaf, data = av_cp15_mean)
+summary(av_cp15_lm)
+```
+
+```{r}
+av_cp15_seg <- segmented(av_cp15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_cp15_seg)
+```
+
+```{r}
+av_cp15_seg$psi
+slope(av_cp15_seg)
+av_cp15_fitted <- fitted(av_cp15_seg)
+av_cp15_model <- data.frame(Leaf = av_cp15_mean$leaf, CP = av_cp15_fitted)
+ggplot(av_cp15_model, aes(x = Leaf, y = CP)) + geom_line()
+av_cp15_lines <- av_cp15_seg$psi[, 2]
+av_cp15_mean_break <- av_cp15_plot + geom_vline(xintercept = av_cp15_lines, linetype = "dashed")
+av_cp15_mean_break
+ggsave("av_cp15_break.png")
+```
+
+We tested leaf number and shape factor for 2013.
+```{r}
+av_sf13_plot <- ggplot(av_13, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_sf13_plot <- av_sf13_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+av_sf13_plot
+```
+
+```{r}
+av_sf13_mean_plot <- ggplot(av_sf13_mean, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
+av_sf13_mean_plot <- av_sf13_mean_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+av_sf13_mean_plot
+```
+
+We ran a LM for leaf number and shape factor in 2013.
+```{r}
+av_sf13_lm <- lm(avg.sf ~ leaf, data = av_sf13_mean)
+summary(av_sf13_lm)
+```
+
+```{r}
+av_sf13_seg <- segmented(av_sf13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(av_sf13_seg)
+```
+
+```{r}
+av_sf13_seg$psi
+slope(av_sf13_seg)
+av_sf13_fitted <- fitted(av_sf13_seg)
+av_sf13_model <- data.frame(Leaf = av_sf13_mean$leaf, SF = av_sf13_fitted)
+ggplot(av_sf13_model, aes(x = Leaf, y = SF)) + geom_line()
+av_sf13_lines <- av_sf13_seg$psi[, 2]
+av_sf13_mean_break <- av_sf13_plot + geom_vline(xintercept = av_sf13_lines, linetype = "dashed")
+av_sf13_mean_break
+ggsave("av_sf13_break.png")
+```
+
+We tested leaf number and shape factor for 2015.
+```{r}
+av_sf15_plot <- ggplot(av_15, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_sf15_plot <- av_sf15_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+av_sf15_plot
+
+```
+
+```{r}
+av_sf15_mean_plot <- ggplot(av_sf15_mean, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
+av_sf15_mean_plot <- av_sf15_mean_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+av_sf15_mean_plot
+```
+
+We ran a LM for leaf number and shape factor in 2015.
+```{r}
+av_sf15_lm <- lm(avg.sf ~ leaf, data = av_sf15_mean)
+summary(av_sf15_lm)
+```
+
+```{r}
+av_sf15_seg <- segmented(av_sf15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(5, 8, 11)))
+summary(av_sf15_seg)
+```
+
+```{r}
+av_sf15_seg$psi
+slope(av_sf15_seg)
+av_sf15_fitted <- fitted(av_sf15_seg)
+av_sf15_model <- data.frame(Leaf = av_sf15_mean$leaf, SF = av_sf15_fitted)
+ggplot(av_sf15_model, aes(x = Leaf, y = SF)) + geom_line()
+av_sf15_lines <- av_sf15_seg$psi[, 2]
+av_sf15_mean_break <- av_sf15_plot + geom_vline(xintercept = av_sf15_lines, linetype = "dashed")
+av_sf15_mean_break
+ggsave("av_sf15_break.png")
+```
+
+Next we tested *Vitis amurensis*.
 
@@
 
-First, we tested *Vitis riparia*.
+Finally we tested *Vitis riparia*.
 

### 2018-10-26 Vitis_GLM.Rmd
V. amurensis and riparia set up

@@
 
-We ran a GLM for leaf number and total teeth.
-```{r}
-amurensis <-filter(R_DiLP_all_species, species == "amurensis")
-```
+We need the means for each leaf character for both years combined.
+```{r}
+amurensis <-filter(Vitis_90_ratios, species == "amurensis")
+am_l_means <-  amurensis %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tt = mean(total.teeth),
+            avg.fdr = mean(feret.diameter.ratio),
+            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
+            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
+            avg.ta = mean(average.tooth.area, na.rm = TRUE),
+            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
+            avg.tp = mean(teeth.perimeter),
+            avg.tip = mean(teeth.int.perimeter),
+            avg.pa = mean(perimeter.area),
+            avg.pr = mean(perimeter.ratio),
+            avg.cp = mean(compactness),
+            avg.sf = mean(shape.factor),
+            avg.tba = mean(teeth.blade.area))
+```
+
+Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF), so we need to split these variables for each year.
 
@@
 
-We ran a GLM for leaf number and total teeth.
-```{r}
-riparia <-filter(R_DiLP_all_species, species == "riparia")
-```
+We need the means for each leaf character for both years combined.
+```{r}
+riparia <-filter(Vitis_90_ratios, species == "riparia")
+r_l_means <-  riparia %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tt = mean(total.teeth),
+            avg.fdr = mean(feret.diameter.ratio),
+            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
+            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
+            avg.ta = mean(average.tooth.area, na.rm = TRUE),
+            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
+            avg.tp = mean(teeth.perimeter),
+            avg.tip = mean(teeth.int.perimeter),
+            avg.pa = mean(perimeter.area),
+            avg.pr = mean(perimeter.ratio),
+            avg.cp = mean(compactness),
+            avg.sf = mean(shape.factor),
+            avg.tba = mean(teeth.blade.area))
+```
+
+Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF), so we need to split these variables for each year.

### 2018-10-26 V_acerifolia_LM.Rmd
Set up new documents

--- 2018-10-23 Vitis_GLM.Rmd
@@
 ---
-title: "Paper 1 GLM"
+title: "Vitis acerifolia LM"
 output: html_notebook
@@
 ```
-
-We then tested *Vitis aestivalis*.
-
-```{r}
-aestivalis <-filter(Vitis_90_ratios, species == "aestivalis")
-```
-
-We need the means for each leaf character for both years combined.
-```{r}
-av_l_means <-  aestivalis %>% 
-  group_by(leaf) %>% 
-  summarize(avg.tt = mean(total.teeth),
-            avg.fdr = mean(feret.diameter.ratio, na.rm = TRUE),
-            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
-            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
-            avg.ta = mean(average.tooth.area, na.rm = TRUE),
-            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
-            avg.tp = mean(teeth.perimeter),
-            avg.tip = mean(teeth.int.perimeter),
-            avg.pa = mean(perimeter.area),
-            avg.pr = mean(perimeter.ratio),
-            avg.cp = mean(compactness),
-            avg.sf = mean(shape.factor),
-            avg.tba = mean(teeth.blade.area))
-skim(av_l_means)
-```
-
-Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF), so we need to split these variables for each year.
-```{r}
-av_13 <- filter(aestivalis, year == 2013)
-av_15 <- filter(aestivalis, year == 2015)
-```
-
-Total teeth
-```{r}
-av_tt13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
-
-ac_tt15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
-```
-
-Tooth area: blade area
-```{r}
-av_taba13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-
-ac_taba15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-```
-
-Compactness
-```{r}
-av_cp13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.cp = mean(compactness, na.rm = TRUE))
-
-ac_cp15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.cp = mean(compactness, na.rm = TRUE))
-```
-
-Shape factor
-```{r}
-av_sf13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
-
-ac_sf15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
-```
-
-We wanted to test leaf number and feret diameter ratio.
-```{r}
-av_fdr_plot <- ggplot(aestivalis, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_fdr_plot <- av_fdr_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-av_fdr_plot
-```
-
-```{r}
-av_fdr_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
-av_fdr_mean_plot <- av_fdr_mean_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-av_fdr_mean_plot
-```
-
-We ran a LM for leaf number and feret diamter ratio.
-```{r}
-av_fdr_lm <- lm(avg.fdr ~ leaf, data = av_l_means)
-summary(av_fdr_lm)
-```
-
-```{r}
-av_fdr_seg <- segmented(av_fdr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_fdr_seg)
-```
-
-```{r}
-av_fdr_seg$psi
-slope(av_fdr_seg)
-av_fdr_fitted <- fitted(av_fdr_seg)
-av_fdr_model <- data.frame(Leaf = av_l_means$leaf, FDR = av_fdr_fitted)
-ggplot(av_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
-av_fdr_lines <- av_fdr_seg$psi[, 2]
-av_fdr_mean_break <- av_fdr_plot + geom_vline(xintercept = av_fdr_lines, linetype = "dashed")
-av_fdr_mean_break
-ggsave("av_fdr_break.png")
-```
-
-We tested leaf number and tooth area:perimeter.
-```{r}
-av_tap_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tap_plot <- av_tap_plot +labs(x = "Leaf number",
-                               y = "Tooth area: perimeter")
-av_tap_plot
-```
-
-```{r}
-av_tap_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
-av_tap_mean_plot <- av_tap_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Perimeter")
-av_tap_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:perimeter.
-```{r}
-av_tap_lm <- lm(avg.tap ~ leaf, data = av_l_means)
-summary(av_tap_lm)
-```
-
-```{r}
-av_tap_seg <- segmented(av_tap_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(4, 7, 10)))
-summary(av_tap_seg)
-```
-
-```{r}
-av_tap_seg$psi
-slope(av_tap_seg)
-av_tap_fitted <- fitted(av_tap_seg)
-av_tap_model <- data.frame(Leaf = av_l_means$leaf, TAP = av_tap_fitted)
-ggplot(av_tap_model, aes(x = Leaf, y = TAP)) + geom_line()
-av_tap_lines <- av_tap_seg$psi[, 2]
-av_tap_mean_break <- av_tap_plot + geom_vline(xintercept = av_tap_lines, linetype = "dashed")
-av_tap_mean_break
-ggsave("av_tap_break.png")
-```
-
-We tested leaf number and tooth area:internal perimeter.
-```{r}
-av_taip_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_taip_plot <- av_taip_plot +labs(x = "Leaf number",
-                               y = "Tooth area: internal perimeter")
-av_taip_plot
-```
-
-```{r}
-av_taip_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
-av_taip_mean_plot <- av_taip_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Internal Perimeter")
-av_taip_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:internal perimeter.
-```{r}
-av_taip_lm <- lm(avg.taip ~ leaf, data = av_l_means)
-summary(av_taip_lm)
-```
-
-```{r}
-av_taip_seg <- segmented(av_taip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_taip_seg)
-```
-
-```{r}
-av_taip_seg$psi
-slope(av_taip_seg)
-av_taip_fitted <- fitted(av_taip_seg)
-av_taip_model <- data.frame(Leaf = av_l_means$leaf, TAIP = av_taip_fitted)
-ggplot(av_taip_model, aes(x = Leaf, y = TAIP)) + geom_line()
-av_taip_lines <- av_taip_seg$psi[, 2]
-av_taip_mean_break <- av_taip_plot + geom_vline(xintercept = av_taip_lines, linetype = "dashed")
-av_taip_mean_break
-ggsave("av_taip_break.png")
-```
-
-We tested leaf number and average tooth area.
-```{r}
-av_ta_plot <- ggplot(aestivalis, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_ta_plot <- av_ta_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-av_ta_plot
-
-```
-
-```{r}
-av_ta_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.ta)) + geom_line() + theme_classic()
-av_ta_mean_plot <- av_ta_mean_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-av_ta_mean_plot
-```
-
-We ran a LM for leaf number and average tooth area.
-```{r}
-av_ta_lm <- lm(avg.ta ~ leaf, data = av_l_means)
-summary(av_ta_lm)
-```
-
-```{r}
-av_ta_seg <- segmented(av_ta_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_ta_seg)
-```
-
-```{r}
-av_ta_seg$psi
-slope(av_ta_seg)
-av_ta_fitted <- fitted(av_ta_seg)
-av_ta_model <- data.frame(Leaf = av_l_means$leaf, TA = av_ta_fitted)
-ggplot(av_ta_model, aes(x = Leaf, y = TA)) + geom_line()
-av_ta_lines <- av_ta_seg$psi[, 2]
-av_ta_mean_break <- av_ta_plot + geom_vline(xintercept = av_ta_lines, linetype = "dashed")
-av_ta_mean_break
-ggsave("av_ta_break.png")
-```
-
-We tested leaf number and teeth:perimeter.
-```{r}
-av_tp_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tp_plot <- av_tp_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-av_tp_plot
-
-```
-
-```{r}
-av_tp_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
-av_tp_mean_plot <- av_tp_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-av_tp_mean_plot
-```
-
-We ran a LM for leaf number and teeth:perimeter.
-```{r}
-av_tp_lm <- lm(avg.tp ~ leaf, data = av_l_means)
-summary(av_tp_lm)
-```
-
-```{r}
-av_tp_seg <- segmented(av_tp_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_tp_seg)
-```
-
-```{r}
-av_tp_seg$psi
-slope(av_tp_seg)
-av_tp_fitted <- fitted(av_tp_seg)
-av_tp_model <- data.frame(Leaf = av_l_means$leaf, TP = av_tp_fitted)
-ggplot(av_tp_model, aes(x = Leaf, y = TP)) + geom_line()
-av_tp_lines <- av_tp_seg$psi[, 2]
-av_tp_mean_break <- av_tp_plot + geom_vline(xintercept = av_tp_lines, linetype = "dashed")
-av_tp_mean_break
-ggsave("av_tp_break.png")
-```
-
-We tested leaf number and teeth:internal perimeter.
-```{r}
-av_tip_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tip_plot <- av_tip_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-av_tip_plot
-
-```
-
-```{r}
-av_tip_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
-av_tip_mean_plot <- av_tip_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-av_tip_mean_plot
-```
-
-We ran a LM for leaf number and teeth:internal perimeter.
-```{r}
-av_tip_lm <- lm(avg.tip ~ leaf, data = av_l_means)
-summary(av_tip_lm)
-```
-
-```{r}
-av_tip_seg <- segmented(av_tip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_tip_seg)
-```
-
-```{r}
-av_tip_seg$psi
-slope(av_tip_seg)
-av_tip_fitted <- fitted(av_tip_seg)
-av_tip_model <- data.frame(Leaf = av_l_means$leaf, TIP = av_tip_fitted)
-ggplot(av_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
-av_tip_lines <- av_tip_seg$psi[, 2]
-av_tip_mean_break <- av_tip_plot + geom_vline(xintercept = av_tip_lines, linetype = "dashed")
-av_tip_mean_break
-ggsave("av_tip_break.png")
-```
-
-We tested leaf number and perimeter:area.
-```{r}
-av_pa_plot <- ggplot(aestivalis, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_pa_plot <- av_pa_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-av_pa_plot
-
-```
-
-```{r}
-av_pa_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
-av_pa_mean_plot <- av_pa_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-av_pa_mean_plot
-```
-
-We ran a LM for leaf number and perimeter:area.
-```{r}
-av_pa_lm <- lm(avg.pa ~ leaf, data = av_l_means)
-summary(av_pa_lm)
-```
-
-```{r}
-av_pa_seg <- segmented(av_pa_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(2, 7, 10)))
-summary(av_pa_seg)
-```
-
-```{r}
-ac_pa_seg$psi
-slope(ac_pa_seg)
-ac_pa_fitted <- fitted(ac_pa_seg)
-ac_pa_model <- data.frame(Leaf = ac_l_means$leaf, PA = ac_pa_fitted)
-ggplot(ac_pa_model, aes(x = Leaf, y = PA)) + geom_line()
-ac_pa_lines <- ac_pa_seg$psi[, 2]
-ac_pa_mean_break <- ac_pa_plot + geom_vline(xintercept = ac_pa_lines, linetype = "dashed")
-ac_pa_mean_break
-ggsave("ac_pa_break.png")
-```
-
-We tested leaf number and perimeter ratio.
-```{r}
-av_pr_plot <- ggplot(aestivalis, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_pr_plot <- av_pr_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-av_pr_plot
-```
-
-```{r}
-av_pr_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
-av_pr_mean_plot <- av_pr_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-av_pr_mean_plot
-```
-
-We ran a LM for leaf number and perimeter ratio.
-```{r}
-av_pr_lm <- lm(avg.pr ~ leaf, data = av_l_means)
-summary(av_pr_lm)
-```
-
-```{r}
-av_pr_seg <- segmented(av_pr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_pr_seg)
-```
-
-```{r}
-av_pr_seg$psi
-slope(av_pr_seg)
-av_pr_fitted <- fitted(av_pr_seg)
-av_pr_model <- data.frame(Leaf = av_l_means$leaf, PR = av_pr_fitted)
-ggplot(av_pr_model, aes(x = Leaf, y = PR)) + geom_line()
-av_pr_lines <- av_pr_seg$psi[, 2]
-av_pr_mean_break <- av_pr_plot + geom_vline(xintercept = av_pr_lines, linetype = "dashed")
-av_pr_mean_break
-ggsave("av_pr_break.png")
-```
-
-We tested leaf number and teeth:blade area.
-```{r}
-av_tba_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tba_plot <- av_tba_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-av_tba_plot
-
-```
-
-```{r}
-av_tba_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
-av_tba_mean_plot <- av_tba_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-av_tba_mean_plot
-```
-
-We ran a LM for leaf number and teeth:blade area.
-```{r}
-av_tba_lm <- lm(avg.tba ~ leaf, data = av_l_means)
-summary(av_tba_lm)
-```
-
-```{r}
-av_tba_seg <- segmented(av_tba_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(2, 4, 6)))
-summary(av_tba_seg)
-```
-
-```{r}
-av_tba_seg$psi
-slope(av_tba_seg)
-av_tba_fitted <- fitted(av_tba_seg)
-av_tba_model <- data.frame(Leaf = av_l_means$leaf, TBA = av_tba_fitted)
-ggplot(av_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
-av_tba_lines <- av_tba_seg$psi[, 2]
-av_tba_mean_break <- av_tba_plot + geom_vline(xintercept = av_tba_lines, linetype = "dashed")
-av_tba_mean_break
-ggsave("av_tba_break.png")
-```
-
-First, we tested *Vitis amurensis*.
-
-We ran a GLM for leaf number and total teeth.
-```{r}
-amurensis <-filter(R_DiLP_all_species, species == "amurensis")
-```
-
-First, we tested *Vitis riparia*.
-
-We ran a GLM for leaf number and total teeth.
-```{r}
-riparia <-filter(R_DiLP_all_species, species == "riparia")
-```

### 2018-10-26 V_aestivalis_LM.Rmd
Set up new documents

--- 2018-10-24 Vitis_GLM.Rmd
@@
 ---
-title: "Paper 1 GLM"
+title: "Vitis aestivalis LM"
 output: html_notebook
@@
 
-First, we tested *Vitis acerifolia*.
-
-We need the means for each leaf character for both years combined.
-```{r}
-acerifolia <-filter(Vitis_90_ratios, species == "acerifolia")
-ac_l_means <-  acerifolia %>% 
-  group_by(leaf) %>% 
-  summarize(avg.tt = mean(total.teeth),
-            avg.fdr = mean(feret.diameter.ratio),
-            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
-            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
-            avg.ta = mean(average.tooth.area, na.rm = TRUE),
-            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
-            avg.tp = mean(teeth.perimeter),
-            avg.tip = mean(teeth.int.perimeter),
-            avg.pa = mean(perimeter.area),
-            avg.pr = mean(perimeter.ratio),
-            avg.cp = mean(compactness),
-            avg.sf = mean(shape.factor),
-            avg.tba = mean(teeth.blade.area))
-```
-
-However we need to split tooth area:blade area for each year.
-```{r}
-ac_13 <- filter(acerifolia, year == 2013)
-ac_15 <- filter(acerifolia, year == 2015)
-```
-
-```{r}
-ac_taba13_mean <-  acerifolia %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-
-ac_taba15_mean <- acerifolia %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-```
-
-We tested leaf number and total teeth.
-```{r}
-skim(acerifolia, total.teeth)
-ac_tt_plot <- ggplot(acerifolia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tt_plot <- ac_tt_plot +labs(x = "Leaf number",
-                               y = "Total teeth")
-ac_tt_plot
-
-```
-
-```{r}
-ac_tt_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
-ac_tt_mean_plot <- ac_tt_mean_plot +labs(x = "Leaf number",
-                               y = "Total teeth")
-ac_tt_mean_plot
-```
-
-We ran a LM for leaf number and total teeth.
-```{r}
-ac_tt_lm <- lm(avg.tt ~ leaf, data = ac_l_means)
-summary(ac_tt_lm)
-```
-
-```{r}
-ac_tt_seg <- segmented(ac_tt_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tt_seg)
-```
-
-```{r}
-ac_tt_seg$psi
-slope(ac_tt_seg)
-ac_tt_fitted <- fitted(ac_tt_seg)
-ac_tt_model <- data.frame(Leaf = ac_l_means$leaf, Teeth = ac_tt_fitted)
-ggplot(ac_tt_model, aes(x = Leaf, y = Teeth)) + geom_line()
-ac_tt_lines <- ac_tt_seg$psi[, 2]
-ac_tt_mean_break <- ac_tt_plot + geom_vline(xintercept = ac_tt_lines, linetype = "dashed")
-ac_tt_mean_break
-ggsave("ac_tt_break.png")
-```
-
-We wanted to test leaf number and feret diameter ratio.
-```{r}
-skim(acerifolia, total.teeth)
-ac_fdr_plot <- ggplot(acerifolia, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_fdr_plot <- ac_fdr_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-ac_fdr_plot
-
-```
-
-```{r}
-ac_fdr_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
-ac_fdr_mean_plot <- ac_fdr_mean_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-ac_fdr_mean_plot
-```
-
-We ran a LM for leaf number and feret diamter ratio.
-```{r}
-ac_fdr_lm <- lm(avg.fdr ~ leaf, data = ac_l_means)
-summary(ac_fdr_lm)
-```
-
-```{r}
-ac_fdr_seg <- segmented(ac_fdr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_fdr_seg)
-```
-
-```{r}
-ac_fdr_seg$psi
-slope(ac_fdr_seg)
-ac_fdr_fitted <- fitted(ac_fdr_seg)
-ac_fdr_model <- data.frame(Leaf = ac_l_means$leaf, FDR = ac_fdr_fitted)
-ggplot(ac_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
-ac_fdr_lines <- ac_fdr_seg$psi[, 2]
-ac_fdr_mean_break <- ac_fdr_plot + geom_vline(xintercept = ac_fdr_lines, linetype = "dashed")
-ac_fdr_mean_break
-ggsave("ac_fdr_break.png")
-```
-
-We tested leaf number and tooth area:perimeter.
-```{r}
-skim(acerifolia, tooth.area.perimeter)
-ac_tap_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tap_plot <- ac_tap_plot +labs(x = "Leaf number",
-                               y = "Tooth area: perimeter")
-ac_tap_plot
-
-```
-
-```{r}
-ac_tap_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
-ac_tap_mean_plot <- ac_tap_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Perimeter")
-ac_tap_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:perimeter.
-```{r}
-ac_tap_lm <- lm(avg.tap ~ leaf, data = ac_l_means)
-summary(ac_tap_lm)
-```
-
-```{r}
-ac_tap_seg <- segmented(ac_tap_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tap_seg)
-```
-
-```{r}
-ac_tap_seg$psi
-slope(ac_tap_seg)
-ac_tap_fitted <- fitted(ac_tap_seg)
-ac_tap_model <- data.frame(Leaf = ac_l_means$leaf, TAP = ac_tap_fitted)
-ggplot(ac_tap_model, aes(x = Leaf, y = TAP)) + geom_line()
-ac_tap_lines <- ac_tap_seg$psi[, 2]
-ac_tap_mean_break <- ac_tap_plot + geom_vline(xintercept = ac_tap_lines, linetype = "dashed")
-ac_tap_mean_break
-ggsave("ac_tap_break.png")
-```
-
-We tested leaf number and tooth area:internal perimeter.
-```{r}
-skim(acerifolia, tooth.area.int.perimeter)
-ac_taip_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_taip_plot <- ac_taip_plot +labs(x = "Leaf number",
-                               y = "Tooth area: internal perimeter")
-ac_taip_plot
-
-```
-
-```{r}
-ac_taip_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
-ac_taip_mean_plot <- ac_taip_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Internal Perimeter")
-ac_taip_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:internal perimeter.
-```{r}
-ac_taip_lm <- lm(avg.taip ~ leaf, data = ac_l_means)
-summary(ac_taip_lm)
-```
-
-```{r}
-ac_taip_seg <- segmented(ac_taip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_taip_seg)
-```
-
-```{r}
-ac_taip_seg$psi
-slope(ac_taip_seg)
-ac_taip_fitted <- fitted(ac_taip_seg)
-ac_taip_model <- data.frame(Leaf = ac_l_means$leaf, TAIP = ac_taip_fitted)
-ggplot(ac_taip_model, aes(x = Leaf, y = TAIP)) + geom_line()
-ac_taip_lines <- ac_taip_seg$psi[, 2]
-ac_taip_mean_break <- ac_taip_plot + geom_vline(xintercept = ac_taip_lines, linetype = "dashed")
-ac_taip_mean_break
-ggsave("ac_taip_break.png")
-```
-
-We tested leaf number and average tooth area.
-```{r}
-skim(acerifolia, average.tooth.area)
-ac_ta_plot <- ggplot(acerifolia, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_ta_plot <- ac_ta_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-ac_ta_plot
-
-```
-
-```{r}
-ac_ta_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.ta)) + geom_line() + theme_classic()
-ac_ta_mean_plot <- ac_ta_mean_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-ac_ta_mean_plot
-```
-
-We ran a LM for leaf number and average tooth area.
-```{r}
-ac_ta_lm <- lm(avg.ta ~ leaf, data = ac_l_means)
-summary(ac_ta_lm)
-```
-
-```{r}
-ac_ta_seg <- segmented(ac_ta_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_ta_seg)
-```
-
-```{r}
-ac_ta_seg$psi
-slope(ac_ta_seg)
-ac_ta_fitted <- fitted(ac_ta_seg)
-ac_ta_model <- data.frame(Leaf = ac_l_means$leaf, TA = ac_ta_fitted)
-ggplot(ac_ta_model, aes(x = Leaf, y = TA)) + geom_line()
-ac_ta_lines <- ac_ta_seg$psi[, 2]
-ac_ta_mean_break <- ac_ta_plot + geom_vline(xintercept = ac_ta_lines, linetype = "dashed")
-ac_ta_mean_break
-ggsave("ac_ta_break.png")
-```
-
-We tested leaf number and teeth:perimeter.
-```{r}
-skim(acerifolia, teeth.perimeter)
-ac_tp_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tp_plot <- ac_tp_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-ac_tp_plot
-
-```
-
-```{r}
-ac_tp_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
-ac_tp_mean_plot <- ac_tp_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-ac_tp_mean_plot
-```
-
-We ran a LM for leaf number and teeth:perimeter.
-```{r}
-ac_tp_lm <- lm(avg.tp ~ leaf, data = ac_l_means)
-summary(ac_tp_lm)
-```
-
-```{r}
-ac_tp_seg <- segmented(ac_tp_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tp_seg)
-```
-
-```{r}
-ac_tp_seg$psi
-slope(ac_tp_seg)
-ac_tp_fitted <- fitted(ac_tp_seg)
-ac_tp_model <- data.frame(Leaf = ac_l_means$leaf, TP = ac_tp_fitted)
-ggplot(ac_tp_model, aes(x = Leaf, y = TP)) + geom_line()
-ac_tp_lines <- ac_tp_seg$psi[, 2]
-ac_tp_mean_break <- ac_tp_plot + geom_vline(xintercept = ac_tp_lines, linetype = "dashed")
-ac_tp_mean_break
-ggsave("ac_tp_break.png")
-```
-
-We tested leaf number and teeth:internal perimeter.
-```{r}
-skim(acerifolia, teeth.int.perimeter)
-ac_tip_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tip_plot <- ac_tip_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-ac_tip_plot
-
-```
-
-```{r}
-ac_tip_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
-ac_tip_mean_plot <- ac_tip_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-ac_tip_mean_plot
-```
-
-We ran a LM for leaf number and teeth:internal perimeter.
-```{r}
-ac_tip_lm <- lm(avg.tip ~ leaf, data = ac_l_means)
-summary(ac_tip_lm)
-```
-
-```{r}
-ac_tip_seg <- segmented(ac_tip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tip_seg)
-```
-
-```{r}
-ac_tip_seg$psi
-slope(ac_tip_seg)
-ac_tip_fitted <- fitted(ac_tip_seg)
-ac_tip_model <- data.frame(Leaf = ac_l_means$leaf, TIP = ac_tip_fitted)
-ggplot(ac_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
-ac_tip_lines <- ac_tip_seg$psi[, 2]
-ac_tip_mean_break <- ac_tip_plot + geom_vline(xintercept = ac_tip_lines, linetype = "dashed")
-ac_tip_mean_break
-ggsave("ac_tip_break.png")
-```
-
-We tested leaf number and perimeter:area.
-```{r}
-skim(acerifolia, perimeter.area)
-ac_pa_plot <- ggplot(acerifolia, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_pa_plot <- ac_pa_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-ac_pa_plot
-
-```
-
-```{r}
-ac_pa_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
-ac_pa_mean_plot <- ac_pa_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-ac_pa_mean_plot
-```
-
-We ran a LM for leaf number and perimeter:area.
-```{r}
-ac_pa_lm <- lm(avg.pa ~ leaf, data = ac_l_means)
-summary(ac_pa_lm)
-```
-
-```{r}
-ac_pa_seg <- segmented(ac_pa_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_pa_seg)
-```
-
-```{r}
-ac_pa_seg$psi
-slope(ac_pa_seg)
-ac_pa_fitted <- fitted(ac_pa_seg)
-ac_pa_model <- data.frame(Leaf = ac_l_means$leaf, PA = ac_pa_fitted)
-ggplot(ac_pa_model, aes(x = Leaf, y = PA)) + geom_line()
-ac_pa_lines <- ac_pa_seg$psi[, 2]
-ac_pa_mean_break <- ac_pa_plot + geom_vline(xintercept = ac_pa_lines, linetype = "dashed")
-ac_pa_mean_break
-ggsave("ac_pa_break.png")
-```
-
-We tested leaf number and perimeter ratio.
-```{r}
-skim(acerifolia, perimeter.ratio)
-ac_pr_plot <- ggplot(acerifolia, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_pr_plot <- ac_pr_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-ac_pr_plot
-
-```
-
-```{r}
-ac_pr_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
-ac_pr_mean_plot <- ac_pr_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-ac_pr_mean_plot
-```
-
-We ran a LM for leaf number and perimeter ratio.
-```{r}
-ac_pr_lm <- lm(avg.pr ~ leaf, data = ac_l_means)
-summary(ac_pr_lm)
-```
-
-```{r}
-ac_pr_seg <- segmented(ac_pr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_pr_seg)
-```
-
-```{r}
-ac_pr_seg$psi
-slope(ac_pr_seg)
-ac_pr_fitted <- fitted(ac_pr_seg)
-ac_pr_model <- data.frame(Leaf = ac_l_means$leaf, PR = ac_pr_fitted)
-ggplot(ac_pr_model, aes(x = Leaf, y = PR)) + geom_line()
-ac_pr_lines <- ac_pr_seg$psi[, 2]
-ac_pr_mean_break <- ac_pr_plot + geom_vline(xintercept = ac_pr_lines, linetype = "dashed")
-ac_pr_mean_break
-ggsave("ac_pr_break.png")
-```
-
-We tested leaf number and compactness.
-```{r}
-skim(acerifolia, compactness)
-ac_cp_plot <- ggplot(acerifolia, aes(x = leaf, y = compactness)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_cp_plot <- ac_cp_plot +labs(x = "Leaf number",
-                               y = "Compactness")
-ac_cp_plot
-
-```
-
-```{r}
-ac_cp_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.cp)) + geom_line() + theme_classic()
-ac_cp_mean_plot <- ac_cp_mean_plot +labs(x = "Leaf number",
-                               y = "Compactness")
-ac_cp_mean_plot
-```
-
-We ran a LM for leaf number and compactness.
-```{r}
-ac_cp_lm <- lm(avg.cp ~ leaf, data = ac_l_means)
-summary(ac_cp_lm)
-```
-
-```{r}
-ac_cp_seg <- segmented(ac_cp_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_cp_seg)
-```
-
-```{r}
-ac_cp_seg$psi
-slope(ac_cp_seg)
-ac_cp_fitted <- fitted(ac_cp_seg)
-ac_cp_model <- data.frame(Leaf = ac_l_means$leaf, CP = ac_cp_fitted)
-ggplot(ac_cp_model, aes(x = Leaf, y = CP)) + geom_line()
-ac_cp_lines <- ac_cp_seg$psi[, 2]
-ac_cp_mean_break <- ac_cp_plot + geom_vline(xintercept = ac_cp_lines, linetype = "dashed")
-ac_cp_mean_break
-ggsave("ac_cp_break.png")
-```
-
-We tested leaf number and shape factor.
-```{r}
-skim(acerifolia, shape.factor)
-ac_sf_plot <- ggplot(acerifolia, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_sf_plot <- ac_sf_plot +labs(x = "Leaf number",
-                               y = "Shape Factor")
-ac_sf_plot
-
-```
-
-```{r}
-ac_sf_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
-ac_sf_mean_plot <- ac_sf_mean_plot +labs(x = "Leaf number",
-                               y = "Shape Factor")
-ac_sf_mean_plot
-```
-
-We ran a LM for leaf number and shape factor.
-```{r}
-ac_sf_lm <- lm(avg.sf ~ leaf, data = ac_l_means)
-summary(ac_sf_lm)
-```
-
-```{r}
-ac_sf_seg <- segmented(ac_sf_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_sf_seg)
-```
-
-```{r}
-ac_sf_seg$psi
-slope(ac_sf_seg)
-ac_sf_fitted <- fitted(ac_sf_seg)
-ac_sf_model <- data.frame(Leaf = ac_l_means$leaf, SF = ac_sf_fitted)
-ggplot(ac_sf_model, aes(x = Leaf, y = SF)) + geom_line()
-ac_sf_lines <- ac_sf_seg$psi[, 2]
-ac_sf_mean_break <- ac_sf_plot + geom_vline(xintercept = ac_sf_lines, linetype = "dashed")
-ac_sf_mean_break
-ggsave("ac_sf_break.png")
-```
-
-We tested leaf number and teeth:blade area.
-```{r}
-skim(acerifolia, teeth.blade.area)
-ac_tba_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tba_plot <- ac_tba_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-ac_tba_plot
-
-```
-
-```{r}
-ac_tba_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
-ac_tba_mean_plot <- ac_tba_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-ac_tba_mean_plot
-```
-
-We ran a LM for leaf number and teeth:blade area.
-```{r}
-ac_tba_lm <- lm(avg.tba ~ leaf, data = ac_l_means)
-summary(ac_tba_lm)
-```
-
-```{r}
-ac_tba_seg <- segmented(ac_tba_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tba_seg)
-```
-
-```{r}
-ac_tba_seg$psi
-slope(ac_tba_seg)
-ac_tba_fitted <- fitted(ac_tba_seg)
-ac_tba_model <- data.frame(Leaf = ac_l_means$leaf, TBA = ac_tba_fitted)
-ggplot(ac_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
-ac_tba_lines <- ac_tba_seg$psi[, 2]
-ac_tba_mean_break <- ac_tba_plot + geom_vline(xintercept = ac_tba_lines, linetype = "dashed")
-ac_tba_mean_break
-ggsave("ac_tba_break.png")
-```
-
-We tested leaf number and tooth area:blade area for 2013.
-```{r}
-ac_taba13_plot <- ggplot(ac_13, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_taba13_plot <- ac_taba13_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba13_plot
-
-```
-
-```{r}
-ac_taba13_mean_plot <- ggplot(ac_taba13_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
-ac_taba13_mean_plot <- ac_taba13_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba13_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:blade area in 2013.
-```{r}
-ac_taba13_lm <- lm(avg.taba ~ leaf, data = ac_taba13_mean)
-summary(ac_taba13_lm)
-```
-
-```{r}
-ac_taba13_seg <- segmented(ac_taba13_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_taba13_seg)
-```
-
-```{r}
-ac_taba13_seg$psi
-slope(ac_taba13_seg)
-ac_taba13_fitted <- fitted(ac_taba13_seg)
-ac_taba13_model <- data.frame(Leaf = ac_taba13_mean$leaf, TABA = ac_taba13_fitted)
-ggplot(ac_taba13_model, aes(x = Leaf, y = TABA)) + geom_line()
-ac_taba13_lines <- ac_taba13_seg$psi[, 2]
-ac_taba13_mean_break <- ac_taba13_plot + geom_vline(xintercept = ac_taba13_lines, linetype = "dashed")
-ac_taba13_mean_break
-ggsave("ac_taba13_break.png")
-```
-
-We tested leaf number and tooth area:blade area for 2015.
-```{r}
-ac_taba15_plot <- ggplot(ac_15, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_taba15_plot <- ac_taba15_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba15_plot
-
-```
-
-```{r}
-ac_taba15_mean_plot <- ggplot(ac_taba15_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
-ac_taba15_mean_plot <- ac_taba15_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba15_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:blade area in 2015.
-```{r}
-ac_taba15_lm <- lm(avg.taba ~ leaf, data = ac_taba15_mean)
-summary(ac_taba15_lm)
-```
-
-```{r}
-ac_taba15_seg <- segmented(ac_taba15_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_taba15_seg)
-```
-
-```{r}
-ac_taba15_seg$psi
-slope(ac_taba15_seg)
-ac_taba15_fitted <- fitted(ac_taba15_seg)
-ac_taba15_model <- data.frame(Leaf = ac_taba15_mean$leaf, TABA = ac_taba15_fitted)
-ggplot(ac_taba15_model, aes(x = Leaf, y = TABA)) + geom_line()
-ac_taba15_lines <- ac_taba15_seg$psi[, 2]
-ac_taba15_mean_break <- ac_taba15_plot + geom_vline(xintercept = ac_taba15_lines, linetype = "dashed")
-ac_taba15_mean_break
-ggsave("ac_taba15_break.png")
-```
-
 We then tested *Vitis aestivalis*.
@@
 ```
-
-Next we tested *Vitis amurensis*.
-
-We ran a GLM for leaf number and total teeth.
-```{r}
-amurensis <-filter(R_DiLP_all_species, species == "amurensis")
-```
-
-Finally we tested *Vitis riparia*.
-
-We ran a GLM for leaf number and total teeth.
-```{r}
-riparia <-filter(R_DiLP_all_species, species == "riparia")
-```

### 2018-10-26 V_amurensis_LM.Rmd
Set up new documents

--- 2018-10-23 Vitis_GLM.Rmd
@@
 ---
-title: "Paper 1 GLM"
+title: "Vitis amurensis LM"
 output: html_notebook
@@
 
-First, we tested *Vitis acerifolia*.
+Next we tested *Vitis amurensis*.
 
@@
 ```{r}
-acerifolia <-filter(Vitis_90_ratios, species == "acerifolia")
-ac_l_means <-  acerifolia %>% 
+amurensis <-filter(Vitis_90_ratios, species == "amurensis")
+am_l_means <-  amurensis %>% 
   group_by(leaf) %>% 
@@
 
-However we need to split tooth area:blade area for each year.
+Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF), so we need to split these variables for each year.
 ```{r}
-ac_13 <- filter(acerifolia, year == 2013)
-ac_15 <- filter(acerifolia, year == 2015)
+am_13 <- filter(amurensis, year == 2013)
+am_15 <- filter(amurensis, year == 2015)
 ```
-
-```{r}
-ac_taba13_mean <-  acerifolia %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-
-ac_taba15_mean <- acerifolia %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-```
-
-We tested leaf number and total teeth.
-```{r}
-skim(acerifolia, total.teeth)
-ac_tt_plot <- ggplot(acerifolia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tt_plot <- ac_tt_plot +labs(x = "Leaf number",
-                               y = "Total teeth")
-ac_tt_plot
-
-```
-
-```{r}
-ac_tt_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
-ac_tt_mean_plot <- ac_tt_mean_plot +labs(x = "Leaf number",
-                               y = "Total teeth")
-ac_tt_mean_plot
-```
-
-We ran a LM for leaf number and total teeth.
-```{r}
-ac_tt_lm <- lm(avg.tt ~ leaf, data = ac_l_means)
-summary(ac_tt_lm)
-```
-
-```{r}
-ac_tt_seg <- segmented(ac_tt_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tt_seg)
-```
-
-```{r}
-ac_tt_seg$psi
-slope(ac_tt_seg)
-ac_tt_fitted <- fitted(ac_tt_seg)
-ac_tt_model <- data.frame(Leaf = ac_l_means$leaf, Teeth = ac_tt_fitted)
-ggplot(ac_tt_model, aes(x = Leaf, y = Teeth)) + geom_line()
-ac_tt_lines <- ac_tt_seg$psi[, 2]
-ac_tt_mean_break <- ac_tt_plot + geom_vline(xintercept = ac_tt_lines, linetype = "dashed")
-ac_tt_mean_break
-ggsave("ac_tt_break.png")
-```
-
-We wanted to test leaf number and feret diameter ratio.
-```{r}
-skim(acerifolia, total.teeth)
-ac_fdr_plot <- ggplot(acerifolia, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_fdr_plot <- ac_fdr_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-ac_fdr_plot
-
-```
-
-```{r}
-ac_fdr_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
-ac_fdr_mean_plot <- ac_fdr_mean_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-ac_fdr_mean_plot
-```
-
-We ran a LM for leaf number and feret diamter ratio.
-```{r}
-ac_fdr_lm <- lm(avg.fdr ~ leaf, data = ac_l_means)
-summary(ac_fdr_lm)
-```
-
-```{r}
-ac_fdr_seg <- segmented(ac_fdr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_fdr_seg)
-```
-
-```{r}
-ac_fdr_seg$psi
-slope(ac_fdr_seg)
-ac_fdr_fitted <- fitted(ac_fdr_seg)
-ac_fdr_model <- data.frame(Leaf = ac_l_means$leaf, FDR = ac_fdr_fitted)
-ggplot(ac_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
-ac_fdr_lines <- ac_fdr_seg$psi[, 2]
-ac_fdr_mean_break <- ac_fdr_plot + geom_vline(xintercept = ac_fdr_lines, linetype = "dashed")
-ac_fdr_mean_break
-ggsave("ac_fdr_break.png")
-```
-
-We tested leaf number and tooth area:perimeter.
-```{r}
-skim(acerifolia, tooth.area.perimeter)
-ac_tap_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tap_plot <- ac_tap_plot +labs(x = "Leaf number",
-                               y = "Tooth area: perimeter")
-ac_tap_plot
-
-```
-
-```{r}
-ac_tap_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
-ac_tap_mean_plot <- ac_tap_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Perimeter")
-ac_tap_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:perimeter.
-```{r}
-ac_tap_lm <- lm(avg.tap ~ leaf, data = ac_l_means)
-summary(ac_tap_lm)
-```
-
-```{r}
-ac_tap_seg <- segmented(ac_tap_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tap_seg)
-```
-
-```{r}
-ac_tap_seg$psi
-slope(ac_tap_seg)
-ac_tap_fitted <- fitted(ac_tap_seg)
-ac_tap_model <- data.frame(Leaf = ac_l_means$leaf, TAP = ac_tap_fitted)
-ggplot(ac_tap_model, aes(x = Leaf, y = TAP)) + geom_line()
-ac_tap_lines <- ac_tap_seg$psi[, 2]
-ac_tap_mean_break <- ac_tap_plot + geom_vline(xintercept = ac_tap_lines, linetype = "dashed")
-ac_tap_mean_break
-ggsave("ac_tap_break.png")
-```
-
-We tested leaf number and tooth area:internal perimeter.
-```{r}
-skim(acerifolia, tooth.area.int.perimeter)
-ac_taip_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_taip_plot <- ac_taip_plot +labs(x = "Leaf number",
-                               y = "Tooth area: internal perimeter")
-ac_taip_plot
-
-```
-
-```{r}
-ac_taip_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
-ac_taip_mean_plot <- ac_taip_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Internal Perimeter")
-ac_taip_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:internal perimeter.
-```{r}
-ac_taip_lm <- lm(avg.taip ~ leaf, data = ac_l_means)
-summary(ac_taip_lm)
-```
-
-```{r}
-ac_taip_seg <- segmented(ac_taip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_taip_seg)
-```
-
-```{r}
-ac_taip_seg$psi
-slope(ac_taip_seg)
-ac_taip_fitted <- fitted(ac_taip_seg)
-ac_taip_model <- data.frame(Leaf = ac_l_means$leaf, TAIP = ac_taip_fitted)
-ggplot(ac_taip_model, aes(x = Leaf, y = TAIP)) + geom_line()
-ac_taip_lines <- ac_taip_seg$psi[, 2]
-ac_taip_mean_break <- ac_taip_plot + geom_vline(xintercept = ac_taip_lines, linetype = "dashed")
-ac_taip_mean_break
-ggsave("ac_taip_break.png")
-```
-
-We tested leaf number and average tooth area.
-```{r}
-skim(acerifolia, average.tooth.area)
-ac_ta_plot <- ggplot(acerifolia, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_ta_plot <- ac_ta_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-ac_ta_plot
-
-```
-
-```{r}
-ac_ta_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.ta)) + geom_line() + theme_classic()
-ac_ta_mean_plot <- ac_ta_mean_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-ac_ta_mean_plot
-```
-
-We ran a LM for leaf number and average tooth area.
-```{r}
-ac_ta_lm <- lm(avg.ta ~ leaf, data = ac_l_means)
-summary(ac_ta_lm)
-```
-
-```{r}
-ac_ta_seg <- segmented(ac_ta_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_ta_seg)
-```
-
-```{r}
-ac_ta_seg$psi
-slope(ac_ta_seg)
-ac_ta_fitted <- fitted(ac_ta_seg)
-ac_ta_model <- data.frame(Leaf = ac_l_means$leaf, TA = ac_ta_fitted)
-ggplot(ac_ta_model, aes(x = Leaf, y = TA)) + geom_line()
-ac_ta_lines <- ac_ta_seg$psi[, 2]
-ac_ta_mean_break <- ac_ta_plot + geom_vline(xintercept = ac_ta_lines, linetype = "dashed")
-ac_ta_mean_break
-ggsave("ac_ta_break.png")
-```
-
-We tested leaf number and teeth:perimeter.
-```{r}
-skim(acerifolia, teeth.perimeter)
-ac_tp_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tp_plot <- ac_tp_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-ac_tp_plot
-
-```
-
-```{r}
-ac_tp_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
-ac_tp_mean_plot <- ac_tp_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-ac_tp_mean_plot
-```
-
-We ran a LM for leaf number and teeth:perimeter.
-```{r}
-ac_tp_lm <- lm(avg.tp ~ leaf, data = ac_l_means)
-summary(ac_tp_lm)
-```
-
-```{r}
-ac_tp_seg <- segmented(ac_tp_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tp_seg)
-```
-
-```{r}
-ac_tp_seg$psi
-slope(ac_tp_seg)
-ac_tp_fitted <- fitted(ac_tp_seg)
-ac_tp_model <- data.frame(Leaf = ac_l_means$leaf, TP = ac_tp_fitted)
-ggplot(ac_tp_model, aes(x = Leaf, y = TP)) + geom_line()
-ac_tp_lines <- ac_tp_seg$psi[, 2]
-ac_tp_mean_break <- ac_tp_plot + geom_vline(xintercept = ac_tp_lines, linetype = "dashed")
-ac_tp_mean_break
-ggsave("ac_tp_break.png")
-```
-
-We tested leaf number and teeth:internal perimeter.
-```{r}
-skim(acerifolia, teeth.int.perimeter)
-ac_tip_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tip_plot <- ac_tip_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-ac_tip_plot
-
-```
-
-```{r}
-ac_tip_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
-ac_tip_mean_plot <- ac_tip_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-ac_tip_mean_plot
-```
-
-We ran a LM for leaf number and teeth:internal perimeter.
-```{r}
-ac_tip_lm <- lm(avg.tip ~ leaf, data = ac_l_means)
-summary(ac_tip_lm)
-```
-
-```{r}
-ac_tip_seg <- segmented(ac_tip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tip_seg)
-```
-
-```{r}
-ac_tip_seg$psi
-slope(ac_tip_seg)
-ac_tip_fitted <- fitted(ac_tip_seg)
-ac_tip_model <- data.frame(Leaf = ac_l_means$leaf, TIP = ac_tip_fitted)
-ggplot(ac_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
-ac_tip_lines <- ac_tip_seg$psi[, 2]
-ac_tip_mean_break <- ac_tip_plot + geom_vline(xintercept = ac_tip_lines, linetype = "dashed")
-ac_tip_mean_break
-ggsave("ac_tip_break.png")
-```
-
-We tested leaf number and perimeter:area.
-```{r}
-skim(acerifolia, perimeter.area)
-ac_pa_plot <- ggplot(acerifolia, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_pa_plot <- ac_pa_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-ac_pa_plot
-
-```
-
-```{r}
-ac_pa_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
-ac_pa_mean_plot <- ac_pa_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-ac_pa_mean_plot
-```
-
-We ran a LM for leaf number and perimeter:area.
-```{r}
-ac_pa_lm <- lm(avg.pa ~ leaf, data = ac_l_means)
-summary(ac_pa_lm)
-```
-
-```{r}
-ac_pa_seg <- segmented(ac_pa_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_pa_seg)
-```
-
-```{r}
-ac_pa_seg$psi
-slope(ac_pa_seg)
-ac_pa_fitted <- fitted(ac_pa_seg)
-ac_pa_model <- data.frame(Leaf = ac_l_means$leaf, PA = ac_pa_fitted)
-ggplot(ac_pa_model, aes(x = Leaf, y = PA)) + geom_line()
-ac_pa_lines <- ac_pa_seg$psi[, 2]
-ac_pa_mean_break <- ac_pa_plot + geom_vline(xintercept = ac_pa_lines, linetype = "dashed")
-ac_pa_mean_break
-ggsave("ac_pa_break.png")
-```
-
-We tested leaf number and perimeter ratio.
-```{r}
-skim(acerifolia, perimeter.ratio)
-ac_pr_plot <- ggplot(acerifolia, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_pr_plot <- ac_pr_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-ac_pr_plot
-
-```
-
-```{r}
-ac_pr_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
-ac_pr_mean_plot <- ac_pr_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-ac_pr_mean_plot
-```
-
-We ran a LM for leaf number and perimeter ratio.
-```{r}
-ac_pr_lm <- lm(avg.pr ~ leaf, data = ac_l_means)
-summary(ac_pr_lm)
-```
-
-```{r}
-ac_pr_seg <- segmented(ac_pr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_pr_seg)
-```
-
-```{r}
-ac_pr_seg$psi
-slope(ac_pr_seg)
-ac_pr_fitted <- fitted(ac_pr_seg)
-ac_pr_model <- data.frame(Leaf = ac_l_means$leaf, PR = ac_pr_fitted)
-ggplot(ac_pr_model, aes(x = Leaf, y = PR)) + geom_line()
-ac_pr_lines <- ac_pr_seg$psi[, 2]
-ac_pr_mean_break <- ac_pr_plot + geom_vline(xintercept = ac_pr_lines, linetype = "dashed")
-ac_pr_mean_break
-ggsave("ac_pr_break.png")
-```
-
-We tested leaf number and compactness.
-```{r}
-skim(acerifolia, compactness)
-ac_cp_plot <- ggplot(acerifolia, aes(x = leaf, y = compactness)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_cp_plot <- ac_cp_plot +labs(x = "Leaf number",
-                               y = "Compactness")
-ac_cp_plot
-
-```
-
-```{r}
-ac_cp_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.cp)) + geom_line() + theme_classic()
-ac_cp_mean_plot <- ac_cp_mean_plot +labs(x = "Leaf number",
-                               y = "Compactness")
-ac_cp_mean_plot
-```
-
-We ran a LM for leaf number and compactness.
-```{r}
-ac_cp_lm <- lm(avg.cp ~ leaf, data = ac_l_means)
-summary(ac_cp_lm)
-```
-
-```{r}
-ac_cp_seg <- segmented(ac_cp_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_cp_seg)
-```
-
-```{r}
-ac_cp_seg$psi
-slope(ac_cp_seg)
-ac_cp_fitted <- fitted(ac_cp_seg)
-ac_cp_model <- data.frame(Leaf = ac_l_means$leaf, CP = ac_cp_fitted)
-ggplot(ac_cp_model, aes(x = Leaf, y = CP)) + geom_line()
-ac_cp_lines <- ac_cp_seg$psi[, 2]
-ac_cp_mean_break <- ac_cp_plot + geom_vline(xintercept = ac_cp_lines, linetype = "dashed")
-ac_cp_mean_break
-ggsave("ac_cp_break.png")
-```
-
-We tested leaf number and shape factor.
-```{r}
-skim(acerifolia, shape.factor)
-ac_sf_plot <- ggplot(acerifolia, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_sf_plot <- ac_sf_plot +labs(x = "Leaf number",
-                               y = "Shape Factor")
-ac_sf_plot
-
-```
-
-```{r}
-ac_sf_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
-ac_sf_mean_plot <- ac_sf_mean_plot +labs(x = "Leaf number",
-                               y = "Shape Factor")
-ac_sf_mean_plot
-```
-
-We ran a LM for leaf number and shape factor.
-```{r}
-ac_sf_lm <- lm(avg.sf ~ leaf, data = ac_l_means)
-summary(ac_sf_lm)
-```
-
-```{r}
-ac_sf_seg <- segmented(ac_sf_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_sf_seg)
-```
-
-```{r}
-ac_sf_seg$psi
-slope(ac_sf_seg)
-ac_sf_fitted <- fitted(ac_sf_seg)
-ac_sf_model <- data.frame(Leaf = ac_l_means$leaf, SF = ac_sf_fitted)
-ggplot(ac_sf_model, aes(x = Leaf, y = SF)) + geom_line()
-ac_sf_lines <- ac_sf_seg$psi[, 2]
-ac_sf_mean_break <- ac_sf_plot + geom_vline(xintercept = ac_sf_lines, linetype = "dashed")
-ac_sf_mean_break
-ggsave("ac_sf_break.png")
-```
-
-We tested leaf number and teeth:blade area.
-```{r}
-skim(acerifolia, teeth.blade.area)
-ac_tba_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tba_plot <- ac_tba_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-ac_tba_plot
-
-```
-
-```{r}
-ac_tba_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
-ac_tba_mean_plot <- ac_tba_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-ac_tba_mean_plot
-```
-
-We ran a LM for leaf number and teeth:blade area.
-```{r}
-ac_tba_lm <- lm(avg.tba ~ leaf, data = ac_l_means)
-summary(ac_tba_lm)
-```
-
-```{r}
-ac_tba_seg <- segmented(ac_tba_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tba_seg)
-```
-
-```{r}
-ac_tba_seg$psi
-slope(ac_tba_seg)
-ac_tba_fitted <- fitted(ac_tba_seg)
-ac_tba_model <- data.frame(Leaf = ac_l_means$leaf, TBA = ac_tba_fitted)
-ggplot(ac_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
-ac_tba_lines <- ac_tba_seg$psi[, 2]
-ac_tba_mean_break <- ac_tba_plot + geom_vline(xintercept = ac_tba_lines, linetype = "dashed")
-ac_tba_mean_break
-ggsave("ac_tba_break.png")
-```
-
-We tested leaf number and tooth area:blade area for 2013.
-```{r}
-ac_taba13_plot <- ggplot(ac_13, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_taba13_plot <- ac_taba13_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba13_plot
-
-```
-
-```{r}
-ac_taba13_mean_plot <- ggplot(ac_taba13_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
-ac_taba13_mean_plot <- ac_taba13_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba13_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:blade area in 2013.
-```{r}
-ac_taba13_lm <- lm(avg.taba ~ leaf, data = ac_taba13_mean)
-summary(ac_taba13_lm)
-```
-
-```{r}
-ac_taba13_seg <- segmented(ac_taba13_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_taba13_seg)
-```
-
-```{r}
-ac_taba13_seg$psi
-slope(ac_taba13_seg)
-ac_taba13_fitted <- fitted(ac_taba13_seg)
-ac_taba13_model <- data.frame(Leaf = ac_taba13_mean$leaf, TABA = ac_taba13_fitted)
-ggplot(ac_taba13_model, aes(x = Leaf, y = TABA)) + geom_line()
-ac_taba13_lines <- ac_taba13_seg$psi[, 2]
-ac_taba13_mean_break <- ac_taba13_plot + geom_vline(xintercept = ac_taba13_lines, linetype = "dashed")
-ac_taba13_mean_break
-ggsave("ac_taba13_break.png")
-```
-
-We tested leaf number and tooth area:blade area for 2015.
-```{r}
-ac_taba15_plot <- ggplot(ac_15, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_taba15_plot <- ac_taba15_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba15_plot
-
-```
-
-```{r}
-ac_taba15_mean_plot <- ggplot(ac_taba15_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
-ac_taba15_mean_plot <- ac_taba15_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba15_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:blade area in 2015.
-```{r}
-ac_taba15_lm <- lm(avg.taba ~ leaf, data = ac_taba15_mean)
-summary(ac_taba15_lm)
-```
-
-```{r}
-ac_taba15_seg <- segmented(ac_taba15_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_taba15_seg)
-```
-
-```{r}
-ac_taba15_seg$psi
-slope(ac_taba15_seg)
-ac_taba15_fitted <- fitted(ac_taba15_seg)
-ac_taba15_model <- data.frame(Leaf = ac_taba15_mean$leaf, TABA = ac_taba15_fitted)
-ggplot(ac_taba15_model, aes(x = Leaf, y = TABA)) + geom_line()
-ac_taba15_lines <- ac_taba15_seg$psi[, 2]
-ac_taba15_mean_break <- ac_taba15_plot + geom_vline(xintercept = ac_taba15_lines, linetype = "dashed")
-ac_taba15_mean_break
-ggsave("ac_taba15_break.png")
-```
-
-We then tested *Vitis aestivalis*.
-
-```{r}
-aestivalis <-filter(Vitis_90_ratios, species == "aestivalis")
-```
-
-We need the means for each leaf character for both years combined.
-```{r}
-av_l_means <-  aestivalis %>% 
-  group_by(leaf) %>% 
-  summarize(avg.tt = mean(total.teeth),
-            avg.fdr = mean(feret.diameter.ratio, na.rm = TRUE),
-            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
-            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
-            avg.ta = mean(average.tooth.area, na.rm = TRUE),
-            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
-            avg.tp = mean(teeth.perimeter),
-            avg.tip = mean(teeth.int.perimeter),
-            avg.pa = mean(perimeter.area),
-            avg.pr = mean(perimeter.ratio),
-            avg.cp = mean(compactness),
-            avg.sf = mean(shape.factor),
-            avg.tba = mean(teeth.blade.area))
-skim(av_l_means)
-```
-
-Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF), so we need to split these variables for each year.
-```{r}
-av_13 <- filter(aestivalis, year == 2013)
-av_15 <- filter(aestivalis, year == 2015)
-```
-
-Total teeth
-```{r}
-av_tt13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
-
-ac_tt15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
-```
-
-Tooth area: blade area
-```{r}
-av_taba13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-
-ac_taba15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-```
-
-Compactness
-```{r}
-av_cp13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.cp = mean(compactness, na.rm = TRUE))
-
-ac_cp15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.cp = mean(compactness, na.rm = TRUE))
-```
-
-Shape factor
-```{r}
-av_sf13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
-
-ac_sf15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
-```
-
-We wanted to test leaf number and feret diameter ratio.
-```{r}
-av_fdr_plot <- ggplot(aestivalis, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_fdr_plot <- av_fdr_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-av_fdr_plot
-```
-
-```{r}
-av_fdr_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
-av_fdr_mean_plot <- av_fdr_mean_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-av_fdr_mean_plot
-```
-
-We ran a LM for leaf number and feret diamter ratio.
-```{r}
-av_fdr_lm <- lm(avg.fdr ~ leaf, data = av_l_means)
-summary(av_fdr_lm)
-```
-
-```{r}
-av_fdr_seg <- segmented(av_fdr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_fdr_seg)
-```
-
-```{r}
-av_fdr_seg$psi
-slope(av_fdr_seg)
-av_fdr_fitted <- fitted(av_fdr_seg)
-av_fdr_model <- data.frame(Leaf = av_l_means$leaf, FDR = av_fdr_fitted)
-ggplot(av_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
-av_fdr_lines <- av_fdr_seg$psi[, 2]
-av_fdr_mean_break <- av_fdr_plot + geom_vline(xintercept = av_fdr_lines, linetype = "dashed")
-av_fdr_mean_break
-ggsave("av_fdr_break.png")
-```
-
-We tested leaf number and tooth area:perimeter.
-```{r}
-av_tap_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tap_plot <- av_tap_plot +labs(x = "Leaf number",
-                               y = "Tooth area: perimeter")
-av_tap_plot
-```
-
-```{r}
-av_tap_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
-av_tap_mean_plot <- av_tap_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Perimeter")
-av_tap_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:perimeter.
-```{r}
-av_tap_lm <- lm(avg.tap ~ leaf, data = av_l_means)
-summary(av_tap_lm)
-```
-
-```{r}
-av_tap_seg <- segmented(av_tap_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(4, 7, 10)))
-summary(av_tap_seg)
-```
-
-```{r}
-av_tap_seg$psi
-slope(av_tap_seg)
-av_tap_fitted <- fitted(av_tap_seg)
-av_tap_model <- data.frame(Leaf = av_l_means$leaf, TAP = av_tap_fitted)
-ggplot(av_tap_model, aes(x = Leaf, y = TAP)) + geom_line()
-av_tap_lines <- av_tap_seg$psi[, 2]
-av_tap_mean_break <- av_tap_plot + geom_vline(xintercept = av_tap_lines, linetype = "dashed")
-av_tap_mean_break
-ggsave("av_tap_break.png")
-```
-
-We tested leaf number and tooth area:internal perimeter.
-```{r}
-av_taip_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_taip_plot <- av_taip_plot +labs(x = "Leaf number",
-                               y = "Tooth area: internal perimeter")
-av_taip_plot
-```
-
-```{r}
-av_taip_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
-av_taip_mean_plot <- av_taip_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Internal Perimeter")
-av_taip_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:internal perimeter.
-```{r}
-av_taip_lm <- lm(avg.taip ~ leaf, data = av_l_means)
-summary(av_taip_lm)
-```
-
-```{r}
-av_taip_seg <- segmented(av_taip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_taip_seg)
-```
-
-```{r}
-av_taip_seg$psi
-slope(av_taip_seg)
-av_taip_fitted <- fitted(av_taip_seg)
-av_taip_model <- data.frame(Leaf = av_l_means$leaf, TAIP = av_taip_fitted)
-ggplot(av_taip_model, aes(x = Leaf, y = TAIP)) + geom_line()
-av_taip_lines <- av_taip_seg$psi[, 2]
-av_taip_mean_break <- av_taip_plot + geom_vline(xintercept = av_taip_lines, linetype = "dashed")
-av_taip_mean_break
-ggsave("av_taip_break.png")
-```
-
-We tested leaf number and average tooth area.
-```{r}
-av_ta_plot <- ggplot(aestivalis, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_ta_plot <- av_ta_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-av_ta_plot
-
-```
-
-```{r}
-av_ta_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.ta)) + geom_line() + theme_classic()
-av_ta_mean_plot <- av_ta_mean_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-av_ta_mean_plot
-```
-
-We ran a LM for leaf number and average tooth area.
-```{r}
-av_ta_lm <- lm(avg.ta ~ leaf, data = av_l_means)
-summary(av_ta_lm)
-```
-
-```{r}
-av_ta_seg <- segmented(av_ta_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_ta_seg)
-```
-
-```{r}
-av_ta_seg$psi
-slope(av_ta_seg)
-av_ta_fitted <- fitted(av_ta_seg)
-av_ta_model <- data.frame(Leaf = av_l_means$leaf, TA = av_ta_fitted)
-ggplot(av_ta_model, aes(x = Leaf, y = TA)) + geom_line()
-av_ta_lines <- av_ta_seg$psi[, 2]
-av_ta_mean_break <- av_ta_plot + geom_vline(xintercept = av_ta_lines, linetype = "dashed")
-av_ta_mean_break
-ggsave("av_ta_break.png")
-```
-
-We tested leaf number and teeth:perimeter.
-```{r}
-av_tp_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tp_plot <- av_tp_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-av_tp_plot
-
-```
-
-```{r}
-av_tp_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
-av_tp_mean_plot <- av_tp_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-av_tp_mean_plot
-```
-
-We ran a LM for leaf number and teeth:perimeter.
-```{r}
-av_tp_lm <- lm(avg.tp ~ leaf, data = av_l_means)
-summary(av_tp_lm)
-```
-
-```{r}
-av_tp_seg <- segmented(av_tp_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_tp_seg)
-```
-
-```{r}
-av_tp_seg$psi
-slope(av_tp_seg)
-av_tp_fitted <- fitted(av_tp_seg)
-av_tp_model <- data.frame(Leaf = av_l_means$leaf, TP = av_tp_fitted)
-ggplot(av_tp_model, aes(x = Leaf, y = TP)) + geom_line()
-av_tp_lines <- av_tp_seg$psi[, 2]
-av_tp_mean_break <- av_tp_plot + geom_vline(xintercept = av_tp_lines, linetype = "dashed")
-av_tp_mean_break
-ggsave("av_tp_break.png")
-```
-
-We tested leaf number and teeth:internal perimeter.
-```{r}
-av_tip_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tip_plot <- av_tip_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-av_tip_plot
-
-```
-
-```{r}
-av_tip_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
-av_tip_mean_plot <- av_tip_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-av_tip_mean_plot
-```
-
-We ran a LM for leaf number and teeth:internal perimeter.
-```{r}
-av_tip_lm <- lm(avg.tip ~ leaf, data = av_l_means)
-summary(av_tip_lm)
-```
-
-```{r}
-av_tip_seg <- segmented(av_tip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_tip_seg)
-```
-
-```{r}
-av_tip_seg$psi
-slope(av_tip_seg)
-av_tip_fitted <- fitted(av_tip_seg)
-av_tip_model <- data.frame(Leaf = av_l_means$leaf, TIP = av_tip_fitted)
-ggplot(av_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
-av_tip_lines <- av_tip_seg$psi[, 2]
-av_tip_mean_break <- av_tip_plot + geom_vline(xintercept = av_tip_lines, linetype = "dashed")
-av_tip_mean_break
-ggsave("av_tip_break.png")
-```
-
-We tested leaf number and perimeter:area.
-```{r}
-av_pa_plot <- ggplot(aestivalis, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_pa_plot <- av_pa_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-av_pa_plot
-
-```
-
-```{r}
-av_pa_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
-av_pa_mean_plot <- av_pa_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-av_pa_mean_plot
-```
-
-We ran a LM for leaf number and perimeter:area.
-```{r}
-av_pa_lm <- lm(avg.pa ~ leaf, data = av_l_means)
-summary(av_pa_lm)
-```
-
-```{r}
-av_pa_seg <- segmented(av_pa_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(2, 7, 10)))
-summary(av_pa_seg)
-```
-
-```{r}
-ac_pa_seg$psi
-slope(ac_pa_seg)
-ac_pa_fitted <- fitted(ac_pa_seg)
-ac_pa_model <- data.frame(Leaf = ac_l_means$leaf, PA = ac_pa_fitted)
-ggplot(ac_pa_model, aes(x = Leaf, y = PA)) + geom_line()
-ac_pa_lines <- ac_pa_seg$psi[, 2]
-ac_pa_mean_break <- ac_pa_plot + geom_vline(xintercept = ac_pa_lines, linetype = "dashed")
-ac_pa_mean_break
-ggsave("ac_pa_break.png")
-```
-
-We tested leaf number and perimeter ratio.
-```{r}
-av_pr_plot <- ggplot(aestivalis, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_pr_plot <- av_pr_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-av_pr_plot
-```
-
-```{r}
-av_pr_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
-av_pr_mean_plot <- av_pr_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-av_pr_mean_plot
-```
-
-We ran a LM for leaf number and perimeter ratio.
-```{r}
-av_pr_lm <- lm(avg.pr ~ leaf, data = av_l_means)
-summary(av_pr_lm)
-```
-
-```{r}
-av_pr_seg <- segmented(av_pr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_pr_seg)
-```
-
-```{r}
-av_pr_seg$psi
-slope(av_pr_seg)
-av_pr_fitted <- fitted(av_pr_seg)
-av_pr_model <- data.frame(Leaf = av_l_means$leaf, PR = av_pr_fitted)
-ggplot(av_pr_model, aes(x = Leaf, y = PR)) + geom_line()
-av_pr_lines <- av_pr_seg$psi[, 2]
-av_pr_mean_break <- av_pr_plot + geom_vline(xintercept = av_pr_lines, linetype = "dashed")
-av_pr_mean_break
-ggsave("av_pr_break.png")
-```
-
-We tested leaf number and teeth:blade area.
-```{r}
-av_tba_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tba_plot <- av_tba_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-av_tba_plot
-
-```
-
-```{r}
-av_tba_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
-av_tba_mean_plot <- av_tba_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-av_tba_mean_plot
-```
-
-We ran a LM for leaf number and teeth:blade area.
-```{r}
-av_tba_lm <- lm(avg.tba ~ leaf, data = av_l_means)
-summary(av_tba_lm)
-```
-
-```{r}
-av_tba_seg <- segmented(av_tba_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(2, 4, 6)))
-summary(av_tba_seg)
-```
-
-```{r}
-av_tba_seg$psi
-slope(av_tba_seg)
-av_tba_fitted <- fitted(av_tba_seg)
-av_tba_model <- data.frame(Leaf = av_l_means$leaf, TBA = av_tba_fitted)
-ggplot(av_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
-av_tba_lines <- av_tba_seg$psi[, 2]
-av_tba_mean_break <- av_tba_plot + geom_vline(xintercept = av_tba_lines, linetype = "dashed")
-av_tba_mean_break
-ggsave("av_tba_break.png")
-```
-
-First, we tested *Vitis amurensis*.
-
-We ran a GLM for leaf number and total teeth.
-```{r}
-amurensis <-filter(R_DiLP_all_species, species == "amurensis")
-```
-
-First, we tested *Vitis riparia*.
-
-We ran a GLM for leaf number and total teeth.
-```{r}
-riparia <-filter(R_DiLP_all_species, species == "riparia")
-```

### 2018-10-26 V_riparia_LM.Rmd
Set up new documents

--- 2018-10-23 Vitis_GLM.Rmd
@@
 ---
-title: "Paper 1 GLM"
+title: "Vitis riparia LM"
 output: html_notebook
@@
 
-First, we tested *Vitis acerifolia*.
+Finally we tested *Vitis riparia*.
 
@@
 ```{r}
-acerifolia <-filter(Vitis_90_ratios, species == "acerifolia")
-ac_l_means <-  acerifolia %>% 
+riparia <-filter(Vitis_90_ratios, species == "riparia")
+r_l_means <-  riparia %>% 
   group_by(leaf) %>% 
@@
 
-However we need to split tooth area:blade area for each year.
+Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF), so we need to split these variables for each year.
 ```{r}
-ac_13 <- filter(acerifolia, year == 2013)
-ac_15 <- filter(acerifolia, year == 2015)
+r_13 <- filter(riparia, year == 2013)
+r_15 <- filter(riparia, year == 2015)
 ```
-
-```{r}
-ac_taba13_mean <-  acerifolia %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-
-ac_taba15_mean <- acerifolia %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-```
-
-We tested leaf number and total teeth.
-```{r}
-skim(acerifolia, total.teeth)
-ac_tt_plot <- ggplot(acerifolia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tt_plot <- ac_tt_plot +labs(x = "Leaf number",
-                               y = "Total teeth")
-ac_tt_plot
-
-```
-
-```{r}
-ac_tt_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
-ac_tt_mean_plot <- ac_tt_mean_plot +labs(x = "Leaf number",
-                               y = "Total teeth")
-ac_tt_mean_plot
-```
-
-We ran a LM for leaf number and total teeth.
-```{r}
-ac_tt_lm <- lm(avg.tt ~ leaf, data = ac_l_means)
-summary(ac_tt_lm)
-```
-
-```{r}
-ac_tt_seg <- segmented(ac_tt_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tt_seg)
-```
-
-```{r}
-ac_tt_seg$psi
-slope(ac_tt_seg)
-ac_tt_fitted <- fitted(ac_tt_seg)
-ac_tt_model <- data.frame(Leaf = ac_l_means$leaf, Teeth = ac_tt_fitted)
-ggplot(ac_tt_model, aes(x = Leaf, y = Teeth)) + geom_line()
-ac_tt_lines <- ac_tt_seg$psi[, 2]
-ac_tt_mean_break <- ac_tt_plot + geom_vline(xintercept = ac_tt_lines, linetype = "dashed")
-ac_tt_mean_break
-ggsave("ac_tt_break.png")
-```
-
-We wanted to test leaf number and feret diameter ratio.
-```{r}
-skim(acerifolia, total.teeth)
-ac_fdr_plot <- ggplot(acerifolia, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_fdr_plot <- ac_fdr_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-ac_fdr_plot
-
-```
-
-```{r}
-ac_fdr_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
-ac_fdr_mean_plot <- ac_fdr_mean_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-ac_fdr_mean_plot
-```
-
-We ran a LM for leaf number and feret diamter ratio.
-```{r}
-ac_fdr_lm <- lm(avg.fdr ~ leaf, data = ac_l_means)
-summary(ac_fdr_lm)
-```
-
-```{r}
-ac_fdr_seg <- segmented(ac_fdr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_fdr_seg)
-```
-
-```{r}
-ac_fdr_seg$psi
-slope(ac_fdr_seg)
-ac_fdr_fitted <- fitted(ac_fdr_seg)
-ac_fdr_model <- data.frame(Leaf = ac_l_means$leaf, FDR = ac_fdr_fitted)
-ggplot(ac_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
-ac_fdr_lines <- ac_fdr_seg$psi[, 2]
-ac_fdr_mean_break <- ac_fdr_plot + geom_vline(xintercept = ac_fdr_lines, linetype = "dashed")
-ac_fdr_mean_break
-ggsave("ac_fdr_break.png")
-```
-
-We tested leaf number and tooth area:perimeter.
-```{r}
-skim(acerifolia, tooth.area.perimeter)
-ac_tap_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tap_plot <- ac_tap_plot +labs(x = "Leaf number",
-                               y = "Tooth area: perimeter")
-ac_tap_plot
-
-```
-
-```{r}
-ac_tap_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
-ac_tap_mean_plot <- ac_tap_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Perimeter")
-ac_tap_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:perimeter.
-```{r}
-ac_tap_lm <- lm(avg.tap ~ leaf, data = ac_l_means)
-summary(ac_tap_lm)
-```
-
-```{r}
-ac_tap_seg <- segmented(ac_tap_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tap_seg)
-```
-
-```{r}
-ac_tap_seg$psi
-slope(ac_tap_seg)
-ac_tap_fitted <- fitted(ac_tap_seg)
-ac_tap_model <- data.frame(Leaf = ac_l_means$leaf, TAP = ac_tap_fitted)
-ggplot(ac_tap_model, aes(x = Leaf, y = TAP)) + geom_line()
-ac_tap_lines <- ac_tap_seg$psi[, 2]
-ac_tap_mean_break <- ac_tap_plot + geom_vline(xintercept = ac_tap_lines, linetype = "dashed")
-ac_tap_mean_break
-ggsave("ac_tap_break.png")
-```
-
-We tested leaf number and tooth area:internal perimeter.
-```{r}
-skim(acerifolia, tooth.area.int.perimeter)
-ac_taip_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_taip_plot <- ac_taip_plot +labs(x = "Leaf number",
-                               y = "Tooth area: internal perimeter")
-ac_taip_plot
-
-```
-
-```{r}
-ac_taip_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
-ac_taip_mean_plot <- ac_taip_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Internal Perimeter")
-ac_taip_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:internal perimeter.
-```{r}
-ac_taip_lm <- lm(avg.taip ~ leaf, data = ac_l_means)
-summary(ac_taip_lm)
-```
-
-```{r}
-ac_taip_seg <- segmented(ac_taip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_taip_seg)
-```
-
-```{r}
-ac_taip_seg$psi
-slope(ac_taip_seg)
-ac_taip_fitted <- fitted(ac_taip_seg)
-ac_taip_model <- data.frame(Leaf = ac_l_means$leaf, TAIP = ac_taip_fitted)
-ggplot(ac_taip_model, aes(x = Leaf, y = TAIP)) + geom_line()
-ac_taip_lines <- ac_taip_seg$psi[, 2]
-ac_taip_mean_break <- ac_taip_plot + geom_vline(xintercept = ac_taip_lines, linetype = "dashed")
-ac_taip_mean_break
-ggsave("ac_taip_break.png")
-```
-
-We tested leaf number and average tooth area.
-```{r}
-skim(acerifolia, average.tooth.area)
-ac_ta_plot <- ggplot(acerifolia, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_ta_plot <- ac_ta_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-ac_ta_plot
-
-```
-
-```{r}
-ac_ta_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.ta)) + geom_line() + theme_classic()
-ac_ta_mean_plot <- ac_ta_mean_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-ac_ta_mean_plot
-```
-
-We ran a LM for leaf number and average tooth area.
-```{r}
-ac_ta_lm <- lm(avg.ta ~ leaf, data = ac_l_means)
-summary(ac_ta_lm)
-```
-
-```{r}
-ac_ta_seg <- segmented(ac_ta_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_ta_seg)
-```
-
-```{r}
-ac_ta_seg$psi
-slope(ac_ta_seg)
-ac_ta_fitted <- fitted(ac_ta_seg)
-ac_ta_model <- data.frame(Leaf = ac_l_means$leaf, TA = ac_ta_fitted)
-ggplot(ac_ta_model, aes(x = Leaf, y = TA)) + geom_line()
-ac_ta_lines <- ac_ta_seg$psi[, 2]
-ac_ta_mean_break <- ac_ta_plot + geom_vline(xintercept = ac_ta_lines, linetype = "dashed")
-ac_ta_mean_break
-ggsave("ac_ta_break.png")
-```
-
-We tested leaf number and teeth:perimeter.
-```{r}
-skim(acerifolia, teeth.perimeter)
-ac_tp_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tp_plot <- ac_tp_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-ac_tp_plot
-
-```
-
-```{r}
-ac_tp_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
-ac_tp_mean_plot <- ac_tp_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-ac_tp_mean_plot
-```
-
-We ran a LM for leaf number and teeth:perimeter.
-```{r}
-ac_tp_lm <- lm(avg.tp ~ leaf, data = ac_l_means)
-summary(ac_tp_lm)
-```
-
-```{r}
-ac_tp_seg <- segmented(ac_tp_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tp_seg)
-```
-
-```{r}
-ac_tp_seg$psi
-slope(ac_tp_seg)
-ac_tp_fitted <- fitted(ac_tp_seg)
-ac_tp_model <- data.frame(Leaf = ac_l_means$leaf, TP = ac_tp_fitted)
-ggplot(ac_tp_model, aes(x = Leaf, y = TP)) + geom_line()
-ac_tp_lines <- ac_tp_seg$psi[, 2]
-ac_tp_mean_break <- ac_tp_plot + geom_vline(xintercept = ac_tp_lines, linetype = "dashed")
-ac_tp_mean_break
-ggsave("ac_tp_break.png")
-```
-
-We tested leaf number and teeth:internal perimeter.
-```{r}
-skim(acerifolia, teeth.int.perimeter)
-ac_tip_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tip_plot <- ac_tip_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-ac_tip_plot
-
-```
-
-```{r}
-ac_tip_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
-ac_tip_mean_plot <- ac_tip_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-ac_tip_mean_plot
-```
-
-We ran a LM for leaf number and teeth:internal perimeter.
-```{r}
-ac_tip_lm <- lm(avg.tip ~ leaf, data = ac_l_means)
-summary(ac_tip_lm)
-```
-
-```{r}
-ac_tip_seg <- segmented(ac_tip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tip_seg)
-```
-
-```{r}
-ac_tip_seg$psi
-slope(ac_tip_seg)
-ac_tip_fitted <- fitted(ac_tip_seg)
-ac_tip_model <- data.frame(Leaf = ac_l_means$leaf, TIP = ac_tip_fitted)
-ggplot(ac_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
-ac_tip_lines <- ac_tip_seg$psi[, 2]
-ac_tip_mean_break <- ac_tip_plot + geom_vline(xintercept = ac_tip_lines, linetype = "dashed")
-ac_tip_mean_break
-ggsave("ac_tip_break.png")
-```
-
-We tested leaf number and perimeter:area.
-```{r}
-skim(acerifolia, perimeter.area)
-ac_pa_plot <- ggplot(acerifolia, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_pa_plot <- ac_pa_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-ac_pa_plot
-
-```
-
-```{r}
-ac_pa_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
-ac_pa_mean_plot <- ac_pa_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-ac_pa_mean_plot
-```
-
-We ran a LM for leaf number and perimeter:area.
-```{r}
-ac_pa_lm <- lm(avg.pa ~ leaf, data = ac_l_means)
-summary(ac_pa_lm)
-```
-
-```{r}
-ac_pa_seg <- segmented(ac_pa_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_pa_seg)
-```
-
-```{r}
-ac_pa_seg$psi
-slope(ac_pa_seg)
-ac_pa_fitted <- fitted(ac_pa_seg)
-ac_pa_model <- data.frame(Leaf = ac_l_means$leaf, PA = ac_pa_fitted)
-ggplot(ac_pa_model, aes(x = Leaf, y = PA)) + geom_line()
-ac_pa_lines <- ac_pa_seg$psi[, 2]
-ac_pa_mean_break <- ac_pa_plot + geom_vline(xintercept = ac_pa_lines, linetype = "dashed")
-ac_pa_mean_break
-ggsave("ac_pa_break.png")
-```
-
-We tested leaf number and perimeter ratio.
-```{r}
-skim(acerifolia, perimeter.ratio)
-ac_pr_plot <- ggplot(acerifolia, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_pr_plot <- ac_pr_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-ac_pr_plot
-
-```
-
-```{r}
-ac_pr_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
-ac_pr_mean_plot <- ac_pr_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-ac_pr_mean_plot
-```
-
-We ran a LM for leaf number and perimeter ratio.
-```{r}
-ac_pr_lm <- lm(avg.pr ~ leaf, data = ac_l_means)
-summary(ac_pr_lm)
-```
-
-```{r}
-ac_pr_seg <- segmented(ac_pr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_pr_seg)
-```
-
-```{r}
-ac_pr_seg$psi
-slope(ac_pr_seg)
-ac_pr_fitted <- fitted(ac_pr_seg)
-ac_pr_model <- data.frame(Leaf = ac_l_means$leaf, PR = ac_pr_fitted)
-ggplot(ac_pr_model, aes(x = Leaf, y = PR)) + geom_line()
-ac_pr_lines <- ac_pr_seg$psi[, 2]
-ac_pr_mean_break <- ac_pr_plot + geom_vline(xintercept = ac_pr_lines, linetype = "dashed")
-ac_pr_mean_break
-ggsave("ac_pr_break.png")
-```
-
-We tested leaf number and compactness.
-```{r}
-skim(acerifolia, compactness)
-ac_cp_plot <- ggplot(acerifolia, aes(x = leaf, y = compactness)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_cp_plot <- ac_cp_plot +labs(x = "Leaf number",
-                               y = "Compactness")
-ac_cp_plot
-
-```
-
-```{r}
-ac_cp_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.cp)) + geom_line() + theme_classic()
-ac_cp_mean_plot <- ac_cp_mean_plot +labs(x = "Leaf number",
-                               y = "Compactness")
-ac_cp_mean_plot
-```
-
-We ran a LM for leaf number and compactness.
-```{r}
-ac_cp_lm <- lm(avg.cp ~ leaf, data = ac_l_means)
-summary(ac_cp_lm)
-```
-
-```{r}
-ac_cp_seg <- segmented(ac_cp_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_cp_seg)
-```
-
-```{r}
-ac_cp_seg$psi
-slope(ac_cp_seg)
-ac_cp_fitted <- fitted(ac_cp_seg)
-ac_cp_model <- data.frame(Leaf = ac_l_means$leaf, CP = ac_cp_fitted)
-ggplot(ac_cp_model, aes(x = Leaf, y = CP)) + geom_line()
-ac_cp_lines <- ac_cp_seg$psi[, 2]
-ac_cp_mean_break <- ac_cp_plot + geom_vline(xintercept = ac_cp_lines, linetype = "dashed")
-ac_cp_mean_break
-ggsave("ac_cp_break.png")
-```
-
-We tested leaf number and shape factor.
-```{r}
-skim(acerifolia, shape.factor)
-ac_sf_plot <- ggplot(acerifolia, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_sf_plot <- ac_sf_plot +labs(x = "Leaf number",
-                               y = "Shape Factor")
-ac_sf_plot
-
-```
-
-```{r}
-ac_sf_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
-ac_sf_mean_plot <- ac_sf_mean_plot +labs(x = "Leaf number",
-                               y = "Shape Factor")
-ac_sf_mean_plot
-```
-
-We ran a LM for leaf number and shape factor.
-```{r}
-ac_sf_lm <- lm(avg.sf ~ leaf, data = ac_l_means)
-summary(ac_sf_lm)
-```
-
-```{r}
-ac_sf_seg <- segmented(ac_sf_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_sf_seg)
-```
-
-```{r}
-ac_sf_seg$psi
-slope(ac_sf_seg)
-ac_sf_fitted <- fitted(ac_sf_seg)
-ac_sf_model <- data.frame(Leaf = ac_l_means$leaf, SF = ac_sf_fitted)
-ggplot(ac_sf_model, aes(x = Leaf, y = SF)) + geom_line()
-ac_sf_lines <- ac_sf_seg$psi[, 2]
-ac_sf_mean_break <- ac_sf_plot + geom_vline(xintercept = ac_sf_lines, linetype = "dashed")
-ac_sf_mean_break
-ggsave("ac_sf_break.png")
-```
-
-We tested leaf number and teeth:blade area.
-```{r}
-skim(acerifolia, teeth.blade.area)
-ac_tba_plot <- ggplot(acerifolia, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_tba_plot <- ac_tba_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-ac_tba_plot
-
-```
-
-```{r}
-ac_tba_mean_plot <- ggplot(ac_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
-ac_tba_mean_plot <- ac_tba_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-ac_tba_mean_plot
-```
-
-We ran a LM for leaf number and teeth:blade area.
-```{r}
-ac_tba_lm <- lm(avg.tba ~ leaf, data = ac_l_means)
-summary(ac_tba_lm)
-```
-
-```{r}
-ac_tba_seg <- segmented(ac_tba_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_tba_seg)
-```
-
-```{r}
-ac_tba_seg$psi
-slope(ac_tba_seg)
-ac_tba_fitted <- fitted(ac_tba_seg)
-ac_tba_model <- data.frame(Leaf = ac_l_means$leaf, TBA = ac_tba_fitted)
-ggplot(ac_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
-ac_tba_lines <- ac_tba_seg$psi[, 2]
-ac_tba_mean_break <- ac_tba_plot + geom_vline(xintercept = ac_tba_lines, linetype = "dashed")
-ac_tba_mean_break
-ggsave("ac_tba_break.png")
-```
-
-We tested leaf number and tooth area:blade area for 2013.
-```{r}
-ac_taba13_plot <- ggplot(ac_13, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_taba13_plot <- ac_taba13_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba13_plot
-
-```
-
-```{r}
-ac_taba13_mean_plot <- ggplot(ac_taba13_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
-ac_taba13_mean_plot <- ac_taba13_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba13_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:blade area in 2013.
-```{r}
-ac_taba13_lm <- lm(avg.taba ~ leaf, data = ac_taba13_mean)
-summary(ac_taba13_lm)
-```
-
-```{r}
-ac_taba13_seg <- segmented(ac_taba13_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_taba13_seg)
-```
-
-```{r}
-ac_taba13_seg$psi
-slope(ac_taba13_seg)
-ac_taba13_fitted <- fitted(ac_taba13_seg)
-ac_taba13_model <- data.frame(Leaf = ac_taba13_mean$leaf, TABA = ac_taba13_fitted)
-ggplot(ac_taba13_model, aes(x = Leaf, y = TABA)) + geom_line()
-ac_taba13_lines <- ac_taba13_seg$psi[, 2]
-ac_taba13_mean_break <- ac_taba13_plot + geom_vline(xintercept = ac_taba13_lines, linetype = "dashed")
-ac_taba13_mean_break
-ggsave("ac_taba13_break.png")
-```
-
-We tested leaf number and tooth area:blade area for 2015.
-```{r}
-ac_taba15_plot <- ggplot(ac_15, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-ac_taba15_plot <- ac_taba15_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba15_plot
-
-```
-
-```{r}
-ac_taba15_mean_plot <- ggplot(ac_taba15_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
-ac_taba15_mean_plot <- ac_taba15_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area:Blade Area")
-ac_taba15_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:blade area in 2015.
-```{r}
-ac_taba15_lm <- lm(avg.taba ~ leaf, data = ac_taba15_mean)
-summary(ac_taba15_lm)
-```
-
-```{r}
-ac_taba15_seg <- segmented(ac_taba15_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(ac_taba15_seg)
-```
-
-```{r}
-ac_taba15_seg$psi
-slope(ac_taba15_seg)
-ac_taba15_fitted <- fitted(ac_taba15_seg)
-ac_taba15_model <- data.frame(Leaf = ac_taba15_mean$leaf, TABA = ac_taba15_fitted)
-ggplot(ac_taba15_model, aes(x = Leaf, y = TABA)) + geom_line()
-ac_taba15_lines <- ac_taba15_seg$psi[, 2]
-ac_taba15_mean_break <- ac_taba15_plot + geom_vline(xintercept = ac_taba15_lines, linetype = "dashed")
-ac_taba15_mean_break
-ggsave("ac_taba15_break.png")
-```
-
-We then tested *Vitis aestivalis*.
-
-```{r}
-aestivalis <-filter(Vitis_90_ratios, species == "aestivalis")
-```
-
-We need the means for each leaf character for both years combined.
-```{r}
-av_l_means <-  aestivalis %>% 
-  group_by(leaf) %>% 
-  summarize(avg.tt = mean(total.teeth),
-            avg.fdr = mean(feret.diameter.ratio, na.rm = TRUE),
-            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
-            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
-            avg.ta = mean(average.tooth.area, na.rm = TRUE),
-            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
-            avg.tp = mean(teeth.perimeter),
-            avg.tip = mean(teeth.int.perimeter),
-            avg.pa = mean(perimeter.area),
-            avg.pr = mean(perimeter.ratio),
-            avg.cp = mean(compactness),
-            avg.sf = mean(shape.factor),
-            avg.tba = mean(teeth.blade.area))
-skim(av_l_means)
-```
-
-Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF), so we need to split these variables for each year.
-```{r}
-av_13 <- filter(aestivalis, year == 2013)
-av_15 <- filter(aestivalis, year == 2015)
-```
-
-Total teeth
-```{r}
-av_tt13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
-
-ac_tt15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
-```
-
-Tooth area: blade area
-```{r}
-av_taba13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-
-ac_taba15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
-```
-
-Compactness
-```{r}
-av_cp13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.cp = mean(compactness, na.rm = TRUE))
-
-ac_cp15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.cp = mean(compactness, na.rm = TRUE))
-```
-
-Shape factor
-```{r}
-av_sf13_mean <-  aestivalis %>% 
-  filter(year == 2013) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
-
-ac_sf15_mean <- aestivalis %>% 
-  filter(year == 2015) %>% 
-  group_by(leaf) %>% 
-  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
-```
-
-We wanted to test leaf number and feret diameter ratio.
-```{r}
-av_fdr_plot <- ggplot(aestivalis, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_fdr_plot <- av_fdr_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-av_fdr_plot
-```
-
-```{r}
-av_fdr_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
-av_fdr_mean_plot <- av_fdr_mean_plot +labs(x = "Leaf number",
-                               y = "Feret Diameter Ratio")
-av_fdr_mean_plot
-```
-
-We ran a LM for leaf number and feret diamter ratio.
-```{r}
-av_fdr_lm <- lm(avg.fdr ~ leaf, data = av_l_means)
-summary(av_fdr_lm)
-```
-
-```{r}
-av_fdr_seg <- segmented(av_fdr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_fdr_seg)
-```
-
-```{r}
-av_fdr_seg$psi
-slope(av_fdr_seg)
-av_fdr_fitted <- fitted(av_fdr_seg)
-av_fdr_model <- data.frame(Leaf = av_l_means$leaf, FDR = av_fdr_fitted)
-ggplot(av_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
-av_fdr_lines <- av_fdr_seg$psi[, 2]
-av_fdr_mean_break <- av_fdr_plot + geom_vline(xintercept = av_fdr_lines, linetype = "dashed")
-av_fdr_mean_break
-ggsave("av_fdr_break.png")
-```
-
-We tested leaf number and tooth area:perimeter.
-```{r}
-av_tap_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tap_plot <- av_tap_plot +labs(x = "Leaf number",
-                               y = "Tooth area: perimeter")
-av_tap_plot
-```
-
-```{r}
-av_tap_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
-av_tap_mean_plot <- av_tap_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Perimeter")
-av_tap_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:perimeter.
-```{r}
-av_tap_lm <- lm(avg.tap ~ leaf, data = av_l_means)
-summary(av_tap_lm)
-```
-
-```{r}
-av_tap_seg <- segmented(av_tap_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(4, 7, 10)))
-summary(av_tap_seg)
-```
-
-```{r}
-av_tap_seg$psi
-slope(av_tap_seg)
-av_tap_fitted <- fitted(av_tap_seg)
-av_tap_model <- data.frame(Leaf = av_l_means$leaf, TAP = av_tap_fitted)
-ggplot(av_tap_model, aes(x = Leaf, y = TAP)) + geom_line()
-av_tap_lines <- av_tap_seg$psi[, 2]
-av_tap_mean_break <- av_tap_plot + geom_vline(xintercept = av_tap_lines, linetype = "dashed")
-av_tap_mean_break
-ggsave("av_tap_break.png")
-```
-
-We tested leaf number and tooth area:internal perimeter.
-```{r}
-av_taip_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_taip_plot <- av_taip_plot +labs(x = "Leaf number",
-                               y = "Tooth area: internal perimeter")
-av_taip_plot
-```
-
-```{r}
-av_taip_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
-av_taip_mean_plot <- av_taip_mean_plot +labs(x = "Leaf number",
-                               y = "Tooth Area: Internal Perimeter")
-av_taip_mean_plot
-```
-
-We ran a LM for leaf number and tooth area:internal perimeter.
-```{r}
-av_taip_lm <- lm(avg.taip ~ leaf, data = av_l_means)
-summary(av_taip_lm)
-```
-
-```{r}
-av_taip_seg <- segmented(av_taip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_taip_seg)
-```
-
-```{r}
-av_taip_seg$psi
-slope(av_taip_seg)
-av_taip_fitted <- fitted(av_taip_seg)
-av_taip_model <- data.frame(Leaf = av_l_means$leaf, TAIP = av_taip_fitted)
-ggplot(av_taip_model, aes(x = Leaf, y = TAIP)) + geom_line()
-av_taip_lines <- av_taip_seg$psi[, 2]
-av_taip_mean_break <- av_taip_plot + geom_vline(xintercept = av_taip_lines, linetype = "dashed")
-av_taip_mean_break
-ggsave("av_taip_break.png")
-```
-
-We tested leaf number and average tooth area.
-```{r}
-av_ta_plot <- ggplot(aestivalis, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_ta_plot <- av_ta_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-av_ta_plot
-
-```
-
-```{r}
-av_ta_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.ta)) + geom_line() + theme_classic()
-av_ta_mean_plot <- av_ta_mean_plot +labs(x = "Leaf number",
-                               y = "Average tooth area")
-av_ta_mean_plot
-```
-
-We ran a LM for leaf number and average tooth area.
-```{r}
-av_ta_lm <- lm(avg.ta ~ leaf, data = av_l_means)
-summary(av_ta_lm)
-```
-
-```{r}
-av_ta_seg <- segmented(av_ta_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_ta_seg)
-```
-
-```{r}
-av_ta_seg$psi
-slope(av_ta_seg)
-av_ta_fitted <- fitted(av_ta_seg)
-av_ta_model <- data.frame(Leaf = av_l_means$leaf, TA = av_ta_fitted)
-ggplot(av_ta_model, aes(x = Leaf, y = TA)) + geom_line()
-av_ta_lines <- av_ta_seg$psi[, 2]
-av_ta_mean_break <- av_ta_plot + geom_vline(xintercept = av_ta_lines, linetype = "dashed")
-av_ta_mean_break
-ggsave("av_ta_break.png")
-```
-
-We tested leaf number and teeth:perimeter.
-```{r}
-av_tp_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tp_plot <- av_tp_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-av_tp_plot
-
-```
-
-```{r}
-av_tp_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
-av_tp_mean_plot <- av_tp_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:perimeter")
-av_tp_mean_plot
-```
-
-We ran a LM for leaf number and teeth:perimeter.
-```{r}
-av_tp_lm <- lm(avg.tp ~ leaf, data = av_l_means)
-summary(av_tp_lm)
-```
-
-```{r}
-av_tp_seg <- segmented(av_tp_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_tp_seg)
-```
-
-```{r}
-av_tp_seg$psi
-slope(av_tp_seg)
-av_tp_fitted <- fitted(av_tp_seg)
-av_tp_model <- data.frame(Leaf = av_l_means$leaf, TP = av_tp_fitted)
-ggplot(av_tp_model, aes(x = Leaf, y = TP)) + geom_line()
-av_tp_lines <- av_tp_seg$psi[, 2]
-av_tp_mean_break <- av_tp_plot + geom_vline(xintercept = av_tp_lines, linetype = "dashed")
-av_tp_mean_break
-ggsave("av_tp_break.png")
-```
-
-We tested leaf number and teeth:internal perimeter.
-```{r}
-av_tip_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tip_plot <- av_tip_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-av_tip_plot
-
-```
-
-```{r}
-av_tip_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
-av_tip_mean_plot <- av_tip_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:internal perimeter")
-av_tip_mean_plot
-```
-
-We ran a LM for leaf number and teeth:internal perimeter.
-```{r}
-av_tip_lm <- lm(avg.tip ~ leaf, data = av_l_means)
-summary(av_tip_lm)
-```
-
-```{r}
-av_tip_seg <- segmented(av_tip_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_tip_seg)
-```
-
-```{r}
-av_tip_seg$psi
-slope(av_tip_seg)
-av_tip_fitted <- fitted(av_tip_seg)
-av_tip_model <- data.frame(Leaf = av_l_means$leaf, TIP = av_tip_fitted)
-ggplot(av_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
-av_tip_lines <- av_tip_seg$psi[, 2]
-av_tip_mean_break <- av_tip_plot + geom_vline(xintercept = av_tip_lines, linetype = "dashed")
-av_tip_mean_break
-ggsave("av_tip_break.png")
-```
-
-We tested leaf number and perimeter:area.
-```{r}
-av_pa_plot <- ggplot(aestivalis, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_pa_plot <- av_pa_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-av_pa_plot
-
-```
-
-```{r}
-av_pa_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
-av_pa_mean_plot <- av_pa_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter:Area")
-av_pa_mean_plot
-```
-
-We ran a LM for leaf number and perimeter:area.
-```{r}
-av_pa_lm <- lm(avg.pa ~ leaf, data = av_l_means)
-summary(av_pa_lm)
-```
-
-```{r}
-av_pa_seg <- segmented(av_pa_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(2, 7, 10)))
-summary(av_pa_seg)
-```
-
-```{r}
-ac_pa_seg$psi
-slope(ac_pa_seg)
-ac_pa_fitted <- fitted(ac_pa_seg)
-ac_pa_model <- data.frame(Leaf = ac_l_means$leaf, PA = ac_pa_fitted)
-ggplot(ac_pa_model, aes(x = Leaf, y = PA)) + geom_line()
-ac_pa_lines <- ac_pa_seg$psi[, 2]
-ac_pa_mean_break <- ac_pa_plot + geom_vline(xintercept = ac_pa_lines, linetype = "dashed")
-ac_pa_mean_break
-ggsave("ac_pa_break.png")
-```
-
-We tested leaf number and perimeter ratio.
-```{r}
-av_pr_plot <- ggplot(aestivalis, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_pr_plot <- av_pr_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-av_pr_plot
-```
-
-```{r}
-av_pr_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
-av_pr_mean_plot <- av_pr_mean_plot +labs(x = "Leaf number",
-                               y = "Perimeter Ratio")
-av_pr_mean_plot
-```
-
-We ran a LM for leaf number and perimeter ratio.
-```{r}
-av_pr_lm <- lm(avg.pr ~ leaf, data = av_l_means)
-summary(av_pr_lm)
-```
-
-```{r}
-av_pr_seg <- segmented(av_pr_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
-summary(av_pr_seg)
-```
-
-```{r}
-av_pr_seg$psi
-slope(av_pr_seg)
-av_pr_fitted <- fitted(av_pr_seg)
-av_pr_model <- data.frame(Leaf = av_l_means$leaf, PR = av_pr_fitted)
-ggplot(av_pr_model, aes(x = Leaf, y = PR)) + geom_line()
-av_pr_lines <- av_pr_seg$psi[, 2]
-av_pr_mean_break <- av_pr_plot + geom_vline(xintercept = av_pr_lines, linetype = "dashed")
-av_pr_mean_break
-ggsave("av_pr_break.png")
-```
-
-We tested leaf number and teeth:blade area.
-```{r}
-av_tba_plot <- ggplot(aestivalis, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
-av_tba_plot <- av_tba_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-av_tba_plot
-
-```
-
-```{r}
-av_tba_mean_plot <- ggplot(av_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
-av_tba_mean_plot <- av_tba_mean_plot +labs(x = "Leaf number",
-                               y = "Teeth:Blade Area")
-av_tba_mean_plot
-```
-
-We ran a LM for leaf number and teeth:blade area.
-```{r}
-av_tba_lm <- lm(avg.tba ~ leaf, data = av_l_means)
-summary(av_tba_lm)
-```
-
-```{r}
-av_tba_seg <- segmented(av_tba_lm,
-                       seg.Z = ~ leaf,
-                       psi = list(leaf = c(2, 4, 6)))
-summary(av_tba_seg)
-```
-
-```{r}
-av_tba_seg$psi
-slope(av_tba_seg)
-av_tba_fitted <- fitted(av_tba_seg)
-av_tba_model <- data.frame(Leaf = av_l_means$leaf, TBA = av_tba_fitted)
-ggplot(av_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
-av_tba_lines <- av_tba_seg$psi[, 2]
-av_tba_mean_break <- av_tba_plot + geom_vline(xintercept = av_tba_lines, linetype = "dashed")
-av_tba_mean_break
-ggsave("av_tba_break.png")
-```
-
-First, we tested *Vitis amurensis*.
-
-We ran a GLM for leaf number and total teeth.
-```{r}
-amurensis <-filter(R_DiLP_all_species, species == "amurensis")
-```
-
-First, we tested *Vitis riparia*.
-
-We ran a GLM for leaf number and total teeth.
-```{r}
-riparia <-filter(R_DiLP_all_species, species == "riparia")
-```

### 2018-10-26 V_amurensis_LM.Rmd
V amurensis setup

@@
 ```
+
+Total teeth
+```{r}
+am_tt13_mean <-  amurensis %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
+
+am_tt15_mean <- amurensis %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
+```
+
+Tooth area:perimeter
+```{r}
+am_tap13_mean <-  amurensis %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tap = mean(tooth.area.perimeter, na.rm = TRUE))
+
+am_tap15_mean <- amurensis %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tap = mean(tooth.area.perimeter, na.rm = TRUE))
+```
+
+Tooth area:internal perimeter
+```{r}
+am_taip13_mean <-  amurensis %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE))
+
+am_taip15_mean <- amurensis %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE))
+```
+
+Average tooth area
+```{r}
+am_ata13_mean <-  amurensis %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.ata = mean(average.tooth.area, na.rm = TRUE))
+
+am_ata15_mean <- amurensis %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.ata = mean(average.tooth.area, na.rm = TRUE))
+```
+
+Total teeth:perimeter
+```{r}
+am_tp13_mean <-  amurensis %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tp = mean(teeth.perimeter, na.rm = TRUE))
+
+am_tp15_mean <- amurensis %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tp = mean(teeth.perimeter, na.rm = TRUE))
+```
+
+Perimeter:area
+```{r}
+am_pa13_mean <-  amurensis %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.pa = mean(perimeter.area, na.rm = TRUE))
+
+am_pa15_mean <- amurensis %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.pa = mean(perimeter.area, na.rm = TRUE))
+```

### 2018-10-29 V_acerifolia_LM.Rmd
V. acerifolia LM complete

@@
          "numeric", "numeric", "numeric"))
+acerifolia <-filter(Vitis_90_ratios, species == "acerifolia")
 ```
@@
 
-We first tested each of the variables for *Vitis acerifolia*. Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).
+We first tested each of the variables for *Vitis acerifolia*.
 
-We next tested each of the variables for *Vitis aestivalis*.
-Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).
-
-We next tested each of the variables for *Vitis amurensis*.
-Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF).
-
-We finally tested each of the variables for *Vitis riparia*.
-Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).
+```{r}
+t.test(total.teeth ~ year, data= acerifolia, var.equal= TRUE)
+t.test(feret.diameter.ratio ~ year, data= acerifolia, var.equal= TRUE)
+t.test(tooth.area.perimeter ~ year, data= acerifolia, var.equal= TRUE)
+t.test(tooth.area.int.perimeter ~ year, data= acerifolia, var.equal= TRUE)
+t.test(average.tooth.area ~ year, data= acerifolia, var.equal= TRUE)
+t.test(tooth.area.blade.area ~ year, data= acerifolia, var.equal= TRUE)
+t.test(teeth.perimeter ~ year, data= acerifolia, var.equal= TRUE)
+t.test(teeth.int.perimeter ~ year, data= acerifolia, var.equal= TRUE)
+t.test(perimeter.area ~ year, data= acerifolia, var.equal= TRUE)
+t.test(perimeter.ratio ~ year, data= acerifolia, var.equal= TRUE)
+t.test(compactness ~ year, data= acerifolia, var.equal= TRUE)
+t.test(shape.factor ~ year, data= acerifolia, var.equal= TRUE)
+t.test(teeth.blade.area ~ year, data= acerifolia, var.equal= TRUE)
+```
+Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).
 
@@
 ```{r}
-acerifolia <-filter(Vitis_90_ratios, species == "acerifolia")
 ac_l_means <-  acerifolia %>% 

### 2018-10-29 V_aestivalis_LM.Rmd
V. aestivalis LM complete

@@
          "numeric", "numeric", "numeric"))
+aestivalis <-filter(Vitis_90_ratios, species == "aestivalis")
 ```
@@
 
-We first tested each of the variables for *Vitis acerifolia*. Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).
+We next tested each of the variables for *Vitis aestivalis*.
 
-We next tested each of the variables for *Vitis aestivalis*.
+```{r}
+t.test(total.teeth ~ year, data= aestivalis, var.equal= TRUE)
+t.test(feret.diameter.ratio ~ year, data= aestivalis, var.equal= TRUE)
+t.test(tooth.area.perimeter ~ year, data= aestivalis, var.equal= TRUE)
+t.test(tooth.area.int.perimeter ~ year, data= aestivalis, var.equal= TRUE)
+t.test(average.tooth.area ~ year, data= aestivalis, var.equal= TRUE)
+t.test(tooth.area.blade.area ~ year, data= aestivalis, var.equal= TRUE)
+t.test(teeth.perimeter ~ year, data= aestivalis, var.equal= TRUE)
+t.test(teeth.int.perimeter ~ year, data= aestivalis, var.equal= TRUE)
+t.test(perimeter.area ~ year, data= aestivalis, var.equal= TRUE)
+t.test(perimeter.ratio ~ year, data= aestivalis, var.equal= TRUE)
+t.test(compactness ~ year, data= aestivalis, var.equal= TRUE)
+t.test(shape.factor ~ year, data= aestivalis, var.equal= TRUE)
+t.test(teeth.blade.area ~ year, data= aestivalis, var.equal= TRUE)
+```
 Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).
-
-We next tested each of the variables for *Vitis amurensis*.
-Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF).
-
-We finally tested each of the variables for *Vitis riparia*.
-Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).
 
@@
 We then tested *Vitis aestivalis*.
-
-```{r}
-aestivalis <-filter(Vitis_90_ratios, species == "aestivalis")
-```
 

### 2018-10-29 V_amurensis_LM.Rmd
V. amurensis prep

@@
          "numeric", "numeric", "numeric"))
+amurensis <-filter(Vitis_90_ratios, species == "amurensis")
 ```
@@
 
-We first tested each of the variables for *Vitis acerifolia*. Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).
+We next tested each of the variables for *Vitis amurensis*.
 
-We next tested each of the variables for *Vitis aestivalis*.
-Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).
-
-We next tested each of the variables for *Vitis amurensis*.
-Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF).
-
-We finally tested each of the variables for *Vitis riparia*.
-Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).
+```{r}
+t.test(total.teeth ~ year, data= amurensis, var.equal= TRUE)
+t.test(feret.diameter.ratio ~ year, data= amurensis, var.equal= TRUE)
+t.test(tooth.area.perimeter ~ year, data= amurensis, var.equal= TRUE)
+t.test(tooth.area.int.perimeter ~ year, data= amurensis, var.equal= TRUE)
+t.test(average.tooth.area ~ year, data= amurensis, var.equal= TRUE)
+t.test(tooth.area.blade.area ~ year, data= amurensis, var.equal= TRUE)
+t.test(teeth.perimeter ~ year, data= amurensis, var.equal= TRUE)
+t.test(teeth.int.perimeter ~ year, data= amurensis, var.equal= TRUE)
+t.test(perimeter.area ~ year, data= amurensis, var.equal= TRUE)
+t.test(perimeter.ratio ~ year, data= amurensis, var.equal= TRUE)
+t.test(compactness ~ year, data= amurensis, var.equal= TRUE)
+t.test(shape.factor ~ year, data= amurensis, var.equal= TRUE)
+t.test(teeth.blade.area ~ year, data= amurensis, var.equal= TRUE)
+```
+Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), perimeter to ratio (PR), and shape factor (SF).
 
@@
 ```{r}
-amurensis <-filter(Vitis_90_ratios, species == "amurensis")
 am_l_means <-  amurensis %>% 
   group_by(leaf) %>% 
-  summarize(avg.tt = mean(total.teeth),
-            avg.fdr = mean(feret.diameter.ratio),
+  summarize(avg.tt = mean(total.teeth, na.rm = TRUE),
+            avg.fdr = mean(feret.diameter.ratio, na.rm = TRUE),
             avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
@@
             avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
-            avg.tp = mean(teeth.perimeter),
-            avg.tip = mean(teeth.int.perimeter),
-            avg.pa = mean(perimeter.area),
-            avg.pr = mean(perimeter.ratio),
-            avg.cp = mean(compactness),
-            avg.sf = mean(shape.factor),
-            avg.tba = mean(teeth.blade.area))
+            avg.tp = mean(teeth.perimeter, na.rm = TRUE),
+            avg.tip = mean(teeth.int.perimeter, na.rm = TRUE),
+            avg.pa = mean(perimeter.area, na.rm = TRUE),
+            avg.pr = mean(perimeter.ratio, na.rm = TRUE),
+            avg.cp = mean(compactness, na.rm = TRUE),
+            avg.sf = mean(shape.factor, na.rm = TRUE),
+            avg.tba = mean(teeth.blade.area, na.rm = TRUE))
 ```
 
-Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF), so we need to split these variables for each year.
+Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), perimeter to ratio (PR), and shape factor (SF), so we need to split these variables for each year.
 ```{r}
@@
 
-Total teeth:perimeter
+Perimeter ratio
 ```{r}
-am_tp13_mean <-  amurensis %>% 
+am_pr13_mean <- amurensis %>% 
   filter(year == 2013) %>% 
   group_by(leaf) %>% 
-  summarize(avg.tp = mean(teeth.perimeter, na.rm = TRUE))
+  summarize(avg.pr = mean(perimeter.ratio, na.rm = TRUE))
 
-am_tp15_mean <- amurensis %>% 
+am_pr15_mean <- amurensis %>% 
   filter(year == 2015) %>% 
   group_by(leaf) %>% 
-  summarize(avg.tp = mean(teeth.perimeter, na.rm = TRUE))
+  summarize(avg.pr = mean(perimeter.ratio, na.rm = TRUE))
 ```
 
-Perimeter:area
+Shape factor
 ```{r}
-am_pa13_mean <-  amurensis %>% 
+am_sf13_mean <- amurensis %>% 
   filter(year == 2013) %>% 
   group_by(leaf) %>% 
-  summarize(avg.pa = mean(perimeter.area, na.rm = TRUE))
+  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
 
-am_pa15_mean <- amurensis %>% 
+am_sf15_mean <- amurensis %>% 
   filter(year == 2015) %>% 
   group_by(leaf) %>% 
-  summarize(avg.pa = mean(perimeter.area, na.rm = TRUE))
+  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
 ```
+
+We tested leaf number and feret diameter ratio.
+```{r}
+am_fdr_plot <- ggplot(amurensis, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_fdr_plot <- am_fdr_plot +labs(x = "Leaf number",
+                               y = "FDR")
+am_fdr_plot
+
+```
+
+```{r}
+am_fdr_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
+am_fdr_mean_plot <- am_fdr_mean_plot +labs(x = "Leaf number",
+                               y = "FDR")
+am_fdr_mean_plot
+```
+
+We ran a LM for leaf number and feret diameter ratio.
+```{r}
+am_fdr_lm <- lm(avg.fdr ~ leaf, data = am_l_means)
+summary(am_fdr_lm)
+```
+
+```{r}
+am_fdr_seg <- segmented(am_fdr_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(4, 7, 11)))
+summary(am_fdr_seg)
+```
+
+```{r}
+am_fdr_seg$psi
+slope(am_fdr_seg)
+am_fdr_fitted <- fitted(am_fdr_seg)
+am_fdr_model <- data.frame(Leaf = am_l_means$leaf, FDR = am_fdr_fitted)
+ggplot(am_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
+am_fdr_lines <- am_fdr_seg$psi[, 2]
+am_fdr_mean_break <- am_fdr_plot + geom_vline(xintercept = am_fdr_lines, linetype = "dashed")
+am_fdr_mean_break
+ggsave("am_fdr_break.png")
+```

### 2018-10-29 V_riparia_LM.Rmd
V. riparia prep

@@
          "numeric", "numeric", "numeric"))
+riparia <-filter(Vitis_90_ratios, species == "riparia")
 ```
@@
 
-We first tested each of the variables for *Vitis acerifolia*. Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).
-
-We next tested each of the variables for *Vitis aestivalis*.
-Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).
-
-We next tested each of the variables for *Vitis amurensis*.
-Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF).
-
-We finally tested each of the variables for *Vitis riparia*.
+We tested each of the variables for *Vitis riparia*. 
+```{r}
+t.test(total.teeth ~ year, data= riparia, var.equal= TRUE)
+t.test(feret.diameter.ratio ~ year, data= riparia, var.equal= TRUE)
+t.test(tooth.area.perimeter ~ year, data= riparia, var.equal= TRUE)
+t.test(tooth.area.int.perimeter ~ year, data= riparia, var.equal= TRUE)
+t.test(average.tooth.area ~ year, data= riparia, var.equal= TRUE)
+t.test(tooth.area.blade.area ~ year, data= riparia, var.equal= TRUE)
+t.test(teeth.perimeter ~ year, data= riparia, var.equal= TRUE)
+t.test(teeth.int.perimeter ~ year, data= riparia, var.equal= TRUE)
+t.test(perimeter.area ~ year, data= riparia, var.equal= TRUE)
+t.test(perimeter.ratio ~ year, data= riparia, var.equal= TRUE)
+t.test(compactness ~ year, data= riparia, var.equal= TRUE)
+t.test(shape.factor ~ year, data= riparia, var.equal= TRUE)
+t.test(teeth.blade.area ~ year, data= riparia, var.equal= TRUE)
+```
 Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).

### 2018-10-30 V_amurensis_LM.Rmd
V. amurensis update

@@
                        seg.Z = ~ leaf,
-                       psi = list(leaf = c(4, 7, 11)))
+                       psi = list(leaf = c(6, 8, 11)))
 summary(am_fdr_seg)
@@
 ```
+
+We tested leaf number and tooth area:blade area.
+```{r}
+am_taba_plot <- ggplot(amurensis, aes(x = leaf, y = tooth.area.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_taba_plot <- am_taba_plot +labs(x = "Leaf number",
+                               y = "Tooth area:blade area")
+am_taba_plot
+
+```
+
+```{r}
+am_taba_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
+am_taba_mean_plot <- am_taba_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:blade area")
+am_taba_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:blade area.
+```{r}
+am_taba_lm <- lm(avg.taba ~ leaf, data = am_l_means)
+summary(am_taba_lm)
+```
+
+```{r}
+am_taba_seg <- segmented(am_taba_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_taba_seg)
+```
+
+```{r}
+am_taba_seg$psi
+slope(am_taba_seg)
+am_taba_fitted <- fitted(am_taba_seg)
+am_taba_model <- data.frame(Leaf = am_l_means$leaf, TABA = am_taba_fitted)
+ggplot(am_taba_model, aes(x = Leaf, y = TABA)) + geom_line()
+am_taba_lines <- am_taba_seg$psi[, 2]
+am_taba_mean_break <- am_taba_plot + geom_vline(xintercept = am_taba_lines, linetype = "dashed")
+am_taba_mean_break
+ggsave("am_taba_break.png")
+```
+
+We tested leaf number and total teeth:perimeter.
+```{r}
+am_tp_plot <- ggplot(amurensis, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_tp_plot <- am_tp_plot +labs(x = "Leaf number",
+                               y = "Teeth:perimeter")
+am_tp_plot
+
+```
+
+```{r}
+am_tp_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
+am_tp_mean_plot <- am_tp_mean_plot +labs(x = "Leaf number",
+                               y = "Teeth.perimeter")
+am_tp_mean_plot
+```
+
+We ran a LM for leaf number and teeth:perimeter.
+```{r}
+am_tp_lm <- lm(avg.tp ~ leaf, data = am_l_means)
+summary(am_tp_lm)
+```
+
+```{r}
+am_tp_seg <- segmented(am_tp_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_tp_seg)
+```
+
+```{r}
+am_tp_seg$psi
+slope(am_tp_seg)
+am_tp_fitted <- fitted(am_tp_seg)
+am_tp_model <- data.frame(Leaf = am_l_means$leaf, TP = am_tp_fitted)
+ggplot(am_tp_model, aes(x = Leaf, y = TP)) + geom_line()
+am_tp_lines <- am_tp_seg$psi[, 2]
+am_tp_mean_break <- am_tp_plot + geom_vline(xintercept = am_tp_lines, linetype = "dashed")
+am_tp_mean_break
+ggsave("am_tp_break.png")
+```
+
+We tested leaf number and total teeth:internal perimeter.
+```{r}
+am_tip_plot <- ggplot(amurensis, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_tip_plot <- am_tip_plot +labs(x = "Leaf number",
+                               y = "Teeth:internal perimeter")
+am_tip_plot
+```
+
+```{r}
+am_tip_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
+am_tip_mean_plot <- am_tip_mean_plot +labs(x = "Leaf number",
+                               y = "Teeth.internal perimeter")
+am_tip_mean_plot
+```
+
+We ran a LM for leaf number and teeth:internal perimeter.
+```{r}
+am_tip_lm <- lm(avg.tip ~ leaf, data = am_l_means)
+summary(am_tip_lm)
+```
+
+```{r}
+am_tip_seg <- segmented(am_tip_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_tip_seg)
+```
+
+```{r}
+am_tip_seg$psi
+slope(am_tip_seg)
+am_tip_fitted <- fitted(am_tip_seg)
+am_tip_model <- data.frame(Leaf = am_l_means$leaf, TIP = am_tip_fitted)
+ggplot(am_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
+am_tip_lines <- am_tip_seg$psi[, 2]
+am_tip_mean_break <- am_tip_plot + geom_vline(xintercept = am_tip_lines, linetype = "dashed")
+am_tip_mean_break
+ggsave("am_tip_break.png")
+```
+
+We tested leaf number and perimeter:area.
+```{r}
+am_pa_plot <- ggplot(amurensis, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_pa_plot <- am_pa_plot +labs(x = "Leaf number",
+                               y = "Perimeter:area")
+am_pa_plot
+```
+
+```{r}
+am_pa_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
+am_pa_mean_plot <- am_pa_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter:area")
+am_pa_mean_plot
+```
+
+We ran a LM for leaf number and perimeter:area.
+```{r}
+am_pa_lm <- lm(avg.pa ~ leaf, data = am_l_means)
+summary(am_pa_lm)
+```
+
+```{r}
+am_pa_seg <- segmented(am_pa_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_pa_seg)
+```
+
+```{r}
+am_pa_seg$psi
+slope(am_pa_seg)
+am_pa_fitted <- fitted(am_pa_seg)
+am_pa_model <- data.frame(Leaf = am_l_means$leaf, PA = am_pa_fitted)
+ggplot(am_pa_model, aes(x = Leaf, y = PA)) + geom_line()
+am_pa_lines <- am_pa_seg$psi[, 2]
+am_pa_mean_break <- am_pa_plot + geom_vline(xintercept = am_pa_lines, linetype = "dashed")
+am_pa_mean_break
+ggsave("am_pa_break.png")
+```

### 2018-11-02 V_amurensis_LM.Rmd
V. amurensis complete

@@
                        seg.Z = ~ leaf,
-                       psi = list(leaf = c(6, 8, 11)))
+                       psi = list(leaf = c(3, 7, 10)))
 summary(am_fdr_seg)
@@
 ```
+
+We tested leaf number and compactness.
+```{r}
+am_cp_plot <- ggplot(amurensis, aes(x = leaf, y = compactness)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_cp_plot <- am_cp_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+am_cp_plot
+```
+
+```{r}
+am_cp_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.cp)) + geom_line() + theme_classic()
+am_cp_mean_plot <- am_cp_mean_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+am_cp_mean_plot
+```
+
+We ran a LM for leaf number and compactness.
+```{r}
+am_cp_lm <- lm(avg.cp ~ leaf, data = am_l_means)
+summary(am_cp_lm)
+```
+
+```{r}
+am_cp_seg <- segmented(am_cp_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_cp_seg)
+```
+
+```{r}
+am_cp_seg$psi
+slope(am_cp_seg)
+am_cp_fitted <- fitted(am_cp_seg)
+am_cp_model <- data.frame(Leaf = am_l_means$leaf, CP = am_cp_fitted)
+ggplot(am_cp_model, aes(x = Leaf, y = CP)) + geom_line()
+am_cp_lines <- am_cp_seg$psi[, 2]
+am_cp_mean_break <- am_cp_plot + geom_vline(xintercept = am_cp_lines, linetype = "dashed")
+am_cp_mean_break
+ggsave("am_cp_break.png")
+```
+
+We tested leaf number and total teeth:blade are.
+```{r}
+am_tba_plot <- ggplot(amurensis, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_tba_plot <- am_tba_plot +labs(x = "Leaf number",
+                               y = "Total Teeth:Blade Area")
+am_tba_plot
+```
+
+```{r}
+am_tba_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
+am_tba_mean_plot <- am_tba_mean_plot +labs(x = "Leaf number",
+                               y = "Total Teeth:Blade Area")
+am_tba_mean_plot
+```
+
+We ran a LM for leaf number and total teeth:blade area.
+```{r}
+am_tba_lm <- lm(avg.tba ~ leaf, data = am_l_means)
+summary(am_tba_lm)
+```
+
+```{r}
+am_tba_seg <- segmented(am_tba_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_tba_seg)
+```
+
+```{r}
+am_tba_seg$psi
+slope(am_tba_seg)
+am_tba_fitted <- fitted(am_tba_seg)
+am_tba_model <- data.frame(Leaf = am_l_means$leaf, TBA = am_tba_fitted)
+ggplot(am_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
+am_tba_lines <- am_tba_seg$psi[, 2]
+am_tba_mean_break <- am_tba_plot + geom_vline(xintercept = am_tba_lines, linetype = "dashed")
+am_tba_mean_break
+ggsave("am_tba_break.png")
+```
+
+We tested leaf number and total teeth for 2013.
+```{r}
+am_tt13_plot <- ggplot(am_13, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_tt13_plot <- am_tt13_plot +labs(x = "Leaf number",
+                               y = "Total Teeth")
+am_tt13_plot
+```
+
+```{r}
+am_tt13_mean_plot <- ggplot(am_tt13_mean, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
+am_tt13_mean_plot <- am_tt13_mean_plot +labs(x = "Leaf number",
+                               y = "Total Teeth")
+am_tt13_mean_plot
+```
+
+We ran a LM for leaf number and total teeth in 2013.
+```{r}
+am_tt13_lm <- lm(avg.tt ~ leaf, data = am_tt13_mean)
+summary(am_tt13_lm)
+```
+
+```{r}
+am_tt13_seg <- segmented(am_tt13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_tt13_seg)
+```
+
+```{r}
+am_tt13_seg$psi
+slope(am_tt13_seg)
+am_tt13_fitted <- fitted(am_tt13_seg)
+am_tt13_model <- data.frame(Leaf = am_tt13_mean$leaf, TT = am_tt13_fitted)
+ggplot(am_tt13_model, aes(x = Leaf, y = TT)) + geom_line()
+am_tt13_lines <- am_tt13_seg$psi[, 2]
+am_tt13_mean_break <- am_tt13_plot + geom_vline(xintercept = am_tt13_lines, linetype = "dashed")
+am_tt13_mean_break
+ggsave("am_tt13_break.png")
+```
+
+We tested leaf number and total teeth for 2015.
+```{r}
+am_tt15_plot <- ggplot(am_15, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_tt15_plot <- am_tt15_plot +labs(x = "Leaf number",
+                               y = "Total Teeth")
+am_tt15_plot
+```
+
+```{r}
+am_tt15_mean_plot <- ggplot(am_tt15_mean, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
+am_tt15_mean_plot <- am_tt15_mean_plot +labs(x = "Leaf number",
+                               y = "Total Teeth")
+am_tt15_mean_plot
+```
+
+We ran a LM for leaf number and total teeth in 2015.
+```{r}
+am_tt15_lm <- lm(avg.tt ~ leaf, data = am_tt15_mean)
+summary(am_tt15_lm)
+```
+
+```{r}
+am_tt15_seg <- segmented(am_tt15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_tt15_seg)
+```
+
+```{r}
+am_tt15_seg$psi
+slope(am_tt15_seg)
+am_tt15_fitted <- fitted(am_tt15_seg)
+am_tt15_model <- data.frame(Leaf = am_tt15_mean$leaf, TT = am_tt15_fitted)
+ggplot(am_tt15_model, aes(x = Leaf, y = TT)) + geom_line()
+am_tt15_lines <- am_tt15_seg$psi[, 2]
+am_tt15_mean_break <- am_tt15_plot + geom_vline(xintercept = am_tt15_lines, linetype = "dashed")
+am_tt15_mean_break
+ggsave("am_tt15_break.png")
+```
+
+We tested leaf number and tooth area:perimeter for 2013.
+```{r}
+am_tap13_plot <- ggplot(am_13, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_tap13_plot <- am_tap13_plot +labs(x = "Leaf number",
+                               y = "Tooth area: perimeter")
+am_tap13_plot
+```
+
+```{r}
+am_tap13_mean_plot <- ggplot(am_tap13_mean, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
+am_tap13_mean_plot <- am_tap13_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area: perimeter")
+am_tap13_mean_plot
+```
+
+We ran a LM for leaf number and toth area perimeter in 2013.
+```{r}
+am_tap13_lm <- lm(avg.tap ~ leaf, data = am_tap13_mean)
+summary(am_tap13_lm)
+```
+
+```{r}
+am_tap13_seg <- segmented(am_tap13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_tap13_seg)
+```
+
+```{r}
+am_tap13_seg$psi
+slope(am_tap13_seg)
+am_tap13_fitted <- fitted(am_tap13_seg)
+am_tap13_model <- data.frame(Leaf = am_tap13_mean$leaf, TAP = am_tap13_fitted)
+ggplot(am_tap13_model, aes(x = Leaf, y = TAP)) + geom_line()
+am_tap13_lines <- am_tap13_seg$psi[, 2]
+am_tap13_mean_break <- am_tap13_plot + geom_vline(xintercept = am_tap13_lines, linetype = "dashed")
+am_tap13_mean_break
+ggsave("am_tap13_break.png")
+```
+
+We tested leaf number and tooth area:perimeter for 2015.
+```{r}
+am_tap15_plot <- ggplot(am_15, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_tap15_plot <- am_tap15_plot +labs(x = "Leaf number",
+                               y = "Tooth area:perimeter")
+am_tap15_plot
+```
+
+```{r}
+am_tap15_mean_plot <- ggplot(am_tap15_mean, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
+am_tap15_mean_plot <- am_tap15_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:perimeter")
+am_tap15_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:perimeter in 2015.
+```{r}
+am_tap15_lm <- lm(avg.tap ~ leaf, data = am_tap15_mean)
+summary(am_tap15_lm)
+```
+
+```{r}
+am_tap15_seg <- segmented(am_tap15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_tap15_seg)
+```
+
+```{r}
+am_tap15_seg$psi
+slope(am_tap15_seg)
+am_tap15_fitted <- fitted(am_tap15_seg)
+am_tap15_model <- data.frame(Leaf = am_tap15_mean$leaf, TAP = am_tap15_fitted)
+ggplot(am_tap15_model, aes(x = Leaf, y = TAP)) + geom_line()
+am_tap15_lines <- am_tap15_seg$psi[, 2]
+am_tap15_mean_break <- am_tap15_plot + geom_vline(xintercept = am_tap15_lines, linetype = "dashed")
+am_tap15_mean_break
+ggsave("am_tap15_break.png")
+```
+
+We tested leaf number and tooth area:internal perimeter for 2013.
+```{r}
+am_taip13_plot <- ggplot(am_13, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_taip13_plot <- am_taip13_plot +labs(x = "Leaf number",
+                               y = "Tooth area: internal perimeter")
+am_taip13_plot
+```
+
+```{r}
+am_taip13_mean_plot <- ggplot(am_taip13_mean, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
+am_taip13_mean_plot <- am_taip13_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:internal perimeter")
+am_taip13_mean_plot
+```
+
+We ran a LM for leaf number and toth area: internal perimeter in 2013.
+```{r}
+am_taip13_lm <- lm(avg.taip ~ leaf, data = am_taip13_mean)
+summary(am_taip13_lm)
+```
+
+```{r}
+am_taip13_seg <- segmented(am_taip13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_taip13_seg)
+```
+
+```{r}
+am_taip13_seg$psi
+slope(am_taip13_seg)
+am_taip13_fitted <- fitted(am_taip13_seg)
+am_taip13_model <- data.frame(Leaf = am_taip13_mean$leaf, TAIP = am_taip13_fitted)
+ggplot(am_taip13_model, aes(x = Leaf, y = TAIP)) + geom_line()
+am_taip13_lines <- am_taip13_seg$psi[, 2]
+am_taip13_mean_break <- am_taip13_plot + geom_vline(xintercept = am_taip13_lines, linetype = "dashed")
+am_taip13_mean_break
+ggsave("am_taip13_break.png")
+```
+
+We tested leaf number and tooth area:internal perimeter for 2015.
+```{r}
+am_taip15_plot <- ggplot(am_15, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_taip15_plot <- am_taip15_plot +labs(x = "Leaf number",
+                               y = "Tooth area:internal perimeter")
+am_taip15_plot
+```
+
+```{r}
+am_taip15_mean_plot <- ggplot(am_taip15_mean, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
+am_taip15_mean_plot <- am_taip15_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:internal perimeter")
+am_taip15_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:internal perimeter in 2015.
+```{r}
+am_taip15_lm <- lm(avg.taip ~ leaf, data = am_taip15_mean)
+summary(am_taip15_lm)
+```
+
+```{r}
+am_taip15_seg <- segmented(am_taip15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_taip15_seg)
+```
+
+```{r}
+am_taip15_seg$psi
+slope(am_taip15_seg)
+am_taip15_fitted <- fitted(am_taip15_seg)
+am_taip15_model <- data.frame(Leaf = am_taip15_mean$leaf, TAIP = am_taip15_fitted)
+ggplot(am_taip15_model, aes(x = Leaf, y = TAIP)) + geom_line()
+am_taip15_lines <- am_taip15_seg$psi[, 2]
+am_taip15_mean_break <- am_taip15_plot + geom_vline(xintercept = am_taip15_lines, linetype = "dashed")
+am_taip15_mean_break
+ggsave("am_taip15_break.png")
+```
+
+We tested leaf number and average tooth area for 2013.
+```{r}
+am_ata13_plot <- ggplot(am_13, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_ata13_plot <- am_ata13_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+am_ata13_plot
+```
+
+```{r}
+am_ata13_mean_plot <- ggplot(am_ata13_mean, aes(x = leaf, y = avg.ata)) + geom_line() + theme_classic()
+am_ata13_mean_plot <- am_ata13_mean_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+am_ata13_mean_plot
+```
+
+We ran a LM for leaf number and average tooth area in 2013.
+```{r}
+am_ata13_lm <- lm(avg.ata ~ leaf, data = am_ata13_mean)
+summary(am_ata13_lm)
+```
+
+```{r}
+am_ata13_seg <- segmented(am_ata13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_ata13_seg)
+```
+
+```{r}
+am_ata13_seg$psi
+slope(am_ata13_seg)
+am_ata13_fitted <- fitted(am_ata13_seg)
+am_ata13_model <- data.frame(Leaf = am_ata13_mean$leaf, ATA = am_ata13_fitted)
+ggplot(am_ata13_model, aes(x = Leaf, y = ATA)) + geom_line()
+am_ata13_lines <- am_ata13_seg$psi[, 2]
+am_ata13_mean_break <- am_ata13_plot + geom_vline(xintercept = am_ata13_lines, linetype = "dashed")
+am_ata13_mean_break
+ggsave("am_ata13_break.png")
+```
+
+We tested leaf number and average tooth area for 2015.
+```{r}
+am_ata15_plot <- ggplot(am_15, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_ata15_plot <- am_ata15_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+am_ata15_plot
+```
+
+```{r}
+am_ata15_mean_plot <- ggplot(am_ata15_mean, aes(x = leaf, y = avg.ata)) + geom_line() + theme_classic()
+am_ata15_mean_plot <- am_ata15_mean_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+am_ata15_mean_plot
+```
+
+We ran a LM for leaf number and average tooth area in 2015.
+```{r}
+am_ata15_lm <- lm(avg.ata ~ leaf, data = am_ata15_mean)
+summary(am_ata15_lm)
+```
+
+```{r}
+am_ata15_seg <- segmented(am_ata15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_ata15_seg)
+```
+
+```{r}
+am_ata15_seg$psi
+slope(am_ata15_seg)
+am_ata15_fitted <- fitted(am_ata15_seg)
+am_ata15_model <- data.frame(Leaf = am_ata15_mean$leaf, ATA = am_ata15_fitted)
+ggplot(am_ata15_model, aes(x = Leaf, y = ATA)) + geom_line()
+am_ata15_lines <- am_ata15_seg$psi[, 2]
+am_ata15_mean_break <- am_ata15_plot + geom_vline(xintercept = am_ata15_lines, linetype = "dashed")
+am_ata15_mean_break
+ggsave("am_ata15_break.png")
+```
+
+We tested leaf number and perimeter ratio for 2013.
+```{r}
+am_pr13_plot <- ggplot(am_13, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_pr13_plot <- am_pr13_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+am_pr13_plot
+```
+
+```{r}
+am_pr13_mean_plot <- ggplot(am_pr13_mean, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
+am_pr13_mean_plot <- am_pr13_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+am_pr13_mean_plot
+```
+
+We ran a LM for leaf number and perimeter ratio in 2013.
+```{r}
+am_pr13_lm <- lm(avg.pr ~ leaf, data = am_pr13_mean)
+summary(am_pr13_lm)
+```
+
+```{r}
+am_pr13_seg <- segmented(am_pr13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_pr13_seg)
+```
+
+```{r}
+am_pr13_seg$psi
+slope(am_pr13_seg)
+am_pr13_fitted <- fitted(am_pr13_seg)
+am_pr13_model <- data.frame(Leaf = am_pr13_mean$leaf, PR = am_pr13_fitted)
+ggplot(am_pr13_model, aes(x = Leaf, y = PR)) + geom_line()
+am_pr13_lines <- am_pr13_seg$psi[, 2]
+am_pr13_mean_break <- am_pr13_plot + geom_vline(xintercept = am_pr13_lines, linetype = "dashed")
+am_pr13_mean_break
+ggsave("am_pr13_break.png")
+```
+
+We tested leaf number and perimeter ratio for 2015.
+```{r}
+am_pr15_plot <- ggplot(am_15, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_pr15_plot <- am_pr15_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+am_pr15_plot
+```
+
+```{r}
+am_pr15_mean_plot <- ggplot(am_pr15_mean, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
+am_pr15_mean_plot <- am_pr15_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+am_pr15_mean_plot
+```
+
+We ran a LM for leaf number and average tooth area in 2015.
+```{r}
+am_pr15_lm <- lm(avg.pr ~ leaf, data = am_pr15_mean)
+summary(am_pr15_lm)
+```
+
+```{r}
+am_pr15_seg <- segmented(am_pr15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_pr15_seg)
+```
+
+```{r}
+am_pr15_seg$psi
+slope(am_pr15_seg)
+am_pr15_fitted <- fitted(am_pr15_seg)
+am_pr15_model <- data.frame(Leaf = am_pr15_mean$leaf, PR = am_pr15_fitted)
+ggplot(am_pr15_model, aes(x = Leaf, y = PR)) + geom_line()
+am_pr15_lines <- am_pr15_seg$psi[, 2]
+am_pr15_mean_break <- am_pr15_plot + geom_vline(xintercept = am_pr15_lines, linetype = "dashed")
+am_pr15_mean_break
+ggsave("am_pr15_break.png")
+```
+
+We tested leaf number and shape factor for 2013.
+```{r}
+am_sf13_plot <- ggplot(am_13, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_sf13_plot <- am_sf13_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+am_sf13_plot
+```
+
+```{r}
+am_sf13_mean_plot <- ggplot(am_sf13_mean, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
+am_sf13_mean_plot <- am_sf13_mean_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+am_sf13_mean_plot
+```
+
+We ran a LM for leaf number and shape factor in 2013.
+```{r}
+am_sf13_lm <- lm(avg.sf ~ leaf, data = am_sf13_mean)
+summary(am_sf13_lm)
+```
+
+```{r}
+am_sf13_seg <- segmented(am_sf13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_sf13_seg)
+```
+
+```{r}
+am_sf13_seg$psi
+slope(am_sf13_seg)
+am_sf13_fitted <- fitted(am_sf13_seg)
+am_sf13_model <- data.frame(Leaf = am_sf13_mean$leaf, SF = am_sf13_fitted)
+ggplot(am_sf13_model, aes(x = Leaf, y = SF)) + geom_line()
+am_sf13_lines <- am_sf13_seg$psi[, 2]
+am_sf13_mean_break <- am_sf13_plot + geom_vline(xintercept = am_sf13_lines, linetype = "dashed")
+am_sf13_mean_break
+ggsave("am_sf13_break.png")
+```
+
+We tested leaf number and shape factor for 2015.
+```{r}
+am_sf15_plot <- ggplot(am_15, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_sf15_plot <- am_sf15_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+am_sf15_plot
+```
+
+```{r}
+am_sf15_mean_plot <- ggplot(am_sf15_mean, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
+am_sf15_mean_plot <- am_sf15_mean_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+am_sf15_mean_plot
+```
+
+We ran a LM for leaf number and shape factor in 2015.
+```{r}
+am_sf15_lm <- lm(avg.sf ~ leaf, data = am_sf15_mean)
+summary(am_sf15_lm)
+```
+
+```{r}
+am_sf15_seg <- segmented(am_sf15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_sf15_seg)
+```
+
+```{r}
+am_sf15_seg$psi
+slope(am_sf15_seg)
+am_sf15_fitted <- fitted(am_sf15_seg)
+am_sf15_model <- data.frame(Leaf = am_sf15_mean$leaf, SF = am_ata15_fitted)
+ggplot(am_sf15_model, aes(x = Leaf, y = SF)) + geom_line()
+am_sf15_lines <- am_sf15_seg$psi[, 2]
+am_sf15_mean_break <- am_sf15_plot + geom_vline(xintercept = am_sf15_lines, linetype = "dashed")
+am_sf15_mean_break
+ggsave("am_sf15_break.png")
+```

### 2018-11-02 V_riparia_LM.Rmd
V. riparia setup

@@
 ```
+
+Tooth area:internal perimeter
+```{r}
+r_taip13_mean <-  riparia %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE))
+
+r_taip15_mean <- riparia %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE))
+```
+
+Tooth area:blade area
+```{r}
+r_taba13_mean <-  riparia %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
+
+r_taba15_mean <- riparia %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.taba = mean(tooth.area.blade.area, na.rm = TRUE))
+```
+
+Total teeth:perimeter
+```{r}
+r_tp13_mean <-  riparia %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tp = mean(teeth.perimeter, na.rm = TRUE))
+
+r_tp15_mean <- riparia %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tp = mean(teeth.perimeter, na.rm = TRUE))
+```
+
+Total teeth: internal perimeter
+```{r}
+r_tip13_mean <-  riparia %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tip = mean(teeth.int.perimeter, na.rm = TRUE))
+
+r_tip15_mean <- riparia %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.tip = mean(teeth.int.perimeter, na.rm = TRUE))
+```
+
+Perimeter ratio
+```{r}
+r_pr13_mean <-  riparia %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.pr = mean(perimeter.ratio, na.rm = TRUE))
+
+r_pr15_mean <- riparia %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.pr = mean(perimeter.ratio, na.rm = TRUE))
+```
+
+Compactness
+```{r}
+r_cp13_mean <-  riparia %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.cp = mean(compactness, na.rm = TRUE))
+
+r_cp15_mean <- riparia %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.cp = mean(compactness, na.rm = TRUE))
+```
+
+Shape factor
+```{r}
+r_sf13_mean <-  riparia %>% 
+  filter(year == 2013) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
+
+r_sf15_mean <- riparia %>% 
+  filter(year == 2015) %>% 
+  group_by(leaf) %>% 
+  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
+```
+
+We tested leaf number and total teeth.
+```{r}
+r_tt_plot <- ggplot(riparia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+r_tt_plot <- r_tt_plot +labs(x = "Leaf number",
+                               y = "Total teeth")
+r_tt_plot
+```
+
+```{r}
+r_tt_mean_plot <- ggplot(r_l_means, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
+r_tt_mean_plot <- r_tt_mean_plot +labs(x = "Leaf number",
+                               y = "Total teeth")
+r_tt_mean_plot
+```
+
+We ran a LM for leaf number and total teeth.
+```{r}
+r_tt_lm <- lm(avg.tt ~ leaf, data = r_l_means)
+summary(r_tt_lm)
+```
+
+```{r}
+r_tt_seg <- segmented(r_tt_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_tt_seg)
+```
+
+```{r}
+r_tt_seg$psi
+slope(r_tt_seg)
+r_tt_fitted <- fitted(r_tt_seg)
+r_tt_model <- data.frame(Leaf = r_l_means$leaf, TT = r_tt_fitted)
+ggplot(r_tt_model, aes(x = Leaf, y = TT)) + geom_line()
+r_tt_lines <- r_tt_seg$psi[, 2]
+r_tt_mean_break <- r_tt_plot + geom_vline(xintercept = r_tt_lines, linetype = "dashed")
+r_tt_mean_break
+ggsave("r_tt_break.png")
+```
+
+We tested leaf number and feret diameter ratio.
+```{r}
+r_fdr_plot <- ggplot(riparia, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+r_fdr_plot <- r_fdr_plot +labs(x = "Leaf number",
+                               y = "FDR")
+r_fdr_plot
+```
+
+```{r}
+r_fdr_mean_plot <- ggplot(r_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
+r_fdr_mean_plot <- r_fdr_mean_plot +labs(x = "Leaf number",
+                               y = "FDR")
+r_fdr_mean_plot
+```
+
+We ran a LM for leaf number and feret diameter ratio.
+```{r}
+r_fdr_lm <- lm(avg.fdr ~ leaf, data = r_l_means)
+summary(r_fdr_lm)
+```
+
+```{r}
+r_fdr_seg <- segmented(r_fdr_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_fdr_seg)
+```
+
+```{r}
+r_fdr_seg$psi
+slope(r_fdr_seg)
+r_fdr_fitted <- fitted(r_fdr_seg)
+r_fdr_model <- data.frame(Leaf = r_l_means$leaf, FDR = r_fdr_fitted)
+ggplot(r_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
+r_fdr_lines <- r_fdr_seg$psi[, 2]
+r_fdr_mean_break <- r_fdr_plot + geom_vline(xintercept = r_fdr_lines, linetype = "dashed")
+r_fdr_mean_break
+ggsave("r_fdr_break.png")
+```

### 2018-11-06 V_riparia_LM.Rmd
V. riparia lm combined

@@
             avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
-            avg.ta = mean(average.tooth.area, na.rm = TRUE),
+            avg.ata = mean(average.tooth.area, na.rm = TRUE),
             avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
@@
             avg.sf = mean(shape.factor),
-            avg.tba = mean(teeth.blade.area))
+            avg.tba = mean(teeth.blade.area, na.rm = TRUE))
 ```
@@
 Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF), so we need to split these variables for each year.
+
 ```{r}
@@
 ```{r}
-r_tt_plot <- ggplot(riparia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+r_tt_plot <- ggplot(riparia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
 r_tt_plot <- r_tt_plot +labs(x = "Leaf number",
@@
 ```{r}
-r_fdr_plot <- ggplot(riparia, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+r_fdr_plot <- ggplot(riparia, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
 r_fdr_plot <- r_fdr_plot +labs(x = "Leaf number",
@@
 ```
+
+We tested leaf number and tooth area:perimeter.
+```{r}
+r_tap_plot <- ggplot(riparia, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_tap_plot <- r_tap_plot +labs(x = "Leaf number",
+                               y = "Tooth area:perimeter")
+r_tap_plot
+```
+
+```{r}
+r_tap_mean_plot <- ggplot(r_l_means, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
+r_tap_mean_plot <- r_tap_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:perimeter")
+r_tap_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:perimeter.
+```{r}
+r_tap_lm <- lm(avg.tap ~ leaf, data = r_l_means)
+summary(r_tap_lm)
+```
+
+```{r}
+r_tap_seg <- segmented(r_tap_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_tap_seg)
+```
+
+```{r}
+r_tap_seg$psi
+slope(r_tap_seg)
+r_tap_fitted <- fitted(r_tap_seg)
+r_tap_model <- data.frame(Leaf = r_l_means$leaf, TAP = r_tap_fitted)
+ggplot(r_tap_model, aes(x = Leaf, y = TAP)) + geom_line()
+r_tap_lines <- r_tap_seg$psi[, 2]
+r_tap_mean_break <- r_tap_plot + geom_vline(xintercept = r_tap_lines, linetype = "dashed")
+r_tap_mean_break
+ggsave("r_tap_break.png")
+```
+
+We tested leaf number and average tooth area.
+```{r}
+r_ata_plot <- ggplot(riparia, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_ata_plot <- r_ata_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+r_ata_plot
+```
+
+```{r}
+r_ata_mean_plot <- ggplot(r_l_means, aes(x = leaf, y = avg.ata)) + geom_line() + theme_classic()
+r_ata_mean_plot <- r_ata_mean_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+r_ata_mean_plot
+```
+
+We ran a LM for leaf number and average tooth area.
+```{r}
+r_ata_lm <- lm(avg.ata ~ leaf, data = r_l_means)
+summary(r_ata_lm)
+```
+
+```{r}
+r_ata_seg <- segmented(r_ata_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_ata_seg)
+```
+
+```{r}
+r_ata_seg$psi
+slope(r_ata_seg)
+r_ata_fitted <- fitted(r_ata_seg)
+r_ata_model <- data.frame(Leaf = r_l_means$leaf, ATA = r_ata_fitted)
+ggplot(r_ata_model, aes(x = Leaf, y = ATA)) + geom_line()
+r_ata_lines <- r_ata_seg$psi[, 2]
+r_ata_mean_break <- r_ata_plot + geom_vline(xintercept = r_ata_lines, linetype = "dashed")
+r_ata_mean_break
+ggsave("r_ata_break.png")
+```
+
+We tested leaf number and perimeter:area.
+```{r}
+r_pa_plot <- ggplot(riparia, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_pa_plot <- r_pa_plot +labs(x = "Leaf number",
+                               y = "Perimeter:area")
+r_pa_plot
+```
+
+```{r}
+r_pa_mean_plot <- ggplot(r_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
+r_pa_mean_plot <- r_pa_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter:area")
+r_pa_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:perimeter.
+```{r}
+r_pa_lm <- lm(avg.pa ~ leaf, data = r_l_means)
+summary(r_pa_lm)
+```
+
+```{r}
+r_pa_seg <- segmented(r_pa_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_pa_seg)
+```
+
+```{r}
+r_pa_seg$psi
+slope(r_pa_seg)
+r_pa_fitted <- fitted(r_pa_seg)
+r_pa_model <- data.frame(Leaf = r_l_means$leaf, PA = r_pa_fitted)
+ggplot(r_pa_model, aes(x = Leaf, y = PA)) + geom_line()
+r_pa_lines <- r_pa_seg$psi[, 2]
+r_pa_mean_break <- r_pa_plot + geom_vline(xintercept = r_pa_lines, linetype = "dashed")
+r_pa_mean_break
+ggsave("r_pa_break.png")
+```
+
+We tested leaf number and teeth:blade area.
+```{r}
+r_tba_plot <- ggplot(riparia, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_tba_plot <- r_pa_plot +labs(x = "Leaf number",
+                               y = "Teeth blade area")
+r_tba_plot
+```
+
+```{r}
+r_tba_mean_plot <- ggplot(r_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
+r_tba_mean_plot <- r_tba_mean_plot +labs(x = "Leaf number",
+                               y = "Teeth: blade area")
+r_tba_mean_plot
+```
+
+We ran a LM for leaf number and teeth:blade area.
+```{r}
+r_tba_lm <- lm(avg.tba ~ leaf, data = r_l_means)
+summary(r_tba_lm)
+```
+
+```{r}
+r_tba_seg <- segmented(r_tba_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_tba_seg)
+```
+
+```{r}
+r_tba_seg$psi
+slope(r_tba_seg)
+r_tba_fitted <- fitted(r_tba_seg)
+r_tba_model <- data.frame(Leaf = r_l_means$leaf, TBA = r_tba_fitted)
+ggplot(r_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
+r_tba_lines <- r_tba_seg$psi[, 2]
+r_tba_mean_break <- r_tba_plot + geom_vline(xintercept = r_tba_lines, linetype = "dashed")
+r_tba_mean_break
+ggsave("r_tba_break.png")
+```

### 2018-11-07 V_riparia_LM.Rmd
V. riparia TAIP & SF

@@
 ```
+
+We tested leaf number and shape factor for 2013.
+```{r}
+r_taip13_plot <- ggplot(r_13, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_taip13_plot <- r_taip13_plot +labs(x = "Leaf number",
+                               y = "Tooth area:internal perimeter")
+r_taip13_plot
+```
+
+```{r}
+r_taip13_mean_plot <- ggplot(r_taip13_mean, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
+r_taip13_mean_plot <- r_taip13_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:internal perimeter")
+r_taip13_mean_plot
+```
+
+We ran a LM for leaf number and shape factor in 2013.
+```{r}
+r_taip13_lm <- lm(avg.taip ~ leaf, data = r_taip13_mean)
+summary(r_taip13_lm)
+```
+
+```{r}
+r_taip13_seg <- segmented(r_taip13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_taip13_seg)
+```
+
+```{r}
+r_taip13_seg$psi
+slope(r_taip13_seg)
+r_taip13_fitted <- fitted(r_taip13_seg)
+r_taip13_model <- data.frame(Leaf = r_taip13_mean$leaf, TAIP = r_taip13_fitted)
+ggplot(r_taip13_model, aes(x = Leaf, y = TAIP)) + geom_line()
+r_taip13_lines <- r_taip13_seg$psi[, 2]
+r_taip13_mean_break <- r_taip13_plot + geom_vline(xintercept = r_taip13_lines, linetype = "dashed")
+r_taip13_mean_break
+ggsave("r_taip13_break.png")
+```
+
+We tested leaf number and tooth area:internal perimeter for 2015.
+```{r}
+r_taip15_plot <- ggplot(r_15, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_taip15_plot <- r_taip15_plot +labs(x = "Leaf number",
+                               y = "Tooth area: internal perimeter")
+r_taip15_plot
+```
+
+```{r}
+r_taip15_mean_plot <- ggplot(r_taip15_mean, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
+r_taip15_mean_plot <- r_taip15_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:internal perimeter")
+r_taip15_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:internal perimeter in 2015.
+```{r}
+r_taip15_lm <- lm(avg.taip ~ leaf, data = r_taip15_mean)
+summary(r_taip15_lm)
+```
+
+```{r}
+r_taip15_seg <- segmented(r_taip15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_taip15_seg)
+```
+
+```{r}
+r_taip15_seg$psi
+slope(r_taip15_seg)
+r_taip15_fitted <- fitted(r_taip15_seg)
+r_taip15_model <- data.frame(Leaf = r_taip15_mean$leaf, TAIP = r_taip15_fitted)
+ggplot(r_taip15_model, aes(x = Leaf, y = TAIP)) + geom_line()
+r_taip15_lines <- r_taip15_seg$psi[, 2]
+r_taip15_mean_break <- r_taip15_plot + geom_vline(xintercept = r_taip15_lines, linetype = "dashed")
+r_taip15_mean_break
+ggsave("r_taip13_break.png")
+```
+
+We tested leaf number and shape factor for 2013.
+```{r}
+r_sf13_plot <- ggplot(r_13, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_sf13_plot <- r_sf13_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+r_sf13_plot
+```
+
+```{r}
+r_sf13_mean_plot <- ggplot(r_sf13_mean, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
+r_sf13_mean_plot <- r_sf13_mean_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+r_sf13_mean_plot
+```
+
+We ran a LM for leaf number and shape factor in 2013.
+```{r}
+r_sf13_lm <- lm(avg.sf ~ leaf, data = r_sf13_mean)
+summary(r_sf13_lm)
+```
+
+```{r}
+r_sf13_seg <- segmented(r_sf13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_sf13_seg)
+```
+
+```{r}
+r_sf13_seg$psi
+slope(r_sf13_seg)
+r_sf13_fitted <- fitted(r_sf13_seg)
+r_sf13_model <- data.frame(Leaf = r_sf13_mean$leaf, SF = r_sf13_fitted)
+ggplot(r_sf13_model, aes(x = Leaf, y = SF)) + geom_line()
+r_sf13_lines <- r_sf13_seg$psi[, 2]
+r_sf13_mean_break <- r_sf13_plot + geom_vline(xintercept = r_sf13_lines, linetype = "dashed")
+r_sf13_mean_break
+ggsave("r_sf13_break.png")
+```
+
+We tested leaf number and shape factor for 2015.
+```{r}
+r_sf15_plot <- ggplot(r_15, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_sf15_plot <- r_sf15_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+r_sf15_plot
+```
+
+```{r}
+r_sf15_mean_plot <- ggplot(r_sf15_mean, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
+r_sf15_mean_plot <- r_sf15_mean_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+r_sf15_mean_plot
+```
+
+We ran a LM for leaf number and shape factor in 2015.
+```{r}
+r_sf15_lm <- lm(avg.sf ~ leaf, data = r_sf15_mean)
+summary(r_sf15_lm)
+```
+
+```{r}
+r_sf15_seg <- segmented(r_sf15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_sf15_seg)
+```
+
+```{r}
+r_sf15_seg$psi
+slope(r_sf15_seg)
+r_sf15_fitted <- fitted(r_sf15_seg)
+r_sf15_model <- data.frame(Leaf = r_sf15_mean$leaf, SF = r_sf15_fitted)
+ggplot(r_sf15_model, aes(x = Leaf, y = SF)) + geom_line()
+r_sf15_lines <- r_sf15_seg$psi[, 2]
+r_sf15_mean_break <- r_sf15_plot + geom_vline(xintercept = r_sf15_lines, linetype = "dashed")
+r_sf15_mean_break
+ggsave("r_sf13_break.png")
+```

### 2018-11-08 V_riparia_LM.Rmd
V. riparia complete

@@
 
-We tested leaf number and shape factor for 2013.
+We tested leaf number and tooth area:internal perimeter for 2013.
 ```{r}
@@
 r_taip15_mean_break
-ggsave("r_taip13_break.png")
+ggsave("r_taip15_break.png")
+```
+
+We tested leaf number and shape factor for 2013.
+```{r}
+r_taba13_plot <- ggplot(r_13, aes(x = leaf, y = tooth.area.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_taba13_plot <- r_taba13_plot +labs(x = "Leaf number",
+                               y = "Tooth area:blade area")
+r_taba13_plot
+```
+
+```{r}
+r_taba13_mean_plot <- ggplot(r_taba13_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
+r_taba13_mean_plot <- r_taba13_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:blade area")
+r_taba13_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:blade area in 2013.
+```{r}
+r_taba13_lm <- lm(avg.taba ~ leaf, data = r_taba13_mean)
+summary(r_taba13_lm)
+```
+
+```{r}
+r_taba13_seg <- segmented(r_taba13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_taba13_seg)
+```
+
+```{r}
+r_taba13_seg$psi
+slope(r_taba13_seg)
+r_taba13_fitted <- fitted(r_taba13_seg)
+r_taba13_model <- data.frame(Leaf = r_taba13_mean$leaf, TABA = r_taba13_fitted)
+ggplot(r_taba13_model, aes(x = Leaf, y = TABA)) + geom_line()
+r_taba13_lines <- r_taba13_seg$psi[, 2]
+r_taba13_mean_break <- r_taba13_plot + geom_vline(xintercept = r_taba13_lines, linetype = "dashed")
+r_taba13_mean_break
+ggsave("r_taba13_break.png")
+```
+
+We tested leaf number and tooth area:blade area for 2015.
+```{r}
+r_taba15_plot <- ggplot(r_15, aes(x = leaf, y = tooth.area.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_taba15_plot <- r_taba15_plot +labs(x = "Leaf number",
+                               y = "Tooth area:blade area")
+r_taba15_plot
+```
+
+```{r}
+r_taba15_mean_plot <- ggplot(r_taba15_mean, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
+r_taba15_mean_plot <- r_taba15_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:blade area")
+r_taba15_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:blade area in 2015.
+```{r}
+r_taba15_lm <- lm(avg.taba ~ leaf, data = r_taba15_mean)
+summary(r_taba15_lm)
+```
+
+```{r}
+r_taba15_seg <- segmented(r_taba15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_taba15_seg)
+```
+
+```{r}
+r_taba15_seg$psi
+slope(r_taba15_seg)
+r_taba15_fitted <- fitted(r_taba15_seg)
+r_taba15_model <- data.frame(Leaf = r_taba15_mean$leaf, TABA = r_taba15_fitted)
+ggplot(r_taba15_model, aes(x = Leaf, y = TABA)) + geom_line()
+r_taba15_lines <- r_taba15_seg$psi[, 2]
+r_taba15_mean_break <- r_taba15_plot + geom_vline(xintercept = r_taba15_lines, linetype = "dashed")
+r_taba15_mean_break
+ggsave("r_taba15_break.png")
+```
+
+We tested leaf number and total teeth:perimeter for 2013.
+```{r}
+r_tp13_plot <- ggplot(r_13, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_tp13_plot <- r_tp13_plot +labs(x = "Leaf number",
+                               y = "Total teeth:perimeter")
+r_tp13_plot
+```
+
+```{r}
+r_tp13_mean_plot <- ggplot(r_tp13_mean, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
+r_tp13_mean_plot <- r_tp13_mean_plot +labs(x = "Leaf number",
+                               y = "Total teeth:perimeter")
+r_tp13_mean_plot
+```
+
+We ran a LM for leaf number and total teeth:perimeter in 2013.
+```{r}
+r_tp13_lm <- lm(avg.tp ~ leaf, data = r_tp13_mean)
+summary(r_tp13_lm)
+```
+
+```{r}
+r_tp13_seg <- segmented(r_tp13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_tp13_seg)
+```
+
+```{r}
+r_tp13_seg$psi
+slope(r_tp13_seg)
+r_tp13_fitted <- fitted(r_tp13_seg)
+r_tp13_model <- data.frame(Leaf = r_tp13_mean$leaf, TP = r_tp13_fitted)
+ggplot(r_tp13_model, aes(x = Leaf, y = TP)) + geom_line()
+r_tp13_lines <- r_tp13_seg$psi[, 2]
+r_tp13_mean_break <- r_tp13_plot + geom_vline(xintercept = r_tp13_lines, linetype = "dashed")
+r_tp13_mean_break
+ggsave("r_tp13_break.png")
+```
+
+We tested leaf number and total teeth:perimeter for 2015.
+```{r}
+r_tp15_plot <- ggplot(r_15, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_tp15_plot <- r_tp15_plot +labs(x = "Leaf number",
+                               y = "Total teeth:perimeter")
+r_tp15_plot
+```
+
+```{r}
+r_tp15_mean_plot <- ggplot(r_tp15_mean, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
+r_tp15_mean_plot <- r_tp15_mean_plot +labs(x = "Leaf number",
+                               y = "Total teeth:perimeter")
+r_tp15_mean_plot
+```
+
+We ran a LM for leaf number and total teeth:perimeter in 2015.
+```{r}
+r_tp15_lm <- lm(avg.tp ~ leaf, data = r_tp15_mean)
+summary(r_tp15_lm)
+```
+
+```{r}
+r_tp15_seg <- segmented(r_tp15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_tp15_seg)
+```
+
+```{r}
+r_tp15_seg$psi
+slope(r_tp15_seg)
+r_tp15_fitted <- fitted(r_tp15_seg)
+r_tp15_model <- data.frame(Leaf = r_tp15_mean$leaf, TP = r_tp15_fitted)
+ggplot(r_tp15_model, aes(x = Leaf, y = TP)) + geom_line()
+r_tp15_lines <- r_tp15_seg$psi[, 2]
+r_tp15_mean_break <- r_tp15_plot + geom_vline(xintercept = r_tp15_lines, linetype = "dashed")
+r_tp15_mean_break
+ggsave("r_tp15_break.png")
+```
+
+We tested leaf number and total teeth:internal perimeter for 2013.
+```{r}
+r_tip13_plot <- ggplot(r_13, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_tip13_plot <- r_tip13_plot +labs(x = "Leaf number",
+                               y = "Total teeth:internal perimeter")
+r_tip13_plot
+```
+
+```{r}
+r_tip13_mean_plot <- ggplot(r_tip13_mean, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
+r_tip13_mean_plot <- r_tip13_mean_plot +labs(x = "Leaf number",
+                               y = "Total teeth:internal perimeter")
+r_tip13_mean_plot
+```
+
+We ran a LM for leaf number and total teeth:internal perimeter in 2013.
+```{r}
+r_tip13_lm <- lm(avg.tip ~ leaf, data = r_tip13_mean)
+summary(r_tip13_lm)
+```
+
+```{r}
+r_tip13_seg <- segmented(r_tip13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_tip13_seg)
+```
+
+```{r}
+r_tip13_seg$psi
+slope(r_tip13_seg)
+r_tip13_fitted <- fitted(r_tip13_seg)
+r_tip13_model <- data.frame(Leaf = r_tip13_mean$leaf, TIP = r_tip13_fitted)
+ggplot(r_tip13_model, aes(x = Leaf, y = TIP)) + geom_line()
+r_tip13_lines <- r_tip13_seg$psi[, 2]
+r_tip13_mean_break <- r_tip13_plot + geom_vline(xintercept = r_tip13_lines, linetype = "dashed")
+r_tip13_mean_break
+ggsave("r_tip13_break.png")
+```
+
+We tested leaf number and total teeth:internal perimeter for 2015.
+```{r}
+r_tip15_plot <- ggplot(r_15, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_tip15_plot <- r_tip15_plot +labs(x = "Leaf number",
+                               y = "Total teeth:internal perimeter")
+r_tip15_plot
+```
+
+```{r}
+r_tip15_mean_plot <- ggplot(r_tip15_mean, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
+r_tip15_mean_plot <- r_tip15_mean_plot +labs(x = "Leaf number",
+                               y = "Total teeth:internal perimeter")
+r_tip15_mean_plot
+```
+
+We ran a LM for leaf number and total teeth:perimeter in 2015.
+```{r}
+r_tip15_lm <- lm(avg.tip ~ leaf, data = r_tip15_mean)
+summary(r_tip15_lm)
+```
+
+```{r}
+r_tip15_seg <- segmented(r_tip15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_tip15_seg)
+```
+
+```{r}
+r_tip15_seg$psi
+slope(r_tip15_seg)
+r_tip15_fitted <- fitted(r_tip15_seg)
+r_tip15_model <- data.frame(Leaf = r_tip15_mean$leaf, TIP = r_tip15_fitted)
+ggplot(r_tip15_model, aes(x = Leaf, y = TIP)) + geom_line()
+r_tip15_lines <- r_tip15_seg$psi[, 2]
+r_tip15_mean_break <- r_tip15_plot + geom_vline(xintercept = r_tip15_lines, linetype = "dashed")
+r_tip15_mean_break
+ggsave("r_tip15_break.png")
 ```
@@
 r_sf15_mean_break
-ggsave("r_sf13_break.png")
+ggsave("r_sf15_break.png")
 ```
+
+We tested leaf number and perimeter ratio for 2013.
+```{r}
+r_pr13_plot <- ggplot(r_13, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_pr13_plot <- r_pr13_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+r_pr13_plot
+```
+
+```{r}
+r_pr13_mean_plot <- ggplot(r_pr13_mean, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
+r_pr13_mean_plot <- r_pr13_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+r_pr13_mean_plot
+```
+
+We ran a LM for leaf number and perimeter ratio in 2013.
+```{r}
+r_pr13_lm <- lm(avg.pr ~ leaf, data = r_pr13_mean)
+summary(r_pr13_lm)
+```
+
+```{r}
+r_pr13_seg <- segmented(r_pr13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 11)))
+summary(r_pr13_seg)
+```
+
+```{r}
+r_pr13_seg$psi
+slope(r_pr13_seg)
+r_pr13_fitted <- fitted(r_pr13_seg)
+r_pr13_model <- data.frame(Leaf = r_pr13_mean$leaf, PR = r_pr13_fitted)
+ggplot(r_pr13_model, aes(x = Leaf, y = PR)) + geom_line()
+r_pr13_lines <- r_pr13_seg$psi[, 2]
+r_pr13_mean_break <- r_pr13_plot + geom_vline(xintercept = r_pr13_lines, linetype = "dashed")
+r_pr13_mean_break
+ggsave("r_pr13_break.png")
+```
+
+We tested leaf number and perimeter ratio for 2015.
+```{r}
+r_pr15_plot <- ggplot(r_15, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_pr15_plot <- r_pr15_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+r_pr15_plot
+```
+
+```{r}
+r_pr15_mean_plot <- ggplot(r_pr15_mean, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
+r_pr15_mean_plot <- r_pr15_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+r_pr15_mean_plot
+```
+
+We ran a LM for leaf number and perimeter ratio in 2015.
+```{r}
+r_pr15_lm <- lm(avg.pr ~ leaf, data = r_pr15_mean)
+summary(r_pr15_lm)
+```
+
+```{r}
+r_pr15_seg <- segmented(r_pr15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_pr15_seg)
+```
+
+```{r}
+r_pr15_seg$psi
+slope(r_pr15_seg)
+r_pr15_fitted <- fitted(r_pr15_seg)
+r_pr15_model <- data.frame(Leaf = r_pr15_mean$leaf, PR = r_pr15_fitted)
+ggplot(r_pr15_model, aes(x = Leaf, y = PR)) + geom_line()
+r_pr15_lines <- r_pr15_seg$psi[, 2]
+r_pr15_mean_break <- r_pr15_plot + geom_vline(xintercept = r_pr15_lines, linetype = "dashed")
+r_pr15_mean_break
+ggsave("r_pr15_break.png")
+```
+
+We tested leaf number and compactness for 2013.
+```{r}
+r_cp13_plot <- ggplot(r_13, aes(x = leaf, y = compactness)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_cp13_plot <- r_cp13_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+r_cp13_plot
+```
+
+```{r}
+r_cp13_mean_plot <- ggplot(r_cp13_mean, aes(x = leaf, y = avg.cp)) + geom_line() + theme_classic()
+r_cp13_mean_plot <- r_cp13_mean_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+r_cp13_mean_plot
+```
+
+We ran a LM for leaf number and compactness in 2013.
+```{r}
+r_cp13_lm <- lm(avg.cp ~ leaf, data = r_cp13_mean)
+summary(r_cp13_lm)
+```
+
+```{r}
+r_cp13_seg <- segmented(r_cp13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 11)))
+summary(r_cp13_seg)
+```
+
+```{r}
+r_cp13_seg$psi
+slope(r_cp13_seg)
+r_cp13_fitted <- fitted(r_cp13_seg)
+r_cp13_model <- data.frame(Leaf = r_cp13_mean$leaf, CP = r_cp13_fitted)
+ggplot(r_cp13_model, aes(x = Leaf, y = CP)) + geom_line()
+r_cp13_lines <- r_cp13_seg$psi[, 2]
+r_cp13_mean_break <- r_cp13_plot + geom_vline(xintercept = r_cp13_lines, linetype = "dashed")
+r_cp13_mean_break
+ggsave("r_cp13_break.png")
+```
+
+We tested leaf number and compactness for 2015.
+```{r}
+r_cp15_plot <- ggplot(r_15, aes(x = leaf, y = compactness)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_cp15_plot <- r_cp15_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+r_cp15_plot
+```
+
+```{r}
+r_cp15_mean_plot <- ggplot(r_cp15_mean, aes(x = leaf, y = avg.cp)) + geom_line() + theme_classic()
+r_cp15_mean_plot <- r_cp15_mean_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+r_cp15_mean_plot
+```
+
+We ran a LM for leaf number and perimeter ratio in 2015.
+```{r}
+r_cp15_lm <- lm(avg.cp ~ leaf, data = r_cp15_mean)
+summary(r_cp15_lm)
+```
+
+```{r}
+r_cp15_seg <- segmented(r_cp15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(r_cp15_seg)
+```
+
+```{r}
+r_cp15_seg$psi
+slope(r_cp15_seg)
+r_cp15_fitted <- fitted(r_cp15_seg)
+r_cp15_model <- data.frame(Leaf = r_cp15_mean$leaf, CP = r_cp15_fitted)
+ggplot(r_cp15_model, aes(x = Leaf, y = CP)) + geom_line()
+r_cp15_lines <- r_cp15_seg$psi[, 2]
+r_cp15_mean_break <- r_cp15_plot + geom_vline(xintercept = r_cp15_lines, linetype = "dashed")
+r_cp15_mean_break
+ggsave("r_cp15_break.png")
+```

### 2018-12-04 V_acerifolia_LM.Rmd
Interpret LM

@@
 summary(ac_tt_lm)
+ac_tt_res <- resid(ac_tt_lm)
+plot(avg.tt ~leaf, ac_l_means,
+     ylab="Residuals", xlab="leaf",
+     main="Total teeth")
 ```
+The average leaf in the dataset has 47 teeth and the number of teeth decreases by one as leaves get older.
+There is a relationship between leaf number and total teeth and about half of the variance in total teeth can be explained by leaf number.
 
@@
 ac_fdr_plot
-
 ```
@@
 ```
+The average leaf in the dataset has an FDR of 0.7 and the FDR increases by ~0.009 as the leaves get older.
+There is a relationship between leaf number and FDR and about 75% of the variance in FDR can be explained by leaf number.
 
@@
 ac_tap_plot
-
 ```
@@
 ```
+
+The average leaf in the dataset has a TAP of 0.04 and the TAP increases by 0.003 as the leaves get older.
+There is a relationship between leaf number and TAP and about 34% of the variance in TAP can be explained by leaf number.
 
@@
 ac_taip_plot
-
 ```
@@
 ```
+
+The average leaf in the dataset has a TAIP of 0.06 and the TAIP increases by 0.003 as the leaves get older.
+There is not a strong relationship between leaf number and TAIP and about 22% of the variance in TAP can be explained by leaf number.
 
@@
 
+The average leaf in the dataset has a ATA of 0.05 and the TAIP increases by 0.007 as the leaves get older.
+There is a weak relationship between leaf number and ATA and about 32% of the variance in ATA can be explained by leaf number.
+
 ```{r}
@@
 ```
+
+The average leaf in the dataset has a TP of 1.12 and the TP decreases by 0.04 as the leaves get older.
+There is a relationship between leaf number and TP and about 30% of the variance in TP can be explained by leaf number.
 
@@
 
+The average leaf in the dataset has a TIP of 1.67 and the TIP decreases by 0.07 as the leaves get older.
+There is a relationship between leaf number and TIP and about 45% of the variance in TIP can be explained by leaf number.
+
 ```{r}
@@
 ```
+
+The average leaf in the dataset has a PA of 2.33 and the PA decreases by 0.14 as the leaves get older.
+There is a relationship between leaf number and PA and about 41% of the variance in PA can be explained by leaf number.
 
@@
 
+The average leaf in the dataset has a PR of 1.5 and the PR decreases by 0.02 as the leaves get older.
+There is a relationship between leaf number and PR and about 68% of the variance in PR can be explained by leaf number.
+
 ```{r}
@@
 ```
+
+The average leaf in the dataset has a CP of 92 and the CP decreases by 4.3 as the leaves get older.
+There is a strong relationship between leaf number and CP and about 75% of the variance in PR can be explained by leaf number.
 
@@
 
+The average leaf in the dataset has a SF of 0.12 and the SF increases by 0.01 as the leaves get older.
+There is a strong relationship between leaf number and SF and about 95% of the variance in SF can be explained by leaf number.
+
 ```{r}
@@
 
+The average leaf in the dataset has a TBA of 2.98 and the TBA decreases by 0.23 as the leaves get older.
+There is a strong relationship between leaf number and TBA and about 40% of the variance in TBA can be explained by leaf number.
+
 ```{r}
@@
 ```
+
+The average leaf in the dataset has a TABA of 0.069 and the TABA decreases by 0.0038 as the leaves get older.
+There is not a relationship between leaf number and TABA and about 0% of the variance in TBA can be explained by leaf number.
 
@@
 
+The average leaf in the dataset has a TABA of 0.081 and the TABA decreases by 0.0014 as the leaves get older.
+There is a weak relationship between leaf number and TABA and about 0% of the variance in TBA can be explained by leaf number.
+
 ```{r}
@@
                        seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
+                       psi = list(leaf = c(5, 7, 10)))
 summary(ac_taba15_seg)

### 2018-12-07 V_acerifolia_LM.Rmd
acerifolia and aestivalis remove averages

@@
 ```{r}
-ac_tt_lm <- lm(avg.tt ~ leaf, data = ac_l_means)
+ac_tt_lm <- lm(total.teeth ~ leaf, data = acerifolia)
 summary(ac_tt_lm)
-ac_tt_res <- resid(ac_tt_lm)
-plot(avg.tt ~leaf, ac_l_means,
-     ylab="Residuals", xlab="leaf",
-     main="Total teeth")
 ```
-The average leaf in the dataset has 47 teeth and the number of teeth decreases by one as leaves get older.
-There is a relationship between leaf number and total teeth and about half of the variance in total teeth can be explained by leaf number.
+The average leaf in the dataset has 48 teeth and the number of teeth decreases by 1.2 as leaves get older.
+There is a relationship between leaf number and total teeth and about 20% of the variance in total teeth can be explained by leaf number.
 
@@
 ```{r}
-skim(acerifolia, total.teeth)
 ac_fdr_plot <- ggplot(acerifolia, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
@@
 
-We ran a LM for leaf number and feret diamter ratio.
+We ran a LM for leaf number and all feret diamter ratio.
 ```{r}
-ac_fdr_lm <- lm(avg.fdr ~ leaf, data = ac_l_means)
+ac_fdr_lm <- lm(feret.diameter.ratio ~ leaf, data = acerifolia)
 summary(ac_fdr_lm)
 ```
-The average leaf in the dataset has an FDR of 0.7 and the FDR increases by ~0.009 as the leaves get older.
-There is a relationship between leaf number and FDR and about 75% of the variance in FDR can be explained by leaf number.
+
+The average leaf in the dataset has a FDR of 0.7 and the FDR increases by 0.008 as the leaves get older.
+There is a relationship between leaf number and FDR and about 49% of the variance in TAP can be explained by leaf number.
 
@@
 ac_fdr_fitted <- fitted(ac_fdr_seg)
-ac_fdr_model <- data.frame(Leaf = ac_l_means$leaf, FDR = ac_fdr_fitted)
+ac_fdr_model <- data.frame(Leaf = acerifolia$leaf, FDR = ac_fdr_fitted)
 ggplot(ac_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
@@
 ```{r}
-ac_tap_lm <- lm(avg.tap ~ leaf, data = ac_l_means)
+ac_tap_lm <- lm(tooth.area.perimeter ~ leaf, data = acerifolia)
 summary(ac_tap_lm)
@@
 The average leaf in the dataset has a TAP of 0.04 and the TAP increases by 0.003 as the leaves get older.
-There is a relationship between leaf number and TAP and about 34% of the variance in TAP can be explained by leaf number.
+There is a relationship between leaf number and TAP and about 20% of the variance in TAP can be explained by leaf number.
 
@@
 ac_tap_fitted <- fitted(ac_tap_seg)
-ac_tap_model <- data.frame(Leaf = ac_l_means$leaf, TAP = ac_tap_fitted)
+ac_tap_model <- data.frame(Leaf = acerifolia$leaf, TAP = ac_tap_fitted)
 ggplot(ac_tap_model, aes(x = Leaf, y = TAP)) + geom_line()
@@
 ```{r}
-ac_taip_lm <- lm(avg.taip ~ leaf, data = ac_l_means)
+ac_taip_lm <- lm(tooth.area.int.perimeter ~ leaf, data = acerifolia)
 summary(ac_taip_lm)
@@
 The average leaf in the dataset has a TAIP of 0.06 and the TAIP increases by 0.003 as the leaves get older.
-There is not a strong relationship between leaf number and TAIP and about 22% of the variance in TAP can be explained by leaf number.
+There is a strong relationship between leaf number and TAIP and about 12% of the variance in TAP can be explained by leaf number.
 
@@
 ac_taip_fitted <- fitted(ac_taip_seg)
-ac_taip_model <- data.frame(Leaf = ac_l_means$leaf, TAIP = ac_taip_fitted)
+ac_taip_model <- data.frame(Leaf = acerifolia$leaf, TAIP = ac_taip_fitted)
 ggplot(ac_taip_model, aes(x = Leaf, y = TAIP)) + geom_line()
@@
 ```{r}
-ac_ta_lm <- lm(avg.ta ~ leaf, data = ac_l_means)
+ac_ta_lm <- lm(average.tooth.area ~ leaf, data = acerifolia)
 summary(ac_ta_lm)
@@
 
-The average leaf in the dataset has a ATA of 0.05 and the TAIP increases by 0.007 as the leaves get older.
-There is a weak relationship between leaf number and ATA and about 32% of the variance in ATA can be explained by leaf number.
+The average leaf in the dataset has a ATA of 0.05 and the TAIP increases by 0.008 as the leaves get older.
+There is a strong relationship between leaf number and ATA and about 18% of the variance in ATA can be explained by leaf number.
 
@@
 ac_ta_fitted <- fitted(ac_ta_seg)
-ac_ta_model <- data.frame(Leaf = ac_l_means$leaf, TA = ac_ta_fitted)
+ac_ta_model <- data.frame(Leaf = acerifolia$leaf, TA = ac_ta_fitted)
 ggplot(ac_ta_model, aes(x = Leaf, y = TA)) + geom_line()
@@
 ```{r}
-ac_tp_lm <- lm(avg.tp ~ leaf, data = ac_l_means)
+ac_tp_lm <- lm(teeth.perimeter ~ leaf, data = acerifolia)
 summary(ac_tp_lm)
@@
 
-The average leaf in the dataset has a TP of 1.12 and the TP decreases by 0.04 as the leaves get older.
-There is a relationship between leaf number and TP and about 30% of the variance in TP can be explained by leaf number.
+The average leaf in the dataset has a TP of 1.08 and the TP decreases by 0.04 as the leaves get older.
+There is a strong relationship between leaf number and TP and about 14% of the variance in TP can be explained by leaf number.
 
@@
 ac_tp_fitted <- fitted(ac_tp_seg)
-ac_tp_model <- data.frame(Leaf = ac_l_means$leaf, TP = ac_tp_fitted)
+ac_tp_model <- data.frame(Leaf = acerifolia$leaf, TP = ac_tp_fitted)
 ggplot(ac_tp_model, aes(x = Leaf, y = TP)) + geom_line()
@@
 ```{r}
-ac_tip_lm <- lm(avg.tip ~ leaf, data = ac_l_means)
+ac_tip_lm <- lm(teeth.int.perimeter ~ leaf, data = acerifolia)
 summary(ac_tip_lm)
@@
 The average leaf in the dataset has a TIP of 1.67 and the TIP decreases by 0.07 as the leaves get older.
-There is a relationship between leaf number and TIP and about 45% of the variance in TIP can be explained by leaf number.
+There is a strong relationship between leaf number and TIP and about 22% of the variance in TIP can be explained by leaf number.
 
@@
 ac_tip_fitted <- fitted(ac_tip_seg)
-ac_tip_model <- data.frame(Leaf = ac_l_means$leaf, TIP = ac_tip_fitted)
+ac_tip_model <- data.frame(Leaf = acerifolia$leaf, TIP = ac_tip_fitted)
 ggplot(ac_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
@@
 ```{r}
-ac_pa_lm <- lm(avg.pa ~ leaf, data = ac_l_means)
+ac_pa_lm <- lm(perimeter.area ~ leaf, data = acerifolia)
 summary(ac_pa_lm)
@@
 
-The average leaf in the dataset has a PA of 2.33 and the PA decreases by 0.14 as the leaves get older.
-There is a relationship between leaf number and PA and about 41% of the variance in PA can be explained by leaf number.
+The average leaf in the dataset has a PA of 2.16 and the PA decreases by 0.13 as the leaves get older.
+There is a relationship between leaf number and PA and about 26% of the variance in PA can be explained by leaf number.
 
@@
 ac_pa_fitted <- fitted(ac_pa_seg)
-ac_pa_model <- data.frame(Leaf = ac_l_means$leaf, PA = ac_pa_fitted)
+ac_pa_model <- data.frame(Leaf = acerifolia$leaf, PA = ac_pa_fitted)
 ggplot(ac_pa_model, aes(x = Leaf, y = PA)) + geom_line()
@@
 ```{r}
-ac_pr_lm <- lm(avg.pr ~ leaf, data = ac_l_means)
+ac_pr_lm <- lm(perimeter.ratio ~ leaf, data = acerifolia)
 summary(ac_pr_lm)
@@
 The average leaf in the dataset has a PR of 1.5 and the PR decreases by 0.02 as the leaves get older.
-There is a relationship between leaf number and PR and about 68% of the variance in PR can be explained by leaf number.
+There is a relationship between leaf number and PR and about 18% of the variance in PR can be explained by leaf number.
 
@@
 ac_pr_fitted <- fitted(ac_pr_seg)
-ac_pr_model <- data.frame(Leaf = ac_l_means$leaf, PR = ac_pr_fitted)
+ac_pr_model <- data.frame(Leaf = acerifolia$leaf, PR = ac_pr_fitted)
 ggplot(ac_pr_model, aes(x = Leaf, y = PR)) + geom_line()
@@
 ```{r}
-ac_cp_lm <- lm(avg.cp ~ leaf, data = ac_l_means)
+ac_cp_lm <- lm(compactness ~ leaf, data = acerifolia)
 summary(ac_cp_lm)
@@
 
-The average leaf in the dataset has a CP of 92 and the CP decreases by 4.3 as the leaves get older.
-There is a strong relationship between leaf number and CP and about 75% of the variance in PR can be explained by leaf number.
+The average leaf in the dataset has a CP of 91 and the CP decreases by 4.3 as the leaves get older.
+There is a strong relationship between leaf number and CP and about 47% of the variance in PR can be explained by leaf number.
 
@@
 ac_cp_fitted <- fitted(ac_cp_seg)
-ac_cp_model <- data.frame(Leaf = ac_l_means$leaf, CP = ac_cp_fitted)
+ac_cp_model <- data.frame(Leaf = acerifolia$leaf, CP = ac_cp_fitted)
 ggplot(ac_cp_model, aes(x = Leaf, y = CP)) + geom_line()
@@
 ```{r}
-ac_sf_lm <- lm(avg.sf ~ leaf, data = ac_l_means)
+ac_sf_lm <- lm(shape.factor ~ leaf, data = acerifolia)
 summary(ac_sf_lm)
@@
 The average leaf in the dataset has a SF of 0.12 and the SF increases by 0.01 as the leaves get older.
-There is a strong relationship between leaf number and SF and about 95% of the variance in SF can be explained by leaf number.
+There is a strong relationship between leaf number and SF and about 48% of the variance in SF can be explained by leaf number.
 
@@
 ac_sf_fitted <- fitted(ac_sf_seg)
-ac_sf_model <- data.frame(Leaf = ac_l_means$leaf, SF = ac_sf_fitted)
+ac_sf_model <- data.frame(Leaf = acerifolia$leaf, SF = ac_sf_fitted)
 ggplot(ac_sf_model, aes(x = Leaf, y = SF)) + geom_line()
@@
 ```{r}
-ac_tba_lm <- lm(avg.tba ~ leaf, data = ac_l_means)
+ac_tba_lm <- lm(teeth.blade.area ~ leaf, data = acerifolia)
 summary(ac_tba_lm)
@@
 
-The average leaf in the dataset has a TBA of 2.98 and the TBA decreases by 0.23 as the leaves get older.
-There is a strong relationship between leaf number and TBA and about 40% of the variance in TBA can be explained by leaf number.
+The average leaf in the dataset has a TBA of 2.69 and the TBA decreases by 0.21 as the leaves get older.
+There is a strong relationship between leaf number and TBA and about 19% of the variance in TBA can be explained by leaf number.
 
@@
 ac_tba_fitted <- fitted(ac_tba_seg)
-ac_tba_model <- data.frame(Leaf = ac_l_means$leaf, TBA = ac_tba_fitted)
+ac_tba_model <- data.frame(Leaf = acerifolia$leaf, TBA = ac_tba_fitted)
 ggplot(ac_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
@@
 ```{r}
-ac_taba13_lm <- lm(avg.taba ~ leaf, data = ac_taba13_mean)
+ac_taba13_lm <- lm(teeth.blade.area ~ leaf, data = ac_13)
 summary(ac_taba13_lm)
@@
 ac_taba13_fitted <- fitted(ac_taba13_seg)
-ac_taba13_model <- data.frame(Leaf = ac_taba13_mean$leaf, TABA = ac_taba13_fitted)
+ac_taba13_model <- data.frame(Leaf = ac_13$leaf, TABA = ac_taba13_fitted)
 ggplot(ac_taba13_model, aes(x = Leaf, y = TABA)) + geom_line()
@@
 ```{r}
-ac_taba15_lm <- lm(avg.taba ~ leaf, data = ac_taba15_mean)
+ac_taba15_lm <- lm(teeth.blade.area ~ leaf, data = ac_15)
 summary(ac_taba15_lm)
@@
 ac_taba15_fitted <- fitted(ac_taba15_seg)
-ac_taba15_model <- data.frame(Leaf = ac_taba15_mean$leaf, TABA = ac_taba15_fitted)
+ac_taba15_model <- data.frame(Leaf = ac_15$leaf, TABA = ac_taba15_fitted)
 ggplot(ac_taba15_model, aes(x = Leaf, y = TABA)) + geom_line()

### 2018-12-07 V_aestivalis_LM.Rmd
acerifolia and aestivalis remove averages

@@
             avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
-            avg.tp = mean(teeth.perimeter),
+            avg.tp = mean(teeth.perimeter, na.rm = TRUE),
             avg.tip = mean(teeth.int.perimeter),
@@
 ```{r}
-av_fdr_lm <- lm(avg.fdr ~ leaf, data = av_l_means)
+av_fdr_lm <- lm(feret.diameter.ratio ~ leaf, data = aestivalis)
 summary(av_fdr_lm)
@@
 av_fdr_fitted <- fitted(av_fdr_seg)
-av_fdr_model <- data.frame(Leaf = av_l_means$leaf, FDR = av_fdr_fitted)
+av_fdr_model <- data.frame(Leaf = aestivalis$leaf, FDR = av_fdr_fitted)
 ggplot(av_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
@@
 ```{r}
-av_tap_lm <- lm(avg.tap ~ leaf, data = av_l_means)
+av_tap_lm <- lm(tooth.area.perimeter ~ leaf, data = aestivalis)
 summary(av_tap_lm)
@@
 ```{r}
-av_taip_lm <- lm(avg.taip ~ leaf, data = av_l_means)
+av_taip_lm <- lm(tooth.area.int.perimeter ~ leaf, data = aestivalis)
 summary(av_taip_lm)
@@
 av_taip_fitted <- fitted(av_taip_seg)
-av_taip_model <- data.frame(Leaf = av_l_means$leaf, TAIP = av_taip_fitted)
+av_taip_model <- data.frame(Leaf = aestivalis$leaf, TAIP = av_taip_fitted)
 ggplot(av_taip_model, aes(x = Leaf, y = TAIP)) + geom_line()
@@
 ```{r}
-av_ta_lm <- lm(avg.ta ~ leaf, data = av_l_means)
+av_ta_lm <- lm(average.tooth.area ~ leaf, data = aestivalis)
 summary(av_ta_lm)
@@
 av_ta_fitted <- fitted(av_ta_seg)
-av_ta_model <- data.frame(Leaf = av_l_means$leaf, TA = av_ta_fitted)
+av_ta_model <- data.frame(Leaf = aestivalis$leaf, TA = av_ta_fitted)
 ggplot(av_ta_model, aes(x = Leaf, y = TA)) + geom_line()
@@
 ```{r}
-av_tp_lm <- lm(avg.tp ~ leaf, data = av_l_means)
+av_tp_lm <- lm(teeth.perimeter ~ leaf, data = aestivalis)
 summary(av_tp_lm)
@@
 av_tp_fitted <- fitted(av_tp_seg)
-av_tp_model <- data.frame(Leaf = av_l_means$leaf, TP = av_tp_fitted)
+av_tp_model <- data.frame(Leaf = aestivalis$leaf, TP = av_tp_fitted)
 ggplot(av_tp_model, aes(x = Leaf, y = TP)) + geom_line()
@@
 ```{r}
-av_tip_lm <- lm(avg.tip ~ leaf, data = av_l_means)
+av_tip_lm <- lm(teeth.int.perimeter ~ leaf, data = aestivalis)
 summary(av_tip_lm)
@@
 av_tip_fitted <- fitted(av_tip_seg)
-av_tip_model <- data.frame(Leaf = av_l_means$leaf, TIP = av_tip_fitted)
+av_tip_model <- data.frame(Leaf = aestivalis$leaf, TIP = av_tip_fitted)
 ggplot(av_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
@@
 ```{r}
-av_pa_lm <- lm(avg.pa ~ leaf, data = av_l_means)
+av_pa_lm <- lm(perimeter.area ~ leaf, data = aestivalis)
 summary(av_pa_lm)
@@
                        seg.Z = ~ leaf,
-                       psi = list(leaf = c(2, 7, 10)))
+                       psi = list(leaf = c(5, 7, 10)))
 summary(av_pa_seg)
@@
 ```{r}
-av_pr_lm <- lm(avg.pr ~ leaf, data = av_l_means)
+av_pr_lm <- lm(perimeter.ratio ~ leaf, data = aestivalis)
 summary(av_pr_lm)
@@
                        seg.Z = ~ leaf,
-                       psi = list(leaf = c(3, 7, 10)))
+                       psi = list(leaf = c(3, 6, 10)))
 summary(av_pr_seg)
@@
 ```{r}
-av_tba_lm <- lm(avg.tba ~ leaf, data = av_l_means)
+av_tba_lm <- lm(teeth.blade.area ~ leaf, data = aestivalis)
 summary(av_tba_lm)
@@
 av_taba13_plot
-
 ```
@@
 ```{r}
-av_taba13_lm <- lm(avg.taba ~ leaf, data = av_taba13_mean)
+av_taba13_lm <- lm(tooth.area.blade.area ~ leaf, data = av_13)
 summary(av_taba13_lm)
@@
 av_taba13_fitted <- fitted(av_taba13_seg)
-av_taba13_model <- data.frame(Leaf = av_taba13_mean$leaf, TABA = av_taba13_fitted)
+av_taba13_model <- data.frame(Leaf = av_13$leaf, TABA = av_taba13_fitted)
 ggplot(av_taba13_model, aes(x = Leaf, y = TABA)) + geom_line()
@@
 ```{r}
-av_taba15_lm <- lm(avg.taba ~ leaf, data = av_taba15_mean)
+av_taba15_lm <- lm(tooth.area.blade.area ~ leaf, data = av_15)
 summary(av_taba15_lm)
@@
 av_taba15_fitted <- fitted(av_taba15_seg)
-av_taba15_model <- data.frame(Leaf = av_taba15_mean$leaf, TABA = av_taba15_fitted)
+av_taba15_model <- data.frame(Leaf = av_15$leaf, TABA = av_taba15_fitted)
 ggplot(av_taba15_model, aes(x = Leaf, y = TABA)) + geom_line()
@@
 ```{r}
-av_tt13_lm <- lm(avg.tt ~ leaf, data = av_tt13_mean)
+av_tt13_lm <- lm(total.teeth ~ leaf, data = av_13)
 summary(av_tt13_lm)
@@
 av_tt13_fitted <- fitted(av_tt13_seg)
-av_tt13_model <- data.frame(Leaf = av_tt13_mean$leaf, TT = av_tt13_fitted)
+av_tt13_model <- data.frame(Leaf = av_13$leaf, TT = av_tt13_fitted)
 ggplot(av_tt13_model, aes(x = Leaf, y = TT)) + geom_line()
@@
 ```{r}
-av_tt15_lm <- lm(avg.tt ~ leaf, data = av_tt15_mean)
+av_tt15_lm <- lm(total.teeth ~ leaf, data = av_15)
 summary(av_tt15_lm)
@@
 av_tt15_fitted <- fitted(av_tt15_seg)
-av_tt15_model <- data.frame(Leaf = av_tt15_mean$leaf, TT = av_tt15_fitted)
+av_tt15_model <- data.frame(Leaf = av_15$leaf, TT = av_tt15_fitted)
 ggplot(av_tt15_model, aes(x = Leaf, y = TT)) + geom_line()
@@
 av_cp13_plot
-
 ```
@@
 ```{r}
-av_cp13_lm <- lm(avg.cp ~ leaf, data = av_cp13_mean)
+av_cp13_lm <- lm(compactness ~ leaf, data = av_13)
 summary(av_cp13_lm)
@@
 av_cp13_fitted <- fitted(av_cp13_seg)
-av_cp13_model <- data.frame(Leaf = av_cp13_mean$leaf, CP = av_cp13_fitted)
+av_cp13_model <- data.frame(Leaf = av_13$leaf, CP = av_cp13_fitted)
 ggplot(av_cp13_model, aes(x = Leaf, y = CP)) + geom_line()
@@
 ```{r}
-av_cp15_lm <- lm(avg.cp ~ leaf, data = av_cp15_mean)
+av_cp15_lm <- lm(compactness ~ leaf, data = av_15)
 summary(av_cp15_lm)
@@
 av_cp15_fitted <- fitted(av_cp15_seg)
-av_cp15_model <- data.frame(Leaf = av_cp15_mean$leaf, CP = av_cp15_fitted)
+av_cp15_model <- data.frame(Leaf = av_15$leaf, CP = av_cp15_fitted)
 ggplot(av_cp15_model, aes(x = Leaf, y = CP)) + geom_line()
@@
 ```{r}
-av_sf13_lm <- lm(avg.sf ~ leaf, data = av_sf13_mean)
+av_sf13_lm <- lm(shape.factor ~ leaf, data = av_13)
 summary(av_sf13_lm)
@@
 av_sf13_fitted <- fitted(av_sf13_seg)
-av_sf13_model <- data.frame(Leaf = av_sf13_mean$leaf, SF = av_sf13_fitted)
+av_sf13_model <- data.frame(Leaf = av_13$leaf, SF = av_sf13_fitted)
 ggplot(av_sf13_model, aes(x = Leaf, y = SF)) + geom_line()
@@
 ```{r}
-av_sf15_lm <- lm(avg.sf ~ leaf, data = av_sf15_mean)
+av_sf15_lm <- lm(shape.factor ~ leaf, data = av_15)
 summary(av_sf15_lm)
@@
 av_sf15_fitted <- fitted(av_sf15_seg)
-av_sf15_model <- data.frame(Leaf = av_sf15_mean$leaf, SF = av_sf15_fitted)
+av_sf15_model <- data.frame(Leaf = av_15$leaf, SF = av_sf15_fitted)
 ggplot(av_sf15_model, aes(x = Leaf, y = SF)) + geom_line()

### 2018-12-07 V_amurensis_LM.Rmd
amurensis remove averages

--- 2018-10-29 V_amurensis_LM.Rmd
@@
 ```{r}
-am_fdr_lm <- lm(avg.fdr ~ leaf, data = am_l_means)
+am_fdr_lm <- lm(feret.diameter.ratio ~ leaf, data = amurensis)
 summary(am_fdr_lm)
@@
                        seg.Z = ~ leaf,
-                       psi = list(leaf = c(4, 7, 11)))
+                       psi = list(leaf = c(3, 7, 10)))
 summary(am_fdr_seg)
@@
 am_fdr_fitted <- fitted(am_fdr_seg)
-am_fdr_model <- data.frame(Leaf = am_l_means$leaf, FDR = am_fdr_fitted)
+am_fdr_model <- data.frame(Leaf = amurensis$leaf, FDR = am_fdr_fitted)
 ggplot(am_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
@@
 ```
+
+We tested leaf number and tooth area:blade area.
+```{r}
+am_taba_plot <- ggplot(amurensis, aes(x = leaf, y = tooth.area.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_taba_plot <- am_taba_plot +labs(x = "Leaf number",
+                               y = "Tooth area:blade area")
+am_taba_plot
+```
+
+```{r}
+am_taba_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
+am_taba_mean_plot <- am_taba_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:blade area")
+am_taba_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:blade area.
+```{r}
+am_taba_lm <- lm(tooth.area.blade.area ~ leaf, data = amurensis)
+summary(am_taba_lm)
+```
+
+```{r}
+am_taba_seg <- segmented(am_taba_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_taba_seg)
+```
+
+```{r}
+am_taba_seg$psi
+slope(am_taba_seg)
+am_taba_fitted <- fitted(am_taba_seg)
+am_taba_model <- data.frame(Leaf = amurensis$leaf, TABA = am_taba_fitted)
+ggplot(am_taba_model, aes(x = Leaf, y = TABA)) + geom_line()
+am_taba_lines <- am_taba_seg$psi[, 2]
+am_taba_mean_break <- am_taba_plot + geom_vline(xintercept = am_taba_lines, linetype = "dashed")
+am_taba_mean_break
+ggsave("am_taba_break.png")
+```
+
+We tested leaf number and total teeth:perimeter.
+```{r}
+am_tp_plot <- ggplot(amurensis, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_tp_plot <- am_tp_plot +labs(x = "Leaf number",
+                               y = "Teeth:perimeter")
+am_tp_plot
+```
+
+```{r}
+am_tp_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
+am_tp_mean_plot <- am_tp_mean_plot +labs(x = "Leaf number",
+                               y = "Teeth.perimeter")
+am_tp_mean_plot
+```
+
+We ran a LM for leaf number and teeth:perimeter.
+```{r}
+am_tp_lm <- lm(teeth.perimeter ~ leaf, data = amurensis)
+summary(am_tp_lm)
+```
+
+```{r}
+am_tp_seg <- segmented(am_tp_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_tp_seg)
+```
+
+```{r}
+am_tp_seg$psi
+slope(am_tp_seg)
+am_tp_fitted <- fitted(am_tp_seg)
+am_tp_model <- data.frame(Leaf = amurensis$leaf, TP = am_tp_fitted)
+ggplot(am_tp_model, aes(x = Leaf, y = TP)) + geom_line()
+am_tp_lines <- am_tp_seg$psi[, 2]
+am_tp_mean_break <- am_tp_plot + geom_vline(xintercept = am_tp_lines, linetype = "dashed")
+am_tp_mean_break
+ggsave("am_tp_break.png")
+```
+
+We tested leaf number and total teeth:internal perimeter.
+```{r}
+am_tip_plot <- ggplot(amurensis, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_tip_plot <- am_tip_plot +labs(x = "Leaf number",
+                               y = "Teeth:internal perimeter")
+am_tip_plot
+```
+
+```{r}
+am_tip_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
+am_tip_mean_plot <- am_tip_mean_plot +labs(x = "Leaf number",
+                               y = "Teeth.internal perimeter")
+am_tip_mean_plot
+```
+
+We ran a LM for leaf number and teeth:internal perimeter.
+```{r}
+am_tip_lm <- lm(teeth.int.perimeter ~ leaf, data = amurensis)
+summary(am_tip_lm)
+```
+
+```{r}
+am_tip_seg <- segmented(am_tip_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_tip_seg)
+```
+
+```{r}
+am_tip_seg$psi
+slope(am_tip_seg)
+am_tip_fitted <- fitted(am_tip_seg)
+am_tip_model <- data.frame(Leaf = amurensis$leaf, TIP = am_tip_fitted)
+ggplot(am_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
+am_tip_lines <- am_tip_seg$psi[, 2]
+am_tip_mean_break <- am_tip_plot + geom_vline(xintercept = am_tip_lines, linetype = "dashed")
+am_tip_mean_break
+ggsave("am_tip_break.png")
+```
+
+We tested leaf number and perimeter:area.
+```{r}
+am_pa_plot <- ggplot(amurensis, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_pa_plot <- am_pa_plot +labs(x = "Leaf number",
+                               y = "Perimeter:area")
+am_pa_plot
+```
+
+```{r}
+am_pa_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
+am_pa_mean_plot <- am_pa_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter:area")
+am_pa_mean_plot
+```
+
+We ran a LM for leaf number and perimeter:area.
+```{r}
+am_pa_lm <- lm(perimeter.area ~ leaf, data = amurensis)
+summary(am_pa_lm)
+```
+
+```{r}
+am_pa_seg <- segmented(am_pa_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_pa_seg)
+```
+
+```{r}
+am_pa_seg$psi
+slope(am_pa_seg)
+am_pa_fitted <- fitted(am_pa_seg)
+am_pa_model <- data.frame(Leaf = amurensis$leaf, PA = am_pa_fitted)
+ggplot(am_pa_model, aes(x = Leaf, y = PA)) + geom_line()
+am_pa_lines <- am_pa_seg$psi[, 2]
+am_pa_mean_break <- am_pa_plot + geom_vline(xintercept = am_pa_lines, linetype = "dashed")
+am_pa_mean_break
+ggsave("am_pa_break.png")
+```
+
+We tested leaf number and compactness.
+```{r}
+am_cp_plot <- ggplot(amurensis, aes(x = leaf, y = compactness)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_cp_plot <- am_cp_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+am_cp_plot
+```
+
+```{r}
+am_cp_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.cp)) + geom_line() + theme_classic()
+am_cp_mean_plot <- am_cp_mean_plot +labs(x = "Leaf number",
+                               y = "Compactness")
+am_cp_mean_plot
+```
+
+We ran a LM for leaf number and compactness.
+```{r}
+am_cp_lm <- lm(compactness ~ leaf, data = amurensis)
+summary(am_cp_lm)
+```
+
+```{r}
+am_cp_seg <- segmented(am_cp_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_cp_seg)
+```
+
+```{r}
+am_cp_seg$psi
+slope(am_cp_seg)
+am_cp_fitted <- fitted(am_cp_seg)
+am_cp_model <- data.frame(Leaf = amurensis$leaf, CP = am_cp_fitted)
+ggplot(am_cp_model, aes(x = Leaf, y = CP)) + geom_line()
+am_cp_lines <- am_cp_seg$psi[, 2]
+am_cp_mean_break <- am_cp_plot + geom_vline(xintercept = am_cp_lines, linetype = "dashed")
+am_cp_mean_break
+ggsave("am_cp_break.png")
+```
+
+We tested leaf number and total teeth:blade are.
+```{r}
+am_tba_plot <- ggplot(amurensis, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_tba_plot <- am_tba_plot +labs(x = "Leaf number",
+                               y = "Total Teeth:Blade Area")
+am_tba_plot
+```
+
+```{r}
+am_tba_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
+am_tba_mean_plot <- am_tba_mean_plot +labs(x = "Leaf number",
+                               y = "Total Teeth:Blade Area")
+am_tba_mean_plot
+```
+
+We ran a LM for leaf number and total teeth:blade area.
+```{r}
+am_tba_lm <- lm(teeth.blade.area ~ leaf, data = amurensis)
+summary(am_tba_lm)
+```
+
+```{r}
+am_tba_seg <- segmented(am_tba_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 8, 11)))
+summary(am_tba_seg)
+```
+
+```{r}
+am_tba_seg$psi
+slope(am_tba_seg)
+am_tba_fitted <- fitted(am_tba_seg)
+am_tba_model <- data.frame(Leaf = amurensis$leaf, TBA = am_tba_fitted)
+ggplot(am_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
+am_tba_lines <- am_tba_seg$psi[, 2]
+am_tba_mean_break <- am_tba_plot + geom_vline(xintercept = am_tba_lines, linetype = "dashed")
+am_tba_mean_break
+ggsave("am_tba_break.png")
+```
+
+We tested leaf number and total teeth for 2013.
+```{r}
+am_tt13_plot <- ggplot(am_13, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_tt13_plot <- am_tt13_plot +labs(x = "Leaf number",
+                               y = "Total Teeth")
+am_tt13_plot
+```
+
+```{r}
+am_tt13_mean_plot <- ggplot(am_tt13_mean, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
+am_tt13_mean_plot <- am_tt13_mean_plot +labs(x = "Leaf number",
+                               y = "Total Teeth")
+am_tt13_mean_plot
+```
+
+We ran a LM for leaf number and total teeth in 2013.
+```{r}
+am_tt13_lm <- lm(total.teeth ~ leaf, data = am_13)
+summary(am_tt13_lm)
+```
+
+```{r}
+am_tt13_seg <- segmented(am_tt13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_tt13_seg)
+```
+
+```{r}
+am_tt13_seg$psi
+slope(am_tt13_seg)
+am_tt13_fitted <- fitted(am_tt13_seg)
+am_tt13_model <- data.frame(Leaf = am_13$leaf, TT = am_tt13_fitted)
+ggplot(am_tt13_model, aes(x = Leaf, y = TT)) + geom_line()
+am_tt13_lines <- am_tt13_seg$psi[, 2]
+am_tt13_mean_break <- am_tt13_plot + geom_vline(xintercept = am_tt13_lines, linetype = "dashed")
+am_tt13_mean_break
+ggsave("am_tt13_break.png")
+```
+
+We tested leaf number and total teeth for 2015.
+```{r}
+am_tt15_plot <- ggplot(am_15, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_tt15_plot <- am_tt15_plot +labs(x = "Leaf number",
+                               y = "Total Teeth")
+am_tt15_plot
+```
+
+```{r}
+am_tt15_mean_plot <- ggplot(am_tt15_mean, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
+am_tt15_mean_plot <- am_tt15_mean_plot +labs(x = "Leaf number",
+                               y = "Total Teeth")
+am_tt15_mean_plot
+```
+
+We ran a LM for leaf number and total teeth in 2015.
+```{r}
+am_tt15_lm <- lm(total.teeth ~ leaf, data = am_15)
+summary(am_tt15_lm)
+```
+
+```{r}
+am_tt15_seg <- segmented(am_tt15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_tt15_seg)
+```
+
+```{r}
+am_tt15_seg$psi
+slope(am_tt15_seg)
+am_tt15_fitted <- fitted(am_tt15_seg)
+am_tt15_model <- data.frame(Leaf = am_15$leaf, TT = am_tt15_fitted)
+ggplot(am_tt15_model, aes(x = Leaf, y = TT)) + geom_line()
+am_tt15_lines <- am_tt15_seg$psi[, 2]
+am_tt15_mean_break <- am_tt15_plot + geom_vline(xintercept = am_tt15_lines, linetype = "dashed")
+am_tt15_mean_break
+ggsave("am_tt15_break.png")
+```
+
+We tested leaf number and tooth area:perimeter for 2013.
+```{r}
+am_tap13_plot <- ggplot(am_13, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_tap13_plot <- am_tap13_plot +labs(x = "Leaf number",
+                               y = "Tooth area: perimeter")
+am_tap13_plot
+```
+
+```{r}
+am_tap13_mean_plot <- ggplot(am_tap13_mean, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
+am_tap13_mean_plot <- am_tap13_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area: perimeter")
+am_tap13_mean_plot
+```
+
+We ran a LM for leaf number and toth area perimeter in 2013.
+```{r}
+am_tap13_lm <- lm(tooth.area.perimeter ~ leaf, data = am_13)
+summary(am_tap13_lm)
+```
+
+```{r}
+am_tap13_seg <- segmented(am_tap13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_tap13_seg)
+```
+
+```{r}
+am_tap13_seg$psi
+slope(am_tap13_seg)
+am_tap13_fitted <- fitted(am_tap13_seg)
+am_tap13_model <- data.frame(Leaf = am_13$leaf, TAP = am_tap13_fitted)
+ggplot(am_tap13_model, aes(x = Leaf, y = TAP)) + geom_line()
+am_tap13_lines <- am_tap13_seg$psi[, 2]
+am_tap13_mean_break <- am_tap13_plot + geom_vline(xintercept = am_tap13_lines, linetype = "dashed")
+am_tap13_mean_break
+ggsave("am_tap13_break.png")
+```
+
+We tested leaf number and tooth area:perimeter for 2015.
+```{r}
+am_tap15_plot <- ggplot(am_15, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_tap15_plot <- am_tap15_plot +labs(x = "Leaf number",
+                               y = "Tooth area:perimeter")
+am_tap15_plot
+```
+
+```{r}
+am_tap15_mean_plot <- ggplot(am_tap15_mean, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
+am_tap15_mean_plot <- am_tap15_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:perimeter")
+am_tap15_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:perimeter in 2015.
+```{r}
+am_tap15_lm <- lm(tooth.area.perimeter ~ leaf, data = am_15)
+summary(am_tap15_lm)
+```
+
+```{r}
+am_tap15_seg <- segmented(am_tap15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_tap15_seg)
+```
+
+```{r}
+am_tap15_seg$psi
+slope(am_tap15_seg)
+am_tap15_fitted <- fitted(am_tap15_seg)
+am_tap15_model <- data.frame(Leaf = am_15$leaf, TAP = am_tap15_fitted)
+ggplot(am_tap15_model, aes(x = Leaf, y = TAP)) + geom_line()
+am_tap15_lines <- am_tap15_seg$psi[, 2]
+am_tap15_mean_break <- am_tap15_plot + geom_vline(xintercept = am_tap15_lines, linetype = "dashed")
+am_tap15_mean_break
+ggsave("am_tap15_break.png")
+```
+
+We tested leaf number and tooth area:internal perimeter for 2013.
+```{r}
+am_taip13_plot <- ggplot(am_13, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_taip13_plot <- am_taip13_plot +labs(x = "Leaf number",
+                               y = "Tooth area: internal perimeter")
+am_taip13_plot
+```
+
+```{r}
+am_taip13_mean_plot <- ggplot(am_taip13_mean, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
+am_taip13_mean_plot <- am_taip13_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:internal perimeter")
+am_taip13_mean_plot
+```
+
+We ran a LM for leaf number and toth area: internal perimeter in 2013.
+```{r}
+am_taip13_lm <- lm(tooth.area.int.perimeter ~ leaf, data = am_13)
+summary(am_taip13_lm)
+```
+
+```{r}
+am_taip13_seg <- segmented(am_taip13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_taip13_seg)
+```
+
+```{r}
+am_taip13_seg$psi
+slope(am_taip13_seg)
+am_taip13_fitted <- fitted(am_taip13_seg)
+am_taip13_model <- data.frame(Leaf = am_13$leaf, TAIP = am_taip13_fitted)
+ggplot(am_taip13_model, aes(x = Leaf, y = TAIP)) + geom_line()
+am_taip13_lines <- am_taip13_seg$psi[, 2]
+am_taip13_mean_break <- am_taip13_plot + geom_vline(xintercept = am_taip13_lines, linetype = "dashed")
+am_taip13_mean_break
+ggsave("am_taip13_break.png")
+```
+
+We tested leaf number and tooth area:internal perimeter for 2015.
+```{r}
+am_taip15_plot <- ggplot(am_15, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_taip15_plot <- am_taip15_plot +labs(x = "Leaf number",
+                               y = "Tooth area:internal perimeter")
+am_taip15_plot
+```
+
+```{r}
+am_taip15_mean_plot <- ggplot(am_taip15_mean, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
+am_taip15_mean_plot <- am_taip15_mean_plot +labs(x = "Leaf number",
+                               y = "Tooth area:internal perimeter")
+am_taip15_mean_plot
+```
+
+We ran a LM for leaf number and tooth area:internal perimeter in 2015.
+```{r}
+am_taip15_lm <- lm(tooth.area.int.perimeter ~ leaf, data = am_15)
+summary(am_taip15_lm)
+```
+
+```{r}
+am_taip15_seg <- segmented(am_taip15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_taip15_seg)
+```
+
+```{r}
+am_taip15_seg$psi
+slope(am_taip15_seg)
+am_taip15_fitted <- fitted(am_taip15_seg)
+am_taip15_model <- data.frame(Leaf = am_15$leaf, TAIP = am_taip15_fitted)
+ggplot(am_taip15_model, aes(x = Leaf, y = TAIP)) + geom_line()
+am_taip15_lines <- am_taip15_seg$psi[, 2]
+am_taip15_mean_break <- am_taip15_plot + geom_vline(xintercept = am_taip15_lines, linetype = "dashed")
+am_taip15_mean_break
+ggsave("am_taip15_break.png")
+```
+
+We tested leaf number and average tooth area for 2013.
+```{r}
+am_ata13_plot <- ggplot(am_13, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_ata13_plot <- am_ata13_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+am_ata13_plot
+```
+
+```{r}
+am_ata13_mean_plot <- ggplot(am_ata13_mean, aes(x = leaf, y = avg.ata)) + geom_line() + theme_classic()
+am_ata13_mean_plot <- am_ata13_mean_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+am_ata13_mean_plot
+```
+
+We ran a LM for leaf number and average tooth area in 2013.
+```{r}
+am_ata13_lm <- lm(average.tooth.area ~ leaf, data = am_13)
+summary(am_ata13_lm)
+```
+
+```{r}
+am_ata13_seg <- segmented(am_ata13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_ata13_seg)
+```
+
+```{r}
+am_ata13_seg$psi
+slope(am_ata13_seg)
+am_ata13_fitted <- fitted(am_ata13_seg)
+am_ata13_model <- data.frame(Leaf = am_13$leaf, ATA = am_ata13_fitted)
+ggplot(am_ata13_model, aes(x = Leaf, y = ATA)) + geom_line()
+am_ata13_lines <- am_ata13_seg$psi[, 2]
+am_ata13_mean_break <- am_ata13_plot + geom_vline(xintercept = am_ata13_lines, linetype = "dashed")
+am_ata13_mean_break
+ggsave("am_ata13_break.png")
+```
+
+We tested leaf number and average tooth area for 2015.
+```{r}
+am_ata15_plot <- ggplot(am_15, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_ata15_plot <- am_ata15_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+am_ata15_plot
+```
+
+```{r}
+am_ata15_mean_plot <- ggplot(am_ata15_mean, aes(x = leaf, y = avg.ata)) + geom_line() + theme_classic()
+am_ata15_mean_plot <- am_ata15_mean_plot +labs(x = "Leaf number",
+                               y = "Average tooth area")
+am_ata15_mean_plot
+```
+
+We ran a LM for leaf number and average tooth area in 2015.
+```{r}
+am_ata15_lm <- lm(average.tooth.area ~ leaf, data = am_15)
+summary(am_ata15_lm)
+```
+
+```{r}
+am_ata15_seg <- segmented(am_ata15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_ata15_seg)
+```
+
+```{r}
+am_ata15_seg$psi
+slope(am_ata15_seg)
+am_ata15_fitted <- fitted(am_ata15_seg)
+am_ata15_model <- data.frame(Leaf = am_15$leaf, ATA = am_ata15_fitted)
+ggplot(am_ata15_model, aes(x = Leaf, y = ATA)) + geom_line()
+am_ata15_lines <- am_ata15_seg$psi[, 2]
+am_ata15_mean_break <- am_ata15_plot + geom_vline(xintercept = am_ata15_lines, linetype = "dashed")
+am_ata15_mean_break
+ggsave("am_ata15_break.png")
+```
+
+We tested leaf number and perimeter ratio for 2013.
+```{r}
+am_pr13_plot <- ggplot(am_13, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_pr13_plot <- am_pr13_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+am_pr13_plot
+```
+
+```{r}
+am_pr13_mean_plot <- ggplot(am_pr13_mean, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
+am_pr13_mean_plot <- am_pr13_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+am_pr13_mean_plot
+```
+
+We ran a LM for leaf number and perimeter ratio in 2013.
+```{r}
+am_pr13_lm <- lm(perimeter.ratio ~ leaf, data = am_13)
+summary(am_pr13_lm)
+```
+
+```{r}
+am_pr13_seg <- segmented(am_pr13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_pr13_seg)
+```
+
+```{r}
+am_pr13_seg$psi
+slope(am_pr13_seg)
+am_pr13_fitted <- fitted(am_pr13_seg)
+am_pr13_model <- data.frame(Leaf = am_13$leaf, PR = am_pr13_fitted)
+ggplot(am_pr13_model, aes(x = Leaf, y = PR)) + geom_line()
+am_pr13_lines <- am_pr13_seg$psi[, 2]
+am_pr13_mean_break <- am_pr13_plot + geom_vline(xintercept = am_pr13_lines, linetype = "dashed")
+am_pr13_mean_break
+ggsave("am_pr13_break.png")
+```
+
+We tested leaf number and perimeter ratio for 2015.
+```{r}
+am_pr15_plot <- ggplot(am_15, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_pr15_plot <- am_pr15_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+am_pr15_plot
+```
+
+```{r}
+am_pr15_mean_plot <- ggplot(am_pr15_mean, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
+am_pr15_mean_plot <- am_pr15_mean_plot +labs(x = "Leaf number",
+                               y = "Perimeter ratio")
+am_pr15_mean_plot
+```
+
+We ran a LM for leaf number and average tooth area in 2015.
+```{r}
+am_pr15_lm <- lm(perimeter.ratio ~ leaf, data = am_15)
+summary(am_pr15_lm)
+```
+
+```{r}
+am_pr15_seg <- segmented(am_pr15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_pr15_seg)
+```
+
+```{r}
+am_pr15_seg$psi
+slope(am_pr15_seg)
+am_pr15_fitted <- fitted(am_pr15_seg)
+am_pr15_model <- data.frame(Leaf = am_15$leaf, PR = am_pr15_fitted)
+ggplot(am_pr15_model, aes(x = Leaf, y = PR)) + geom_line()
+am_pr15_lines <- am_pr15_seg$psi[, 2]
+am_pr15_mean_break <- am_pr15_plot + geom_vline(xintercept = am_pr15_lines, linetype = "dashed")
+am_pr15_mean_break
+ggsave("am_pr15_break.png")
+```
+
+We tested leaf number and shape factor for 2013.
+```{r}
+am_sf13_plot <- ggplot(am_13, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_sf13_plot <- am_sf13_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+am_sf13_plot
+```
+
+```{r}
+am_sf13_mean_plot <- ggplot(am_sf13_mean, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
+am_sf13_mean_plot <- am_sf13_mean_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+am_sf13_mean_plot
+```
+
+We ran a LM for leaf number and shape factor in 2013.
+```{r}
+am_sf13_lm <- lm(shape.factor ~ leaf, data = am_13)
+summary(am_sf13_lm)
+```
+
+```{r}
+am_sf13_seg <- segmented(am_sf13_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_sf13_seg)
+```
+
+```{r}
+am_sf13_seg$psi
+slope(am_sf13_seg)
+am_sf13_fitted <- fitted(am_sf13_seg)
+am_sf13_model <- data.frame(Leaf = am_13$leaf, SF = am_sf13_fitted)
+ggplot(am_sf13_model, aes(x = Leaf, y = SF)) + geom_line()
+am_sf13_lines <- am_sf13_seg$psi[, 2]
+am_sf13_mean_break <- am_sf13_plot + geom_vline(xintercept = am_sf13_lines, linetype = "dashed")
+am_sf13_mean_break
+ggsave("am_sf13_break.png")
+```
+
+We tested leaf number and shape factor for 2015.
+```{r}
+am_sf15_plot <- ggplot(am_15, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_sf15_plot <- am_sf15_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+am_sf15_plot
+```
+
+```{r}
+am_sf15_mean_plot <- ggplot(am_sf15_mean, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
+am_sf15_mean_plot <- am_sf15_mean_plot +labs(x = "Leaf number",
+                               y = "Shape factor")
+am_sf15_mean_plot
+```
+
+We ran a LM for leaf number and shape factor in 2015.
+```{r}
+am_sf15_lm <- lm(shape.factor ~ leaf, data = am_15)
+summary(am_sf15_lm)
+```
+
+```{r}
+am_sf15_seg <- segmented(am_sf15_lm,
+                       seg.Z = ~ leaf,
+                       psi = list(leaf = c(3, 7, 10)))
+summary(am_sf15_seg)
+```
+
+```{r}
+am_sf15_seg$psi
+slope(am_sf15_seg)
+am_sf15_fitted <- fitted(am_sf15_seg)
+am_sf15_model <- data.frame(Leaf = am_15$leaf, SF = am_ata15_fitted)
+ggplot(am_sf15_model, aes(x = Leaf, y = SF)) + geom_line()
+am_sf15_lines <- am_sf15_seg$psi[, 2]
+am_sf15_mean_break <- am_sf15_plot + geom_vline(xintercept = am_sf15_lines, linetype = "dashed")
+am_sf15_mean_break
+ggsave("am_sf15_break.png")
+```

### 2018-12-11 V_acerifolia_LM.Rmd
Vitis comparison

@@
 skim(acerifolia, total.teeth)
-ac_tt_plot <- ggplot(acerifolia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_tt_plot <- ggplot(acerifolia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + ylim(25, 100) + 
+  scale_x_continuous(breaks = seq(0, 12, 1))
 ac_tt_plot <- ac_tt_plot +labs(x = "Leaf number",
@@
 ac_tt_plot
-
+ggsave("ac_tt.png")
 ```
@@
 ```{r}
-ac_fdr_plot <- ggplot(acerifolia, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_fdr_plot <- ggplot(acerifolia, aes(x = leaf, y = feret.diameter.ratio)) + 
+  geom_smooth() + theme_classic() + ylim(0.65, 0.9) + scale_x_continuous(breaks = seq(0, 12, 1))
 ac_fdr_plot <- ac_fdr_plot +labs(x = "Leaf number",
@@
 ac_fdr_plot
+ggsave("ac_fdr.png")
 ```
@@
 skim(acerifolia, tooth.area.perimeter)
-ac_tap_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_tap_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.perimeter)) + 
+  geom_smooth() + theme_classic() + ylim(0, 0.1) + scale_x_continuous(breaks = seq(0, 12, 1))
 ac_tap_plot <- ac_tap_plot +labs(x = "Leaf number",
@@
 ac_tap_plot
+ggsave("ac_tap.png")
 ```
@@
 skim(acerifolia, tooth.area.int.perimeter)
-ac_taip_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_taip_plot <- ggplot(acerifolia, aes(x = leaf, y = tooth.area.int.perimeter)) + 
+  geom_smooth() + theme_classic() + ylim(0.02, 0.16) + scale_x_continuous(breaks = seq(0, 12, 1))
 ac_taip_plot <- ac_taip_plot +labs(x = "Leaf number",
@@
 ac_taip_plot
+ggsave("ac_taip.png")
 ```
@@
 skim(acerifolia, average.tooth.area)
-ac_ta_plot <- ggplot(acerifolia, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+ac_ta_plot <- ggplot(acerifolia, aes(x = leaf, y = average.tooth.area)) + 
+  geom_smooth() + theme_classic() + ylim(0, 0.25) + scale_x_continuous(breaks = seq(0, 12, 1))
 ac_ta_plot <- ac_ta_plot +labs(x = "Leaf number",
@@
 ac_ta_plot
-
+ggsave("ac_ata.png")
 ```

### 2018-12-11 V_aestivalis_LM.Rmd
Vitis comparison

@@
 ```{r}
-av_fdr_plot <- ggplot(aestivalis, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_fdr_plot <- ggplot(aestivalis, aes(x = leaf, y = feret.diameter.ratio)) + 
+  geom_smooth() + theme_classic() + ylim(0.65, 0.9) + scale_x_continuous(breaks = seq(0, 12, 1))
 av_fdr_plot <- av_fdr_plot +labs(x = "Leaf number",
@@
 av_fdr_plot
+ggsave("av_fdr.png")
 ```
@@
 ```{r}
-av_tap_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_tap_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.perimeter)) + 
+  geom_smooth() + theme_classic() + ylim(0, 0.1) + scale_x_continuous(breaks = seq(0, 12, 1))
 av_tap_plot <- av_tap_plot +labs(x = "Leaf number",
@@
 av_tap_plot
+ggsave("av_tap.png")
 ```
@@
 ```{r}
-av_taip_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_taip_plot <- ggplot(aestivalis, aes(x = leaf, y = tooth.area.int.perimeter)) + 
+  geom_smooth() + theme_classic() + ylim(0.02, 0.16) + scale_x_continuous(breaks = seq(0, 12, 1))
 av_taip_plot <- av_taip_plot +labs(x = "Leaf number",
@@
 av_taip_plot
+ggsave("av_taip.png")
 ```
@@
 ```{r}
-av_ta_plot <- ggplot(aestivalis, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_ta_plot <- ggplot(aestivalis, aes(x = leaf, y = average.tooth.area)) + 
+  geom_smooth() + theme_classic() + ylim(0, 0.25) + scale_x_continuous(breaks = seq(0, 12, 1))
 av_ta_plot <- av_ta_plot +labs(x = "Leaf number",
@@
 av_ta_plot
-
+ggsave("av_ata.png")
 ```
@@
 ```{r}
-av_tt13_plot <- ggplot(av_13, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_tt13_plot <- ggplot(av_13, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + ylim(25, 100) + scale_x_continuous(breaks = seq(0, 12, 1), limits = c(1, 12))
 av_tt13_plot <- av_tt13_plot +labs(x = "Leaf number",
@@
 av_tt13_plot
-
+ggsave("av_tt13.png")
 ```
@@
 ```{r}
-av_tt15_plot <- ggplot(av_15, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+av_tt15_plot <- ggplot(av_15, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + ylim(25, 100) + scale_x_continuous(breaks = seq(0, 12, 1), limits = c(1, 12))
 av_tt15_plot <- av_tt15_plot +labs(x = "Leaf number",
@@
 av_tt15_plot
-
+ggsave("av_tt15.png")
 ```

### 2018-12-11 V_amurensis_LM.Rmd
Vitis comparison

@@
 ```{r}
-am_fdr_plot <- ggplot(amurensis, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_fdr_plot <- ggplot(amurensis, aes(x = leaf, y = feret.diameter.ratio)) + 
+  geom_smooth() + theme_classic() + ylim(0.65, 0.9) + scale_x_continuous(breaks = seq(0, 13, 1))
 am_fdr_plot <- am_fdr_plot +labs(x = "Leaf number",
@@
 am_fdr_plot
-
+ggsave("am_fdr.png")
 ```
@@
 ```{r}
-am_tt13_plot <- ggplot(am_13, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_tt13_plot <- ggplot(am_13, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + ylim(25, 100) + scale_x_continuous(breaks = seq(0, 12, 1))
 am_tt13_plot <- am_tt13_plot +labs(x = "Leaf number",
@@
 am_tt13_plot
+ggsave("am_tt13.png")
 ```
@@
 ```{r}
-am_tt15_plot <- ggplot(am_15, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_tt15_plot <- ggplot(am_15, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + ylim(25, 100) + scale_x_continuous(breaks = seq(0, 12, 1))
 am_tt15_plot <- am_tt15_plot +labs(x = "Leaf number",
@@
 am_tt15_plot
+ggsave("am_tt15.png")
 ```
@@
 ```{r}
-am_tap13_plot <- ggplot(am_13, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_tap13_plot <- ggplot(am_13, aes(x = leaf, y = tooth.area.perimeter)) + 
+  geom_smooth() + theme_classic() + ylim(0, 0.1) + scale_x_continuous(breaks = seq(0, 13, 1))
 am_tap13_plot <- am_tap13_plot +labs(x = "Leaf number",
@@
 am_tap13_plot
+ggsave("am_tap13.png")
 ```
@@
 ```{r}
-am_tap15_plot <- ggplot(am_15, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_tap15_plot <- ggplot(am_15, aes(x = leaf, y = tooth.area.perimeter)) + 
+  geom_smooth() + theme_classic() + ylim(0, 0.1) + scale_x_continuous(breaks = seq(0, 12, 1))
 am_tap15_plot <- am_tap15_plot +labs(x = "Leaf number",
@@
 am_tap15_plot
+ggsave("am_tap15.png")
 ```
@@
 ```{r}
-am_taip13_plot <- ggplot(am_13, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
+am_taip13_plot <- ggplot(am_13, aes(x = leaf, y = tooth.area.int.perimeter)) + 
+  geom_smooth() + theme_classic() + ylim(0.02, 0.16) + scale_x_continuous(breaks = seq(0, 13, 1))
 am_taip13_plot <- am_taip13_plot +labs(x = "Leaf number",
@@
 am_taip13_plot
+ggsave("am_taip13.png")
 ```
@@
 ```{r}
-am_taip15_plot <- ggplot(am_15, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_taip15_plot <- ggplot(am_15, aes(x = leaf, y = tooth.area.int.perimeter)) + 
+  geom_smooth() + theme_classic() + ylim(0.02, 0.16) + scale_x_continuous(breaks = seq(0, 13, 1))
 am_taip15_plot <- am_taip15_plot +labs(x = "Leaf number",
@@
 am_taip15_plot
+ggsave("am_taip15.png")
 ```
@@
 ```{r}
-am_ata13_plot <- ggplot(am_13, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_ata13_plot <- ggplot(am_13, aes(x = leaf, y = average.tooth.area)) + 
+  geom_smooth() + theme_classic() + ylim(0, 0.25) + scale_x_continuous(breaks = seq(0, 13, 1))
 am_ata13_plot <- am_ata13_plot +labs(x = "Leaf number",
@@
 am_ata13_plot
+ggsave("am_ata13.png")
 ```
@@
 ```{r}
-am_ata15_plot <- ggplot(am_15, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
+am_ata15_plot <- ggplot(am_15, aes(x = leaf, y = average.tooth.area)) + 
+  geom_smooth() + theme_classic() + ylim(0, 0.25) + scale_x_continuous(breaks = seq(0, 13, 1))
 am_ata15_plot <- am_ata15_plot +labs(x = "Leaf number",
@@
 am_ata15_plot
+ggsave("am_ata15.png")
 ```

### 2018-12-11 V_riparia_LM.Rmd
Vitis comparison

@@
 ```{r}
-r_tt_plot <- ggplot(riparia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_tt_plot <- ggplot(riparia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + ylim(25, 100) + scale_x_continuous(breaks = seq(0, 14, 1))
 r_tt_plot <- r_tt_plot +labs(x = "Leaf number",
@@
 r_tt_plot
+ggsave("r_tt.png")
 ```
@@
 ```{r}
-r_tt_lm <- lm(avg.tt ~ leaf, data = r_l_means)
+r_tt_lm <- lm(total.teeth ~ leaf, data = riparia)
 summary(r_tt_lm)
@@
 r_tt_fitted <- fitted(r_tt_seg)
-r_tt_model <- data.frame(Leaf = r_l_means$leaf, TT = r_tt_fitted)
+r_tt_model <- data.frame(Leaf = riparia$leaf, TT = r_tt_fitted)
 ggplot(r_tt_model, aes(x = Leaf, y = TT)) + geom_line()
@@
 ```{r}
-r_fdr_plot <- ggplot(riparia, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_fdr_plot <- ggplot(riparia, aes(x = leaf, y = feret.diameter.ratio)) + 
+  geom_smooth() + theme_classic() + ylim(0.65, 0.9) + scale_x_continuous(breaks = seq(0, 14, 1))
 r_fdr_plot <- r_fdr_plot +labs(x = "Leaf number",
@@
 r_fdr_plot
+ggsave("r_fdr.png")
 ```
@@
 ```{r}
-r_fdr_lm <- lm(avg.fdr ~ leaf, data = r_l_means)
+r_fdr_lm <- lm(feret.diameter.ratio ~ leaf, data = riparia)
 summary(r_fdr_lm)
@@
 r_fdr_fitted <- fitted(r_fdr_seg)
-r_fdr_model <- data.frame(Leaf = r_l_means$leaf, FDR = r_fdr_fitted)
+r_fdr_model <- data.frame(Leaf = riparia$leaf, FDR = r_fdr_fitted)
 ggplot(r_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
@@
 ```{r}
-r_tap_plot <- ggplot(riparia, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_tap_plot <- ggplot(riparia, aes(x = leaf, y = tooth.area.perimeter)) + 
+  geom_smooth() + theme_classic() + ylim(0, 0.1) + scale_x_continuous(breaks = seq(0, 14, 1))
 r_tap_plot <- r_tap_plot +labs(x = "Leaf number",
@@
 r_tap_plot
+ggsave("r_tap.png")
 ```
@@
 ```{r}
-r_tap_lm <- lm(avg.tap ~ leaf, data = r_l_means)
+r_tap_lm <- lm(tooth.area.perimeter ~ leaf, data = riparia)
 summary(r_tap_lm)
@@
 r_tap_fitted <- fitted(r_tap_seg)
-r_tap_model <- data.frame(Leaf = r_l_means$leaf, TAP = r_tap_fitted)
+r_tap_model <- data.frame(Leaf = riparia$leaf, TAP = r_tap_fitted)
 ggplot(r_tap_model, aes(x = Leaf, y = TAP)) + geom_line()
@@
 ```{r}
-r_ata_plot <- ggplot(riparia, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_ata_plot <- ggplot(riparia, aes(x = leaf, y = average.tooth.area)) + 
+  geom_smooth() + theme_classic() + ylim(0, 0.25) + scale_x_continuous(breaks = seq(0, 14, 1))
 r_ata_plot <- r_ata_plot +labs(x = "Leaf number",
@@
 r_ata_plot
+ggsave("r_ata.png")
 ```
@@
 ```{r}
-r_ata_lm <- lm(avg.ata ~ leaf, data = r_l_means)
+r_ata_lm <- lm(average.tooth.area ~ leaf, data = riparia)
 summary(r_ata_lm)
@@
 r_ata_fitted <- fitted(r_ata_seg)
-r_ata_model <- data.frame(Leaf = r_l_means$leaf, ATA = r_ata_fitted)
+r_ata_model <- data.frame(Leaf = riparia$leaf, ATA = r_ata_fitted)
 ggplot(r_ata_model, aes(x = Leaf, y = ATA)) + geom_line()
@@
 ```{r}
-r_pa_lm <- lm(avg.pa ~ leaf, data = r_l_means)
+r_pa_lm <- lm(perimeter.area ~ leaf, data = riparia)
 summary(r_pa_lm)
@@
 r_pa_fitted <- fitted(r_pa_seg)
-r_pa_model <- data.frame(Leaf = r_l_means$leaf, PA = r_pa_fitted)
+r_pa_model <- data.frame(Leaf = riparia$leaf, PA = r_pa_fitted)
 ggplot(r_pa_model, aes(x = Leaf, y = PA)) + geom_line()
@@
 ```{r}
-r_tba_lm <- lm(avg.tba ~ leaf, data = r_l_means)
+r_tba_lm <- lm(teeth.blade.area ~ leaf, data = riparia)
 summary(r_tba_lm)
@@
 r_tba_fitted <- fitted(r_tba_seg)
-r_tba_model <- data.frame(Leaf = r_l_means$leaf, TBA = r_tba_fitted)
+r_tba_model <- data.frame(Leaf = riparia$leaf, TBA = r_tba_fitted)
 ggplot(r_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
@@
 ```{r}
-r_taip13_plot <- ggplot(r_13, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_taip13_plot <- ggplot(r_13, aes(x = leaf, y = tooth.area.int.perimeter)) + 
+  geom_smooth() + theme_classic() + ylim(0.02, 0.16) + scale_x_continuous(breaks = seq(0, 14, 1))
 r_taip13_plot <- r_taip13_plot +labs(x = "Leaf number",
@@
 r_taip13_plot
+ggsave("r_taip13.png")
 ```
@@
 ```{r}
-r_taip13_lm <- lm(avg.taip ~ leaf, data = r_taip13_mean)
+r_taip13_lm <- lm(tooth.area.int.perimeter ~ leaf, data = r_13)
 summary(r_taip13_lm)
@@
 r_taip13_fitted <- fitted(r_taip13_seg)
-r_taip13_model <- data.frame(Leaf = r_taip13_mean$leaf, TAIP = r_taip13_fitted)
+r_taip13_model <- data.frame(Leaf = r_13$leaf, TAIP = r_taip13_fitted)
 ggplot(r_taip13_model, aes(x = Leaf, y = TAIP)) + geom_line()
@@
 ```{r}
-r_taip15_plot <- ggplot(r_15, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 14, 1))
+r_taip15_plot <- ggplot(r_15, aes(x = leaf, y = tooth.area.int.perimeter)) + 
+  geom_smooth() + theme_classic() + ylim(0.02, 0.16) + scale_x_continuous(breaks = seq(0, 14, 1))
 r_taip15_plot <- r_taip15_plot +labs(x = "Leaf number",
@@
 r_taip15_plot
+ggsave("r_taip15.png")
 ```
@@
 ```{r}
-r_taip15_lm <- lm(avg.taip ~ leaf, data = r_taip15_mean)
+r_taip15_lm <- lm(tooth.area.int.perimeter ~ leaf, data = r_15)
 summary(r_taip15_lm)
@@
 r_taip15_fitted <- fitted(r_taip15_seg)
-r_taip15_model <- data.frame(Leaf = r_taip15_mean$leaf, TAIP = r_taip15_fitted)
+r_taip15_model <- data.frame(Leaf = r_15$leaf, TAIP = r_taip15_fitted)
 ggplot(r_taip15_model, aes(x = Leaf, y = TAIP)) + geom_line()
@@
 
-We tested leaf number and shape factor for 2013.
+We tested leaf number and tooth area:blade area for 2013.
 ```{r}
@@
 ```{r}
-r_taba13_lm <- lm(avg.taba ~ leaf, data = r_taba13_mean)
+r_taba13_lm <- lm(tooth.area.blade.area ~ leaf, data = r_13)
 summary(r_taba13_lm)
@@
 r_taba13_fitted <- fitted(r_taba13_seg)
-r_taba13_model <- data.frame(Leaf = r_taba13_mean$leaf, TABA = r_taba13_fitted)
+r_taba13_model <- data.frame(Leaf = r_13$leaf, TABA = r_taba13_fitted)
 ggplot(r_taba13_model, aes(x = Leaf, y = TABA)) + geom_line()
@@
 ```{r}
-r_taba15_lm <- lm(avg.taba ~ leaf, data = r_taba15_mean)
+r_taba15_lm <- lm(tooth.area.blade.area ~ leaf, data = r_15)
 summary(r_taba15_lm)
@@
 r_taba15_fitted <- fitted(r_taba15_seg)
-r_taba15_model <- data.frame(Leaf = r_taba15_mean$leaf, TABA = r_taba15_fitted)
+r_taba15_model <- data.frame(Leaf = r_15$leaf, TABA = r_taba15_fitted)
 ggplot(r_taba15_model, aes(x = Leaf, y = TABA)) + geom_line()
@@
 ```{r}
-r_tp13_lm <- lm(avg.tp ~ leaf, data = r_tp13_mean)
+r_tp13_lm <- lm(teeth.perimeter ~ leaf, data = r_13)
 summary(r_tp13_lm)
@@
 r_tp13_fitted <- fitted(r_tp13_seg)
-r_tp13_model <- data.frame(Leaf = r_tp13_mean$leaf, TP = r_tp13_fitted)
+r_tp13_model <- data.frame(Leaf = r_13$leaf, TP = r_tp13_fitted)
 ggplot(r_tp13_model, aes(x = Leaf, y = TP)) + geom_line()
@@
 ```{r}
-r_tp15_lm <- lm(avg.tp ~ leaf, data = r_tp15_mean)
+r_tp15_lm <- lm(teeth.perimeter ~ leaf, data = r_15)
 summary(r_tp15_lm)
@@
 r_tp15_fitted <- fitted(r_tp15_seg)
-r_tp15_model <- data.frame(Leaf = r_tp15_mean$leaf, TP = r_tp15_fitted)
+r_tp15_model <- data.frame(Leaf = r_15$leaf, TP = r_tp15_fitted)
 ggplot(r_tp15_model, aes(x = Leaf, y = TP)) + geom_line()
@@
 ```{r}
-r_tip13_lm <- lm(avg.tip ~ leaf, data = r_tip13_mean)
+r_tip13_lm <- lm(teeth.int.perimeter ~ leaf, data = r_13)
 summary(r_tip13_lm)
@@
 r_tip13_fitted <- fitted(r_tip13_seg)
-r_tip13_model <- data.frame(Leaf = r_tip13_mean$leaf, TIP = r_tip13_fitted)
+r_tip13_model <- data.frame(Leaf = r_13$leaf, TIP = r_tip13_fitted)
 ggplot(r_tip13_model, aes(x = Leaf, y = TIP)) + geom_line()
@@
 ```{r}
-r_tip15_lm <- lm(avg.tip ~ leaf, data = r_tip15_mean)
+r_tip15_lm <- lm(teeth.int.perimeter ~ leaf, data = r_15)
 summary(r_tip15_lm)
@@
 r_tip15_fitted <- fitted(r_tip15_seg)
-r_tip15_model <- data.frame(Leaf = r_tip15_mean$leaf, TIP = r_tip15_fitted)
+r_tip15_model <- data.frame(Leaf = r_15$leaf, TIP = r_tip15_fitted)
 ggplot(r_tip15_model, aes(x = Leaf, y = TIP)) + geom_line()
@@
 ```{r}
-r_sf13_lm <- lm(avg.sf ~ leaf, data = r_sf13_mean)
+r_sf13_lm <- lm(shape.factor ~ leaf, data = r_13)
 summary(r_sf13_lm)
@@
 r_sf13_fitted <- fitted(r_sf13_seg)
-r_sf13_model <- data.frame(Leaf = r_sf13_mean$leaf, SF = r_sf13_fitted)
+r_sf13_model <- data.frame(Leaf = r_13$leaf, SF = r_sf13_fitted)
 ggplot(r_sf13_model, aes(x = Leaf, y = SF)) + geom_line()
@@
 ```{r}
-r_sf15_lm <- lm(avg.sf ~ leaf, data = r_sf15_mean)
+r_sf15_lm <- lm(shape.factor ~ leaf, data = r_15)
 summary(r_sf15_lm)
@@
 r_sf15_fitted <- fitted(r_sf15_seg)
-r_sf15_model <- data.frame(Leaf = r_sf15_mean$leaf, SF = r_sf15_fitted)
+r_sf15_model <- data.frame(Leaf = r_15$leaf, SF = r_sf15_fitted)
 ggplot(r_sf15_model, aes(x = Leaf, y = SF)) + geom_line()
@@
 ```{r}
-r_pr13_lm <- lm(avg.pr ~ leaf, data = r_pr13_mean)
+r_pr13_lm <- lm(perimeter.ratio ~ leaf, data = r_13)
 summary(r_pr13_lm)
@@
 r_pr13_fitted <- fitted(r_pr13_seg)
-r_pr13_model <- data.frame(Leaf = r_pr13_mean$leaf, PR = r_pr13_fitted)
+r_pr13_model <- data.frame(Leaf = r_13$leaf, PR = r_pr13_fitted)
 ggplot(r_pr13_model, aes(x = Leaf, y = PR)) + geom_line()
@@
 ```{r}
-r_pr15_lm <- lm(avg.pr ~ leaf, data = r_pr15_mean)
+r_pr15_lm <- lm(perimeter.ratio ~ leaf, data = r_15)
 summary(r_pr15_lm)
@@
 r_pr15_fitted <- fitted(r_pr15_seg)
-r_pr15_model <- data.frame(Leaf = r_pr15_mean$leaf, PR = r_pr15_fitted)
+r_pr15_model <- data.frame(Leaf = r_15$leaf, PR = r_pr15_fitted)
 ggplot(r_pr15_model, aes(x = Leaf, y = PR)) + geom_line()
@@
 ```{r}
-r_cp13_lm <- lm(avg.cp ~ leaf, data = r_cp13_mean)
+r_cp13_lm <- lm(compactness ~ leaf, data = r_13)
 summary(r_cp13_lm)
@@
 r_cp13_fitted <- fitted(r_cp13_seg)
-r_cp13_model <- data.frame(Leaf = r_cp13_mean$leaf, CP = r_cp13_fitted)
+r_cp13_model <- data.frame(Leaf = r_13$leaf, CP = r_cp13_fitted)
 ggplot(r_cp13_model, aes(x = Leaf, y = CP)) + geom_line()
@@
 ```{r}
-r_cp15_lm <- lm(avg.cp ~ leaf, data = r_cp15_mean)
+r_cp15_lm <- lm(compactness ~ leaf, data = r_15)
 summary(r_cp15_lm)
@@
 r_cp15_fitted <- fitted(r_cp15_seg)
-r_cp15_model <- data.frame(Leaf = r_cp15_mean$leaf, CP = r_cp15_fitted)
+r_cp15_model <- data.frame(Leaf = r_15$leaf, CP = r_cp15_fitted)
 ggplot(r_cp15_model, aes(x = Leaf, y = CP)) + geom_line()
